---
title: "Weighted SDID"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weighted SDID}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(synthdid)
set.seed(42)
```

# Weighted estimation

The standard SDID estimator gives equal weight to all treated units.
When treated units differ in size, you may want a population-weighted estimate instead.
The `synthdid_estimate_weighted` function allows user-specified weights on treated units.

We demonstrate with simulated data where three treated units have different sizes and effects.

```{r simulate}
# Use the California Prop 99 dataset included in the package
data('california_prop99')
setup = panel.matrices(california_prop99)
N0 = setup$N0
N1 = nrow(setup$Y) - N0
T0 = setup$T0

# Population weights (e.g., some states are larger)
# Here we just illustrate with fabricated weights
set.seed(42)
pop = runif(N1, 1, 10)
pop_weights = pop / sum(pop)
```

# Compare equally-weighted and population-weighted estimates

```{r estimates}
tau_sdid = synthdid_estimate(setup$Y, N0, T0)
tau_sdid_w = synthdid_estimate_weighted(setup$Y, N0, T0,
               treated.weights = pop_weights)
tau_did = did_estimate(setup$Y, N0, T0)
tau_did_w = did_estimate_weighted(setup$Y, N0, T0,
              treated.weights = pop_weights)

cat(sprintf("Equally-weighted SDID:     %1.2f\n", tau_sdid))
cat(sprintf("Population-weighted SDID:  %1.2f\n", tau_sdid_w))
cat(sprintf("Equally-weighted DiD:      %1.2f\n", tau_did))
cat(sprintf("Population-weighted DiD:   %1.2f\n", tau_did_w))
```

# Standard errors

Variance estimators are adapted for weight renormalization during resampling.
Bootstrap, jackknife, and placebo methods are available.

```{r standard-errors}
se_jk = sqrt(vcov(tau_sdid_w, method = "jackknife"))
cat(sprintf("Weighted SDID: %1.2f (SE = %1.2f)\n", tau_sdid_w, se_jk))
cat(sprintf("95%% CI: (%1.2f, %1.2f)\n",
    tau_sdid_w - 1.96 * se_jk, tau_sdid_w + 1.96 * se_jk))
```

# Cluster-robust standard errors

When treatment is assigned at a higher level (e.g., states) than the unit of observation
(e.g., counties), pass a cluster variable. The bootstrap then resamples clusters.

```{r cluster-se}
# Assign units to 10 clusters
cluster = rep(1:10, length.out = nrow(setup$Y))

tau_cl = synthdid_estimate_weighted(setup$Y, N0, T0,
           treated.weights = pop_weights,
           cluster = cluster)

se_cl = suppressWarnings(sqrt(vcov(tau_cl, method = "bootstrap", replications = 50)))
cat(sprintf("Cluster-bootstrapped SE: %1.2f\n", se_cl))
```

# Event study

Decompose the treatment effect into period-specific estimates. Pre-treatment
coefficients serve as a visual parallel-trends check — they should be near zero
if the synthetic control is well-matched to the treated unit before treatment.

```{r event-study}
es = synthdid_event_study(tau_sdid_w)
cat(sprintf("Post-treatment effects: %s\n",
    paste(round(es$estimate[es$relative_time >= 0], 2), collapse = ", ")))
```

The vertical dashed line separates pre-treatment periods (left) from post-treatment
periods (right). Pre-treatment estimates near zero indicate good pre-treatment fit.
Post-treatment estimates show how the effect evolves over time.

# Summary

```{r summary}
s = summary(tau_sdid_w, fast = TRUE)
cat(sprintf("Estimate: %1.3f (SE: %1.3f)\n", s$estimate, s$se))
cat(sprintf("Effective N0: %1.1f, Effective N1: %1.1f\n",
    s$dimensions["N0.effective"], s$dimensions["N1.effective"]))
```
