---
title: "Weighted SDID"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weighted SDID}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(synthdid)
set.seed(42)
```

# Weighted estimation

The standard SDID estimator gives equal weight to all treated units.
When treated units differ in size, you may want a population-weighted estimate instead.
The `synthdid_estimate_weighted` function allows user-specified weights on treated units.

We demonstrate with the California Prop 99 dataset. In its original form
California is the *only* treated unit, so any set of treated-unit weights
collapses to a single weight of one — weighted and unweighted estimates are
then identical. To make weighting bite, we randomly promote four additional
states to "treated", sharing California's 1989 onset. This yields five treated
units that differ in size, so a population-weighted average differs from a
simple average.

```{r simulate}
# Use the California Prop 99 dataset included in the package
data('california_prop99')
ca = california_prop99

# California alone is treated here; with one treated unit the weighted and
# unweighted estimators coincide. Randomly add four more "treated" states
# (same 1989 onset) so that treated-unit weighting has an effect.
set.seed(42)
control_states = setdiff(as.character(unique(ca$State)), "California")
extra_treated  = sample(control_states, 4)
treated_states = c("California", extra_treated)
ca$treated     = as.integer(ca$State %in% treated_states & ca$Year >= 1989)

setup = panel.matrices(ca)
N0 = setup$N0
N1 = nrow(setup$Y) - N0
T0 = setup$T0

# Population weights (e.g., some states are larger).
# Here we just illustrate with fabricated weights, aligned to the treated rows.
treated_names = rownames(setup$Y)[(N0 + 1):nrow(setup$Y)]
pop           = setNames(runif(N1, 1, 10), treated_names)
pop_weights   = pop / sum(pop)

cat(sprintf("Treated units (N1 = %d): %s\n", N1,
    paste(treated_names, collapse = ", ")))
```

# Compare equally-weighted and population-weighted estimates

```{r estimates}
tau_sdid = synthdid_estimate(setup$Y, N0, T0)
tau_sdid_w = synthdid_estimate_weighted(setup$Y, N0, T0,
               treated.weights = pop_weights)
tau_did = did_estimate(setup$Y, N0, T0)
tau_did_w = did_estimate_weighted(setup$Y, N0, T0,
              treated.weights = pop_weights)

cat(sprintf("Equally-weighted SDID:     %1.2f\n", tau_sdid))
cat(sprintf("Population-weighted SDID:  %1.2f\n", tau_sdid_w))
cat(sprintf("Equally-weighted DiD:      %1.2f\n", tau_did))
cat(sprintf("Population-weighted DiD:   %1.2f\n", tau_did_w))
```

# Standard errors

Variance estimators are adapted for weight renormalization during resampling.
Bootstrap, jackknife, and placebo methods are available.

```{r standard-errors}
se_jk = sqrt(vcov(tau_sdid_w, method = "jackknife"))
cat(sprintf("Weighted SDID: %1.2f (SE = %1.2f)\n", tau_sdid_w, se_jk))
cat(sprintf("95%% CI: (%1.2f, %1.2f)\n",
    tau_sdid_w - 1.96 * se_jk, tau_sdid_w + 1.96 * se_jk))
```

# Cluster-robust standard errors

When treatment is assigned at a higher level (e.g., states) than the unit of observation
(e.g., counties), pass a cluster variable. The bootstrap then resamples clusters.

```{r cluster-se}
# Assign units to 10 clusters
cluster = rep(1:10, length.out = nrow(setup$Y))

tau_cl = synthdid_estimate_weighted(setup$Y, N0, T0,
           treated.weights = pop_weights,
           cluster = cluster)

se_cl = suppressWarnings(sqrt(vcov(tau_cl, method = "bootstrap", replications = 50)))
cat(sprintf("Cluster-bootstrapped SE: %1.2f\n", se_cl))
```

# Event study

Decompose the treatment effect into period-specific estimates. Pre-treatment
coefficients serve as a visual parallel-trends check — they should be near zero
if the synthetic control is well-matched to the treated unit before treatment.

```{r event-study}
es = synthdid_event_study(tau_sdid_w)
cat(sprintf("Post-treatment effects: %s\n",
    paste(round(es$estimate[es$relative_time >= 0], 2), collapse = ", ")))
```

Passing `se.method` adds period-by-period confidence intervals, and calling
`plot()` on the resulting object draws the event-study figure. We compare the
equally-weighted and population-weighted event studies; differences between the
two panels are driven entirely by the treated-unit weights.

```{r event-study-plot, fig.width = 7, fig.height = 4.5, fig.alt = "Equally-weighted SDID event study"}
es_uw = synthdid_event_study(tau_sdid,   se.method = "placebo")
es_w  = synthdid_event_study(tau_sdid_w, se.method = "placebo")

plot(es_uw, main = "Equally-weighted SDID event study")
```

```{r event-study-plot-weighted, fig.width = 7, fig.height = 4.5, fig.alt = "Population-weighted SDID event study"}
plot(es_w, main = "Population-weighted SDID event study")
```

The vertical dotted line separates pre-treatment periods (left) from post-treatment
periods (right), and the dashed horizontal line marks zero. Pre-treatment estimates
near zero indicate good pre-treatment fit; post-treatment estimates show how the
effect evolves over time. The blue whiskers are 95% placebo confidence intervals.

# Summary

```{r summary}
s = summary(tau_sdid_w, fast = TRUE)
cat(sprintf("Estimate: %1.3f (SE: %1.3f)\n", s$estimate, s$se))
cat(sprintf("Effective N0: %1.1f, Effective N1: %1.1f\n",
    s$dimensions["N0.effective"], s$dimensions["N1.effective"]))
```
