---
title: "Weighted Synthetic Difference-in-Differences"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weighted Synthetic Difference-in-Differences}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
ragg_png = function(..., res = 192) {
  ragg::agg_png(..., res = res, units = "in")
}
knitr::opts_chunk$set(dev = "ragg_png", fig.ext = "png")
```

```{r setup}
library(synthdid)
library(ggplot2)
set.seed(42)

# Source the weighted functions (in correct order for dependencies)
# Note: The base synthdid package provides helper functions like sc.weight.fw, contract3
source("../R_weighted/utils.R")
source("../R_weighted/solver.R")
source("../R_weighted/synthdid.R")
source("../R_weighted/vcov.R")
source("../R_weighted/summary.R")
source("../R_weighted/print.R")
```

# Introduction

This vignette demonstrates the weighted synthetic difference-in-differences estimator, which extends the standard SDID to allow for user-specified weights on treated units. This is particularly useful when:

1. There are multiple treated units of different sizes (e.g., states with different populations)
2. You want to estimate a population-weighted average treatment effect rather than a simple average
3. Some treated units are more policy-relevant than others

# Simulating Panel Data with Heterogeneous Treatment Effects

We simulate a panel dataset with:
- 40 control units
- 3 treated units with different populations and treatment effects
- 20 pre-treatment periods and 5 post-treatment periods

```{r simulate-data}
# Parameters
N0 <- 40        # Control units
N1 <- 3         # Treated units
T0 <- 20        # Pre-treatment periods
T1 <- 5         # Post-treatment periods
N <- N0 + N1
T_total <- T0 + T1

# Population weights for treated units (e.g., reflecting state populations)
# Unit 1: Large state (60% of treated population)
# Unit 2: Medium state (30% of treated population)
# Unit 3: Small state (10% of treated population)
pop_weights <- c(0.60, 0.30, 0.10)

# Heterogeneous treatment effects by unit
# Large state: small effect (tau = 5)
# Medium state: medium effect (tau = 15)
# Small state: large effect (tau = 30)
tau_by_unit <- c(5, 15, 30)

# True population-weighted average treatment effect
tau_true_weighted <- sum(pop_weights * tau_by_unit)
# True simple average treatment effect
tau_true_simple <- mean(tau_by_unit)

cat("True treatment effects by unit:\n")
cat(sprintf("  Unit 1 (pop weight %.0f%%): tau = %.1f\n", pop_weights[1]*100, tau_by_unit[1]))
cat(sprintf("  Unit 2 (pop weight %.0f%%): tau = %.1f\n", pop_weights[2]*100, tau_by_unit[2]))
cat(sprintf("  Unit 3 (pop weight %.0f%%): tau = %.1f\n", pop_weights[3]*100, tau_by_unit[3]))
cat(sprintf("\nTrue simple average effect: %.1f\n", tau_true_simple))
cat(sprintf("True population-weighted effect: %.1f\n", tau_true_weighted))
```

```{r generate-Y}
# Generate outcome matrix Y
# Y_it = alpha_i + beta_t + epsilon_it + tau_i * W_it

# Unit fixed effects
alpha <- c(rnorm(N0, mean = 50, sd = 10),
           rnorm(N1, mean = 55, sd = 5))

# Time fixed effects (with a trend)
beta <- cumsum(rnorm(T_total, mean = 0.5, sd = 1))

# Generate Y matrix
Y <- matrix(0, nrow = N, ncol = T_total)
for (i in 1:N) {
  for (t in 1:T_total) {
    Y[i, t] <- alpha[i] + beta[t] + rnorm(1, sd = 2)
  }
}

# Add treatment effects for treated units in post-period
for (j in 1:N1) {
  i <- N0 + j
  for (t in (T0 + 1):T_total) {
    Y[i, t] <- Y[i, t] + tau_by_unit[j]
  }
}

# Add row and column names
rownames(Y) <- c(paste0("Control_", 1:N0), paste0("Treated_", 1:N1))
colnames(Y) <- paste0("t", 1:T_total)

cat("Dimensions of Y:", nrow(Y), "units x", ncol(Y), "periods\n")
cat("N0 =", N0, "control units, N1 =", N1, "treated units\n")
cat("T0 =", T0, "pre-treatment periods, T1 =", T1, "post-treatment periods\n")
```

# Standard (Unweighted) Estimation

First, we estimate the treatment effect using the standard SDID estimator, which gives equal weight to all treated units.

```{r standard-estimate}
# Standard SDID (equal weights on treated units)
tau_sdid <- synthdid_estimate(Y, N0, T0)
tau_sc <- sc_estimate(Y, N0, T0)
tau_did <- did_estimate(Y, N0, T0)

cat("Standard Estimates (equal weights on treated units):\n")
cat(sprintf("  Diff-in-Diff:              %.2f\n", tau_did))
cat(sprintf("  Synthetic Control:         %.2f\n", tau_sc))
cat(sprintf("  Synthetic Diff-in-Diff:    %.2f\n", tau_sdid))
cat(sprintf("\nTrue simple average effect:  %.2f\n", tau_true_simple))
```

# Weighted Estimation

Now we use the weighted estimator with population weights. This gives more weight to larger treated units.

```{r weighted-estimate}
# Weighted SDID (population weights on treated units)
tau_sdid_w <- synthdid_estimate_weighted(Y, N0, T0, treated.weights = pop_weights)
tau_sc_w <- sc_estimate_weighted(Y, N0, T0, treated.weights = pop_weights)
tau_did_w <- did_estimate_weighted(Y, N0, T0, treated.weights = pop_weights)

cat("Weighted Estimates (population weights on treated units):\n")
cat(sprintf("  Diff-in-Diff:              %.2f\n", tau_did_w))
cat(sprintf("  Synthetic Control:         %.2f\n", tau_sc_w))
cat(sprintf("  Synthetic Diff-in-Diff:    %.2f\n", tau_sdid_w))
cat(sprintf("\nTrue population-weighted effect: %.2f\n", tau_true_weighted))
```

# Comparison of Results

```{r comparison-table}
results <- data.frame(
  Estimator = c("Diff-in-Diff", "Synthetic Control", "Synthetic Diff-in-Diff"),
  Unweighted = c(as.numeric(tau_did), as.numeric(tau_sc), as.numeric(tau_sdid)),
  Weighted = c(as.numeric(tau_did_w), as.numeric(tau_sc_w), as.numeric(tau_sdid_w)),
  True_Simple = rep(tau_true_simple, 3),
  True_Weighted = rep(tau_true_weighted, 3)
)

results$Unweighted_Error <- results$Unweighted - results$True_Simple
results$Weighted_Error <- results$Weighted - results$True_Weighted

knitr::kable(results, digits = 2,
             caption = "Comparison of Weighted vs Unweighted Estimates")
```

# Understanding the Difference

The key insight is that with heterogeneous treatment effects across treated units, the choice of weights matters:

- **Unweighted (simple average)**: Each treated unit contributes equally to the average treatment effect. This answers: "What is the average effect across treated units?"

- **Weighted (population-weighted)**: Larger units contribute more to the average. This answers: "What is the average effect per person in the treated population?"

In our simulation:
- The largest treated unit (60% of population) has the smallest effect (tau = 5)
- The smallest treated unit (10% of population) has the largest effect (tau = 30)

Therefore, the population-weighted effect (`r round(tau_true_weighted, 1)`) is smaller than the simple average (`r round(tau_true_simple, 1)`).

# Visualizing the Weights

```{r weight-visualization, fig.width=8, fig.height=4}
# Compare the treated unit weights
weight_df <- data.frame(
  Unit = rep(paste0("Treated_", 1:N1), 2),
  Weight = c(rep(1/N1, N1), pop_weights),
  Type = rep(c("Unweighted (equal)", "Population-weighted"), each = N1),
  Effect = rep(tau_by_unit, 2)
)

ggplot(weight_df, aes(x = Unit, y = Weight, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  geom_text(aes(label = sprintf("tau=%.0f", Effect)),
            position = position_dodge(width = 0.9), vjust = -0.5, size = 3) +
  scale_fill_manual(values = c("steelblue", "coral")) +
  labs(title = "Treated Unit Weights: Equal vs Population-Weighted",
       subtitle = sprintf("Simple avg effect = %.1f, Population-weighted effect = %.1f",
                         tau_true_simple, tau_true_weighted),
       y = "Weight", x = "Treated Unit") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

# Examining the Synthetic Control Weights

```{r examine-weights}
# Get the omega weights from both estimates
omega_unweighted <- attr(tau_sdid, 'weights')$omega
omega_weighted <- attr(tau_sdid_w, 'weights')$omega

# Top control units by weight
top_n <- 10
top_unweighted <- head(sort(omega_unweighted, decreasing = TRUE), top_n)
top_weighted <- head(sort(omega_weighted, decreasing = TRUE), top_n)

cat("Top", top_n, "control unit weights (unweighted SDID):\n")
print(round(top_unweighted, 4))

cat("\nTop", top_n, "control unit weights (weighted SDID):\n")
print(round(top_weighted, 4))
```

# Standard Errors

We can compute standard errors for the weighted estimates using bootstrap, jackknife, or placebo methods.

```{r standard-errors}
# Compute standard errors (using jackknife for speed)
se_sdid <- sqrt(vcov(tau_sdid, method = "jackknife"))
se_sdid_w <- sqrt(vcov(tau_sdid_w, method = "jackknife"))

cat("Standard Errors (jackknife):\n")
cat(sprintf("  Unweighted SDID: %.2f (SE = %.2f)\n", tau_sdid, se_sdid))
cat(sprintf("  Weighted SDID:   %.2f (SE = %.2f)\n", tau_sdid_w, se_sdid_w))

# 95% confidence intervals
cat("\n95% Confidence Intervals:\n")
cat(sprintf("  Unweighted: (%.2f, %.2f)\n",
            tau_sdid - 1.96*se_sdid, tau_sdid + 1.96*se_sdid))
cat(sprintf("  Weighted:   (%.2f, %.2f)\n",
            tau_sdid_w - 1.96*se_sdid_w, tau_sdid_w + 1.96*se_sdid_w))
```

# Summary

```{r summary}
cat("Summary of Weighted SDID Estimate:\n")

# Extract components manually to avoid S4 dispatch issues
setup = attr(tau_sdid_w, 'setup')
weights = attr(tau_sdid_w, 'weights')
tw = attr(tau_sdid_w, 'treated.weights')
pw = attr(tau_sdid_w, 'period.weights')

cat(sprintf("  Estimate: %.3f\n", as.numeric(tau_sdid_w)))
cat(sprintf("  N0 = %d, N1 = %d, T0 = %d, T1 = %d\n",
            setup$N0, nrow(setup$Y) - setup$N0, setup$T0, ncol(setup$Y) - setup$T0))
cat(sprintf("  Effective N0: %.1f\n", 1/sum(weights$omega^2)))
cat(sprintf("  Effective N1: %.1f\n", 1/sum(tw^2)))
cat("\nTreated unit weights:\n")
print(round(tw, 3))
cat("\nTop 5 control unit weights (omega):\n")
print(round(head(sort(weights$omega, decreasing = TRUE), 5), 3))
```

# When to Use Weighted vs Unweighted

| Scenario | Recommended Approach |
|----------|---------------------|
| Policy evaluation for population | Use population weights |
| Effect on "typical" treated unit | Use equal weights |
| Treatment effects vary by unit size | Consider both; report as robustness |
| Units represent different subgroups | Weight by subgroup importance |

# Conclusion

The weighted SDID estimator provides flexibility when treatment effects are heterogeneous across treated units. By allowing user-specified weights, researchers can:

1. Estimate population-weighted average treatment effects
2. Up-weight policy-relevant units
3. Account for differential sampling probabilities

The choice between weighted and unweighted estimates should be guided by the research question: Are you interested in the average effect per treated unit, or the average effect per person/entity in the treated population?
