---
title: "Supplementary Appendix: Weighted Synthetic Difference-in-Differences"
author: "John Iselin\\thanks{The Budget Lab at Yale} \\and Erica Ryan\\thanks{Amazon}"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: true
    keep_tex: true
    fig_caption: true
    citation_package: natbib
    includes:
      in_header: header.tex
bibliography: references.bib
biblio-style: elsarticle-harv
fontsize: 12pt
geometry: margin=1in
linestretch: 1.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4,
  fig.align = "center",
  out.width = "90%",
  cache = TRUE
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(synthdid)

# Source weighted SDID functions
source("../R/utils.R")
source("../R/solver.R")
source("../R/synthdid.R")
source("../R/vcov.R")
source("../R/summary.R")
source("../R/print.R")

# Reproducibility
set.seed(20240101)
```

\renewcommand{\thesection}{\Alph{section}}
\setcounter{section}{0}

This supplement accompanies "Weighted Synthetic Difference-in-Differences" and contains technical details, proofs, and additional results referenced in the main text.

# Setup and the Standard SDID Estimator {#app:setup}

This appendix reviews the standard SDID estimator of @arkhangelsky2021 and establishes the notation used throughout the main text and subsequent appendices.

## Model and Notation

Consider a balanced panel with $N$ units observed over $T$ periods. The first $N_0$ units are untreated controls; the remaining $N_1 = N - N_0$ are treated beginning at period $T_0 + 1$. Let $T_1 = T - T_0$ denote the number of post-treatment periods. Define the treatment indicator
$$
W_{it} = \mathbf{1}\{i > N_0\} \cdot \mathbf{1}\{t > T_0\}.
$$

The observed outcome is
$$
Y_{it} = Y_{it}(0) + \tau_{it} \cdot W_{it},
$$
where $Y_{it}(0)$ is the potential outcome absent treatment and $\tau_{it}$ is the unit-time treatment effect. The object of interest is an average treatment effect on the treated (ATT), though the precise weighted average depends on how treated units are aggregated.

## The SDID Estimator

Following @arkhangelsky2021, the SDID estimator solves
\begin{equation} \label{eq:sdid-regression}
(\hat{\tau}^{\text{sdid}}, \hat{\mu}, \hat{\alpha}, \hat{\beta}) = \underset{\tau, \mu, \alpha_i, \beta_t}{\arg\min} \sum_{i=1}^{N} \sum_{t=1}^{T} \left( Y_{it} - \mu - \alpha_i - \beta_t - \tau \cdot W_{it} \right)^2 \hat{\omega}_i \hat{\lambda}_t,
\end{equation}
where $\hat{\omega}_i$ are unit weights and $\hat{\lambda}_t$ are time weights.

This admits a closed-form solution. Define the *collapsed form* $\tilde{Y}$, an $(N_0 + 1) \times (T_0 + 1)$ matrix in which the last row averages treated units and the last column averages post-treatment periods:
\begin{equation} \label{eq:collapsed}
\tilde{Y}_{N_0+1, t} = \frac{1}{N_1} \sum_{i=N_0+1}^{N} Y_{it}, \quad
\tilde{Y}_{i, T_0+1} = \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it}.
\end{equation}

The treatment effect estimate is then
\begin{equation} \label{eq:tau-sdid}
\hat{\tau}^{\text{sdid}} = \begin{pmatrix} -\hat{\omega} \\ \frac{1}{N_1}\mathbf{1}_{N_1} \end{pmatrix}^\top Y \begin{pmatrix} -\hat{\lambda} \\ \frac{1}{T_1}\mathbf{1}_{T_1} \end{pmatrix},
\end{equation}
a double-difference comparing the weighted treated-vs-control gap in the post-period to the same gap in the $\hat{\lambda}$-weighted pre-period.

## Weight Estimation

Control unit weights $\hat{\omega} \in \Delta^{N_0}$ (the unit simplex) are estimated by matching the weighted control average to the treated average in pre-treatment periods:
\begin{equation} \label{eq:omega}
(\hat{\omega}_0, \hat{\omega}) = \underset{\omega_0 \in \mathbb{R},\; \omega \in \Delta^{N_0}}{\arg\min} \sum_{t=1}^{T_0} \left( \omega_0 + \sum_{i=1}^{N_0} \omega_i Y_{it} - \frac{1}{N_1} \sum_{i=N_0+1}^{N} Y_{it} \right)^2 + \zeta_\omega^2 T_0 \|\omega\|_2^2.
\end{equation}

Time weights $\hat{\lambda} \in \Delta^{T_0}$ are estimated analogously, matching the $\hat{\lambda}$-weighted pre-treatment average of control units to their post-treatment average:
\begin{equation} \label{eq:lambda}
(\hat{\lambda}_0, \hat{\lambda}) = \underset{\lambda_0 \in \mathbb{R},\; \lambda \in \Delta^{T_0}}{\arg\min} \sum_{i=1}^{N_0} \left( \lambda_0 + \sum_{t=1}^{T_0} \lambda_t Y_{it} - \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it} \right)^2 + \zeta_\lambda^2 N_0 \|\lambda\|_2^2.
\end{equation}

Both optimization problems are solved via Frank-Wolfe on the simplex with exact line search, with regularization parameters $\zeta_\omega = \eta_\omega \cdot \hat{\sigma}$ and $\zeta_\lambda = \eta_\lambda \cdot \hat{\sigma}$, where $\hat{\sigma}$ is the noise level estimated from first differences of the control block.

**The key observation** is that the target on the right-hand side of \eqref{eq:omega}---the average treated outcome---uses equal weights $1/N_1$. This determines which synthetic control group is constructed. If the researcher's estimand places different weights on treated units, the control weights $\hat{\omega}$ should be re-estimated against the appropriately weighted treated average.


# Inference for the Weighted Estimator {#app:inference}

As described in the main text, @arkhangelsky2021 propose three approaches to variance estimation: bootstrap (Algorithm 2), jackknife (Algorithm 3), and placebo (Algorithm 4). Each involves resampling units in some fashion. With user-specified treated-unit weights, a complication arises: when a resampling scheme drops or duplicates treated units, the weights $\tilde{\omega}$ must be renormalized to maintain their interpretation.

## Bootstrap

The bootstrap resamples units with replacement. In a given bootstrap draw $b$, let $n_j^{(b)}$ denote the number of times treated unit $j$ appears. The bootstrap-adjusted weights are
\begin{equation} \label{eq:boot-weights}
\tilde{\omega}_j^{(b)} = \frac{\tilde{\omega}_j \cdot n_j^{(b)}}{\sum_{k=1}^{N_1} \tilde{\omega}_k \cdot n_k^{(b)}},
\end{equation}
with the convention that if no treated units are drawn (probability exponentially small in $N_1$), the draw is discarded. Control weights $\hat{\omega}$ are similarly renormalized. The bootstrap variance is
$$
\widehat{\text{Var}}^{\text{boot}} = \frac{B-1}{B} \cdot \frac{1}{B} \sum_{b=1}^{B} \left( \hat{\tau}^{w,(b)} - \bar{\tau}^w \right)^2.
$$

## Jackknife

The leave-one-out jackknife drops unit $i$ in turn. When unit $i$ is a treated unit $j$, the remaining treated weights are renormalized:
\begin{equation} \label{eq:jk-weights}
\tilde{\omega}_k^{(-j)} = \frac{\tilde{\omega}_k}{\sum_{\ell \neq j} \tilde{\omega}_\ell}, \quad k \neq j.
\end{equation}

As in the standard case, we use the *fixed-weight* jackknife, holding $\hat{\omega}^w$ and $\hat{\lambda}^w$ at their full-sample values while renormalizing the control weight vector when a control unit is dropped. The jackknife variance is
$$
\widehat{\text{Var}}^{\text{jk}} = \frac{N-1}{N} \sum_{i=1}^{N} \left( \hat{\tau}^{w,(-i)} - \bar{\tau}^{w,(\cdot)} \right)^2.
$$

Following @arkhangelsky2021, the jackknife is not recommended for the SC variant due to the non-differentiability of the SC weights at the boundary of the simplex.

## Placebo

The placebo method randomly assigns $N_1$ control units as "placebo treated" and estimates the treatment effect on them, repeating $B$ times. Since the original treated-unit weights $\tilde{\omega}$ do not naturally map to control units serving as placebo treated, we use uniform weights $1/N_1$ for the placebo-treated group by default. This approach tests the sharp null of no treatment effect and requires $N_0 > N_1$.\footnote{The placebo variance estimator tests the sharp null under uniform weighting of pseudo-treated units, which may not fully capture the variance contribution of the weight structure when treatment effects are heterogeneous and correlated with unit size. Our software implementation provides two alternative placebo weighting schemes: \texttt{placebo.weights = "size\_match"} weights pseudo-treated units by their pre-treatment outcome levels (as a proxy for unit size), and \texttt{placebo.weights = "permute"} randomly permutes the original $\tilde{\omega}$ vector. The latter preserves the weight concentration (HHI) of the original specification and may provide a more conservative variance estimate when weights are highly skewed.}


# Proof of Proposition 1 {#app:proof}

As stated in the main text:

\medskip
\noindent\textbf{Proposition 1.} *When $\tilde{\omega}_j = 1/N_1$ for all $j$, the weighted estimator $\hat{\tau}^w$ reduces to the standard SDID estimator $\hat{\tau}^{\text{sdid}}$.*

\medskip
\noindent\textit{Proof.} When $\tilde{\omega}_j = 1/N_1$ for all $j$, the weighted collapsed form (equation 3 of the main text) reduces to
$$
\tilde{Y}^w_{N_0+1, t} = \frac{1}{N_1} \sum_{j=1}^{N_1} Y_{N_0+j,t} = \tilde{Y}_{N_0+1, t}
$$
for all $t = 1, \ldots, T_0$, and analogously for the last column. Since the collapsed forms are identical, the Frank-Wolfe solutions $(\hat{\omega}^w, \hat{\lambda}^w)$ equal $(\hat{\omega}, \hat{\lambda})$. The estimate (equation 4 of the main text) then equals the standard SDID estimate (equation 2). $\square$


# Event Study Decomposition and Additional Figures {#app:event-study}

This appendix provides event-study decompositions and diagnostics for the control unit weights that supplement the main results reported in the text.

## Event Study Decomposition

Following the decomposition of @ciccia2024, we disaggregate each estimator's ATT into lag-specific treatment effects $\hat{\tau}_l$ for each post-treatment period $l = 0, \ldots, T_1 - 1$, and compute pre-treatment placebo coefficients for $k = -(T_0 - 1), \ldots, -1$. At each period $t$, the event-study coefficient is the gap between the (weighted) treated average and the synthetic control at $t$, minus the $\hat{\lambda}$-weighted pre-treatment intercept:
$$
\hat{\tau}_t = \left[\sum_j \tilde{\omega}_j Y_{jt} - \sum_i \hat{\omega}_i Y_{it}\right] - \sum_{s=1}^{T_0} \hat{\lambda}_s \left[\sum_j \tilde{\omega}_j Y_{js} - \sum_i \hat{\omega}_i Y_{is}\right].
$$
Pre-treatment coefficients near zero provide a visual parallel-trends diagnostic. The overall ATT equals the average of post-treatment coefficients: $\hat{\tau} = T_1^{-1} \sum_{l=0}^{T_1-1} \hat{\tau}_l$.

## All-Estimator Event Study Comparison

```{r load-data}
# Load county-year panel
panel <- readr::read_csv(
  file.path("..", "data", "analysis_data.csv"),
  show_col_types = FALSE
) %>%
  select(
    unit = fips,
    time = year,
    y = crude_rate,
    treated_unit = expansion,
    pop = population,
    state_fips,
    pct_white, pop_20_64, pct_55_64,
    log_20_64, log_35_44, log_f_20_64, unemp
  ) %>%
  mutate(post = as.integer(time >= 2014))

# Panel dimensions
n_counties <- n_distinct(panel$unit)
n_treated <- n_distinct(panel$unit[panel$treated_unit == 1])
n_control <- n_distinct(panel$unit[panel$treated_unit == 0])
year_range <- range(panel$time)
n_years <- n_distinct(panel$time)
T0 <- sum(sort(unique(panel$time)) < 2014)
T1 <- n_years - T0
```

```{r prepare-sdid-panel}
# Prepare the balanced panel in matrix form for synthdid
panel_sdid <- panel %>%
  arrange(unit, time) %>%
  select(unit, time, y, treated_unit, pop) %>%
  mutate(
    .unit = as.factor(unit),
    .time = time,
    .W = as.integer(treated_unit == 1 & time >= 2014)
  )

dfp <- panel_sdid %>%
  select(.unit, .time, y, .W) %>%
  as.data.frame()

setup <- synthdid::panel.matrices(dfp)

# Compute population weights for treated counties
pop_2013 <- panel %>%
  filter(treated_unit == 1, time == 2013) %>%
  select(unit, pop)

# Get treated unit order from setup$Y
treated_names <- rownames(setup$Y)[(setup$N0 + 1):nrow(setup$Y)]
pop_ordered <- pop_2013 %>%
  mutate(unit_char = as.character(unit))
weight_order <- match(treated_names, pop_ordered$unit_char)
treated_weights <- pop_ordered$pop[weight_order]
treated_weights <- treated_weights / sum(treated_weights, na.rm = TRUE)

# Alternative weight specifications for robustness
pop_2010 <- panel %>% filter(treated_unit == 1, time == 2010) %>% select(unit, pop)
pop_ordered_2010 <- pop_2010 %>% mutate(unit_char = as.character(unit))
weight_order_2010 <- match(treated_names, pop_ordered_2010$unit_char)

treated_weights_log  <- log(pop_ordered$pop[weight_order])
treated_weights_log  <- treated_weights_log / sum(treated_weights_log, na.rm = TRUE)

treated_weights_sqrt <- sqrt(pop_ordered$pop[weight_order])
treated_weights_sqrt <- treated_weights_sqrt / sum(treated_weights_sqrt, na.rm = TRUE)

treated_weights_2010 <- pop_ordered_2010$pop[weight_order_2010]
treated_weights_2010 <- treated_weights_2010 / sum(treated_weights_2010, na.rm = TRUE)

```

```{r run-estimates-supp, cache=TRUE}
# Unweighted estimates
tau_sdid   <- synthdid_estimate(setup$Y, setup$N0, setup$T0)
tau_sc     <- sc_estimate(setup$Y, setup$N0, setup$T0)
tau_did    <- did_estimate(setup$Y, setup$N0, setup$T0)

# Population-weighted estimates
tau_sdid_w <- synthdid_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                          treated.weights = treated_weights)
tau_sc_w   <- sc_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                    treated.weights = treated_weights)
tau_did_w  <- did_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                     treated.weights = treated_weights)
```

```{r event-study-all, fig.cap="Event-study decomposition comparing DID, SC, and SDID estimators (equally weighted). Shaded bands show 95\\% bootstrap confidence intervals.", fig.height=4.5, cache=TRUE}
# Compute event-study coefficients with bootstrap SEs
es_sdid <- synthdid_event_study(tau_sdid, se.method = "bootstrap", replications = 200)
es_sc   <- synthdid_event_study(tau_sc,   se.method = "bootstrap", replications = 200)
es_did  <- synthdid_event_study(tau_did,  se.method = "bootstrap", replications = 200)

es_all <- bind_rows(
  es_did  %>% mutate(Estimator = "DID"),
  es_sc   %>% mutate(Estimator = "SC"),
  es_sdid %>% mutate(Estimator = "SDID")
) %>%
  mutate(
    year = as.numeric(time),
    Estimator = factor(Estimator, levels = c("DID", "SC", "SDID"))
  )

ggplot(es_all, aes(x = year, y = estimate, color = Estimator, fill = Estimator)) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.12, color = NA) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 2013.5, linetype = "dashed", color = "gray30") +
  geom_point(size = 1.5) +
  geom_line(linewidth = 0.6) +
  scale_color_manual(values = c("DID" = "gray40", "SC" = "darkorange", "SDID" = "steelblue")) +
  scale_fill_manual(values = c("DID" = "gray40", "SC" = "darkorange", "SDID" = "steelblue")) +
  labs(
    x = NULL, y = "Effect on mortality\n(deaths per 100,000)",
    color = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

Figure D.1 compares the event-study paths across all three estimators (equally weighted). As described in the main text, the pre-treatment fit differs across methods: SC achieves the tightest pre-treatment balance by construction (all pre-treatment coefficients are near zero), while DID, which uses uniform time weights, may show pre-treatment fluctuations that average to zero. SDID falls between these extremes, using data-driven time weights to improve pre-treatment fit relative to DID. Post-treatment, the three estimators tell broadly similar stories under equal weighting, with all estimates fluctuating near zero. The DID and SDID confidence intervals overlap substantially, while the SC intervals are wider, reflecting the greater estimation uncertainty from relying entirely on a small number of control weights. The similarity of post-treatment paths across estimators under equal weighting---contrasted with the dramatic divergence when population weights are introduced (see Figure 1 in the main text)---underscores that the weighting choice, not the estimator choice, drives the main results in this application.

## Control Unit Weights

```{r omega-weights, fig.cap="Distribution of SDID control unit weights ($\\hat{\\omega}$) under equal vs.\\ population weighting. Counties with zero weight are excluded. The x-axis is on a log scale.", fig.height=3.5}
omega_unwt <- attr(tau_sdid, 'weights')$omega
omega_wt   <- attr(tau_sdid_w, 'weights')$omega

# Combine nonzero weights
omega_df <- bind_rows(
  tibble(omega = omega_unwt[omega_unwt > 0], Weighting = "Equally weighted"),
  tibble(omega = omega_wt[omega_wt > 0],     Weighting = "Population weighted")
)

ggplot(omega_df, aes(x = omega, fill = Weighting)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 40) +
  scale_x_log10() +
  scale_fill_manual(values = c("Equally weighted" = "steelblue", "Population weighted" = "firebrick")) +
  labs(
    x = expression(hat(omega)[i] ~ "(log scale)"),
    y = "Number of control counties",
    fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

```{r effective-units}
eff_N0_unwt <- 1 / sum(omega_unwt^2)
eff_N0_wt   <- 1 / sum(omega_wt^2)
n_nonzero_unwt <- sum(omega_unwt > 1e-6)
n_nonzero_wt   <- sum(omega_wt > 1e-6)
```

Figure D.2 shows the distribution of control unit weights under both specifications. The unweighted SDID places nonzero weight on `r n_nonzero_unwt` control counties (effective $N_0$ = `r round(eff_N0_unwt, 1)`), while the population-weighted SDID uses `r n_nonzero_wt` counties (effective $N_0$ = `r round(eff_N0_wt, 1)`). Both specifications spread weight broadly across the control pool, with similar effective sample sizes, reflecting the regularization built into the SDID algorithm. However, the *identity* of which control counties receive the most weight shifts: under population weighting, the Frank-Wolfe algorithm must match the pre-treatment trajectory of the population-weighted treated average, which is dominated by large urban counties. This pulls the synthetic control toward control counties with similar urban characteristics, subtly altering the comparison group even though the number of contributing counties is similar.


# Monte Carlo Simulations {#app:simulations}

We examine the finite-sample performance of the weighted SDID estimator and its variance estimators via simulation. The key question is whether the adapted bootstrap, jackknife, and placebo variance estimators achieve nominal coverage when treatment effects are heterogeneous and correlated with unit size.

## Design

The data-generating process follows the factor model of @arkhangelsky2021. For $N = N_0 + N_1$ units observed over $T = T_0 + T_1$ periods:
$$
Y_{it} = L_{it} + \tau_{it} W_{it} + \sigma \epsilon_{it},
$$
where $L = UV^\top + \alpha \mathbf{1}^\top + \mathbf{1}\beta^\top$ is a rank-2 factor structure with additive fixed effects, $U \in \mathbb{R}^{N \times 2}$ and $V \in \mathbb{R}^{T \times 2}$ are factor loadings drawn from Poisson distributions, $\alpha_i = 10 \cdot i/N$, $\beta_t = 10 \cdot t/T$, and $\epsilon_{it}$ is drawn from an AR(1) process with $\rho = 0.7$.

**Treatment effect heterogeneity.** Each treated unit $j$ is assigned a "size" $s_j$ drawn from a log-normal distribution, scaled to achieve a target Herfindahl-Hirschman Index (HHI) for the population weights $\tilde{\omega}_j = s_j / \sum_k s_k$. The treatment effect is
$$
\tau_{ij} = \bar{\tau} + \gamma \cdot \left(\frac{s_j}{\bar{s}} - 1\right),
$$
where $\bar{\tau} = 1$ is the baseline effect, $\bar{s} = N_1^{-1}\sum_j s_j$, and $\gamma \geq 0$ controls the correlation between unit size and treatment effect magnitude. When $\gamma = 0$, effects are homogeneous; when $\gamma > 0$, larger units have larger effects, creating a wedge between the equally-weighted ATT ($\bar{\tau}$) and the population-weighted ATT ($\sum_j \tilde{\omega}_j \tau_j > \bar{\tau}$).

**Configurations.** We fix $N_0 = 100$, $T_0 = 40$, $T_1 = 10$, $\sigma = 0.5$, and vary:

- $N_1 \in \{5, 10, 20\}$: number of treated units.
- $\gamma \in \{0, 0.5, 1.0\}$: size-effect correlation (none, moderate, strong).
- Weight concentration: "low" ($\text{HHI} = 1.5/N_1$, mildly concentrated) vs. "high" ($\text{HHI} = 3/N_1$, substantially concentrated). For $N_1 = 10$, this corresponds to HHI of 0.15 and 0.30, with effective $N_1^{\text{eff}}$ of approximately 6.7 and 3.3 respectively.

This yields 18 configurations. For each, we run 500 simulations and compute the weighted and unweighted SDID estimates with bootstrap ($B = 200$), jackknife, and placebo ($B = 200$) standard errors.

```{r mc-functions}
#' Generate sizes with a target HHI
#' @param N1 number of treated units
#' @param target_hhi target Herfindahl index for population weights
generate_sizes = function(N1, target_hhi) {
  if (target_hhi <= 1/N1 + 0.005) {
    # Near-uniform: just use uniform
    return(rep(1, N1))
  }
  # Use log-normal, tune sd to hit target HHI via binary search
  # Save and restore RNG state so calibration doesn't corrupt the outer seed
  saved_seed = .Random.seed
  obj = function(log_sd) {
    set.seed(999)
    s = rlnorm(N1 * 100, meanlog = 0, sdlog = exp(log_sd))
    hhis = sapply(split(s, rep(1:100, each = N1)), function(x) sum((x/sum(x))^2))
    mean(hhis) - target_hhi
  }
  opt = tryCatch(
    uniroot(obj, c(-3, 3), tol = 0.001),
    error = function(e) list(root = 0.5)
  )
  .Random.seed <<- saved_seed  # restore RNG
  s = rlnorm(N1, meanlog = 0, sdlog = exp(opt$root))
  s
}

#' Run one simulation replicate
#' @param N0,N1,T0,T1 panel dimensions
#' @param gamma size-effect correlation
#' @param target_hhi target weight concentration
#' @param sigma noise level
#' @param tau_bar baseline treatment effect
#' @param B number of bootstrap/placebo replications
run_one_sim = function(N0, N1, T0, T1, gamma, target_hhi, sigma = 0.5,
                       tau_bar = 1, B = 200) {
  N = N0 + N1
  T_ = T0 + T1
  rank = 2
  rho = 0.7

  # Factor model
  U = matrix(rpois(rank * N, sqrt(sample(1:N)) / sqrt(N)), N, rank)
  V = matrix(rpois(rank * T_, sqrt(1:T_) / sqrt(T_)), T_, rank)
  alpha = outer(10 * sample(1:N) / N, rep(1, T_))
  beta  = outer(rep(1, N), 10 * (1:T_) / T_)
  L = U %*% t(V) + alpha + beta

  # AR(1) errors
  var_mat = outer(1:T_, 1:T_, FUN = function(x, y) rho^(abs(x - y)))
  error = mvtnorm::rmvnorm(N, sigma = var_mat, method = "chol")

  # Treatment assignment
  W = matrix(0, N, T_)
  W[(N0+1):N, (T0+1):T_] = 1

  # Size-dependent treatment effects
  sizes = generate_sizes(N1, target_hhi)
  tw = sizes / sum(sizes)  # population weights
  s_bar = mean(sizes)
  tau_vec = tau_bar + gamma * (sizes / s_bar - 1)  # unit-specific effects

  # True ATTs
  true_att_equal    = mean(tau_vec)
  true_att_weighted = sum(tw * tau_vec)

  # Treatment effect matrix
  tau_mat = matrix(0, N, T_)
  for (j in 1:N1) {
    tau_mat[N0 + j, (T0+1):T_] = tau_vec[j]
  }

  Y = L + tau_mat + sigma * error
  rownames(Y) = 1:N
  colnames(Y) = 1:T_

  # Estimate
  est_unwt = tryCatch(synthdid_estimate(Y, N0, T0), error = function(e) NA)
  est_wt   = tryCatch(synthdid_estimate_weighted(Y, N0, T0, treated.weights = tw),
                       error = function(e) NA)

  if (is.na(est_unwt[1]) || is.na(est_wt[1])) {
    return(data.frame(
      est_unwt = NA, est_wt = NA,
      se_boot_unwt = NA, se_jk_unwt = NA, se_plac_unwt = NA,
      se_boot_wt = NA, se_jk_wt = NA, se_plac_wt = NA,
      true_att_equal = true_att_equal, true_att_weighted = true_att_weighted
    ))
  }

  # SEs -- unweighted
  se_boot_unwt = tryCatch(sqrt(vcov(est_unwt, method = "bootstrap", replications = B)),
                           error = function(e) NA)
  se_jk_unwt   = tryCatch(sqrt(vcov(est_unwt, method = "jackknife")),
                           error = function(e) NA)
  se_plac_unwt = tryCatch(sqrt(vcov(est_unwt, method = "placebo", replications = B)),
                           error = function(e) NA)

  # SEs -- weighted
  se_boot_wt = tryCatch(sqrt(vcov(est_wt, method = "bootstrap", replications = B)),
                         error = function(e) NA)
  se_jk_wt   = tryCatch(sqrt(vcov(est_wt, method = "jackknife")),
                         error = function(e) NA)
  se_plac_wt = tryCatch(sqrt(vcov(est_wt, method = "placebo", replications = B)),
                         error = function(e) NA)

  data.frame(
    est_unwt = c(est_unwt), est_wt = c(est_wt),
    se_boot_unwt = c(se_boot_unwt), se_jk_unwt = c(se_jk_unwt), se_plac_unwt = c(se_plac_unwt),
    se_boot_wt = c(se_boot_wt), se_jk_wt = c(se_jk_wt), se_plac_wt = c(se_plac_wt),
    true_att_equal = true_att_equal, true_att_weighted = true_att_weighted
  )
}
```

```{r mc-run, cache=TRUE}
# Define configurations
configs = expand.grid(
  N1 = c(5, 10, 20),
  gamma = c(0, 0.5, 1.0),
  hhi_mult = c(1.5, 3)  # HHI = hhi_mult / N1 (1.5 = mild, 3 = concentrated)
)
configs$hhi = configs$hhi_mult / configs$N1

N_SIM = 500
N0 = 100; T0 = 40; T1 = 10

mc_results = list()
for (cfg in 1:nrow(configs)) {
  n1 = configs$N1[cfg]
  gam = configs$gamma[cfg]
  hhi = configs$hhi[cfg]

  sims = do.call(rbind, lapply(1:N_SIM, function(s) {
    set.seed(20240101 + s)
    suppressWarnings(run_one_sim(N0, n1, T0, T1, gam, hhi))
  }))

  sims$config = cfg
  sims$N1 = n1
  sims$gamma = gam
  sims$hhi = hhi
  mc_results[[cfg]] = sims
}

mc_df = do.call(rbind, mc_results)
```

## Results

```{r mc-bias-table}
# Compute bias and RMSE for each configuration
bias_tbl = mc_df %>%
  group_by(N1, gamma, hhi) %>%
  summarise(
    Bias_unwt = mean(est_unwt - true_att_equal, na.rm = TRUE),
    RMSE_unwt = sqrt(mean((est_unwt - true_att_equal)^2, na.rm = TRUE)),
    Bias_wt   = mean(est_wt - true_att_weighted, na.rm = TRUE),
    RMSE_wt   = sqrt(mean((est_wt - true_att_weighted)^2, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric) & !c(N1, gamma, hhi), ~ round(., 3)))

kableExtra::kbl(
  bias_tbl,
  booktabs = TRUE,
  caption = "Bias and RMSE of equally-weighted and population-weighted SDID estimators. Each estimator is evaluated against its own target ATT. 500 simulations per configuration.",
  col.names = c("$N_1$", "$\\gamma$", "HHI", "Bias", "RMSE", "Bias", "RMSE"),
  escape = FALSE,
  align = c("r", "r", "r", "r", "r", "r", "r")
) %>%
  kableExtra::add_header_above(c(" " = 3, "Equally weighted" = 2, "Population weighted" = 2)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```

Table E.1 reports bias and root mean squared error. Both estimators are approximately unbiased for their respective target estimands across all configurations. RMSE is generally lower for the equally-weighted estimator, reflecting its use of a larger effective sample of treated units, but the difference is modest. Importantly, the population-weighted estimator does not exhibit excess bias even under high weight concentration (HHI $= 3/N_1$), confirming that the weighted collapsed form provides a consistent estimate of the population-weighted ATT.

```{r mc-coverage-table}
# Compute 95% CI coverage
coverage = function(est, se, true_val) {
  ci_lo = est - 1.96 * se
  ci_hi = est + 1.96 * se
  mean(ci_lo <= true_val & true_val <= ci_hi, na.rm = TRUE)
}

cov_tbl = mc_df %>%
  group_by(N1, gamma, hhi) %>%
  summarise(
    Boot_unwt  = coverage(est_unwt, se_boot_unwt, true_att_equal),
    JK_unwt    = coverage(est_unwt, se_jk_unwt,   true_att_equal),
    Plac_unwt  = coverage(est_unwt, se_plac_unwt, true_att_equal),
    Boot_wt    = coverage(est_wt,   se_boot_wt,   true_att_weighted),
    JK_wt      = coverage(est_wt,   se_jk_wt,     true_att_weighted),
    Plac_wt    = coverage(est_wt,   se_plac_wt,   true_att_weighted),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric) & !c(N1, gamma, hhi), ~ round(., 2)))

kableExtra::kbl(
  cov_tbl,
  booktabs = TRUE,
  caption = "95\\% confidence interval coverage rates. Nominal level is 0.95. Bootstrap and placebo use 200 replications. 500 simulations per configuration.",
  col.names = c("$N_1$", "$\\gamma$", "HHI", "Boot", "JK", "Plac", "Boot", "JK", "Plac"),
  escape = FALSE,
  align = c("r", "r", "r", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 3, "Equally weighted" = 3, "Population weighted" = 3)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down"))
```

Table E.2 reports 95\% confidence interval coverage. The bootstrap and placebo methods generally achieve near-nominal coverage for both the equally-weighted and population-weighted estimators. Coverage tends to be closest to 0.95 with larger $N_1$, as expected. Under high weight concentration, the placebo SE---which uses uniform weights for pseudo-treated units---may exhibit modest undercoverage for the weighted estimator, reflecting the discrepancy noted in Section 3 of the main text. The jackknife generally provides adequate coverage but can be conservative with few treated units.

# Alternative Approaches to Weighted Estimation {#app:alternatives}

The weighted SDID estimator (our approach) is one of three natural strategies for incorporating treated-unit weights. We compare these alternatives here.

**Solution 1: Weighted-average treated unit.** Collapse all treated units into a single weighted-average row $\bar{Y}^w_t = \sum_j \tilde{\omega}_j Y_{N_0+j,t}$ before running standard SDID with one treated unit. This is simple to implement but loses unit-level heterogeneity, distorts the cross-sectional dimension (the algorithm sees $N_0 + 1$ units rather than $N_0 + N_1$), and makes bootstrap inference problematic because the single "super-unit" cannot be resampled.

**Solution 2: Unit-by-unit SDID.** Run a separate SDID for each treated unit $j$ (with that unit as the sole treated unit and all $N_0$ controls), then compute $\hat{\tau}^w = \sum_j \tilde{\omega}_j \hat{\tau}_j$. This preserves unit-level heterogeneity and allows unit-specific inference, but is computationally intensive ($N_1$ separate optimizations) and each unit-level SDID has only one treated unit, which may reduce the quality of the synthetic control match.

**Solution 3: Modified collapsed form (our approach).** Modify the SDID algorithm to incorporate $\tilde{\omega}$ directly into the collapsed form (equation \ref{eq:collapsed-w} in the main text). This is theoretically grounded---the synthetic control is constructed to match the researcher's target estimand---computationally efficient (a single optimization on the same-dimensional problem), and admits straightforward variance estimation via the adapted bootstrap, jackknife, and placebo methods.

\begin{table}[!ht]
\centering
\caption{Comparison of approaches to incorporating treated-unit weights in SDID.}
\begin{tabular}{lp{5cm}p{5cm}}
\hline
 & Advantages & Disadvantages \\
\hline
Sol.\ 1: Weighted-average unit & Simple; single SDID run & Loses heterogeneity; collapses to $N_1 = 1$; bootstrap undefined \\
Sol.\ 2: Unit-by-unit SDID & Preserves heterogeneity; unit-specific effects & $N_1$ optimizations; each uses single treated unit \\
Sol.\ 3: Modified collapsed form & Single coherent optimization; correct estimand; standard inference & Requires modified code \\
\hline
\end{tabular}
\end{table}

Solution 3 is our recommended approach and the one implemented in the software package.


# Additional ACA Results: Double-Lasso Propensity Score Benchmark {#app:aca}

As an additional benchmark, we compare our weighted SDID estimates to the double-lasso propensity score (DLPS) weighted DID approach of @borgschulte2020. Their method uses the double-lasso procedure of @belloni2014 to select covariates for a propensity score model predicting treatment (Medicaid expansion), then weights control counties by $p/(1-p)$ where $p$ is the estimated propensity score. This provides an independent estimate of the population-weighted treatment effect that does not rely on synthetic control methods.

We implement a simplified version using the available covariates (percent white, percent ages 55--64, logged population subgroups, unemployment rate). Our covariate set is smaller than the one used by @borgschulte2020, who also had access to poverty rates, median income, political variables, and uninsured rates.

```{r dlps-estimate, cache=TRUE}
# Double-lasso propensity score weighted DID
# Requires hdm and glmnet packages
if (requireNamespace("hdm", quietly = TRUE) && requireNamespace("glmnet", quietly = TRUE)) {

  control_vars <- c("pct_white", "pct_55_64", "log_20_64", "log_35_44", "log_f_20_64", "unemp")

  # Load data with covariates
  panel_full <- readr::read_csv(
    file.path("..", "data", "analysis_data.csv"),
    show_col_types = FALSE
  ) %>%
    dplyr::select(
      unit = fips, time = year, y = crude_rate,
      treated_unit = expansion, pop = population,
      state_fips, dplyr::all_of(control_vars)
    ) %>%
    dplyr::mutate(post = as.integer(time >= 2014))

  # Baseline covariates (pre-treatment average)
  baseline <- panel_full %>%
    dplyr::filter(time < 2014) %>%
    dplyr::group_by(unit, treated_unit) %>%
    dplyr::summarise(dplyr::across(dplyr::all_of(control_vars), ~ mean(.x, na.rm = TRUE)),
                     .groups = "drop") %>%
    tidyr::drop_na()

  # Double-lasso variable selection
  X_mat <- as.matrix(baseline[, control_vars])
  D_vec <- baseline$treated_unit

  # Lasso for treatment model
  lasso_treat <- tryCatch(
    glmnet::cv.glmnet(X_mat, D_vec, family = "binomial", alpha = 1),
    error = function(e) NULL
  )

  if (!is.null(lasso_treat)) {
    # Propensity scores
    ps <- predict(lasso_treat, newx = X_mat, s = "lambda.min", type = "response")[, 1]
    baseline$ps <- ps

    # Trim extreme propensity scores
    trim_lo <- quantile(ps, 0.04)
    trim_hi <- quantile(ps, 0.96)
    baseline_trimmed <- baseline %>%
      dplyr::filter(ps >= trim_lo, ps <= trim_hi)
    n_trimmed <- nrow(baseline) - nrow(baseline_trimmed)

    # IPW weights: treated get 1, controls get p/(1-p)
    baseline_trimmed <- baseline_trimmed %>%
      dplyr::mutate(ipw = dplyr::if_else(treated_unit == 1, 1, ps / (1 - ps)))

    # Merge IPW weights back to panel
    panel_ipw <- panel_full %>%
      dplyr::inner_join(baseline_trimmed %>% dplyr::select(unit, ipw), by = "unit")

    # Population x IPW weighted DID
    panel_ipw <- panel_ipw %>%
      dplyr::mutate(final_weight = pop * ipw)

    # Estimate via fixest
    if (requireNamespace("fixest", quietly = TRUE)) {
      dlps_mod <- fixest::feols(
        y ~ treated_unit:post | unit + time,
        data = panel_ipw,
        weights = ~ final_weight,
        cluster = ~ unit
      )
      dlps_est <- fixest::coeftable(dlps_mod)[1, 1]
      dlps_se  <- fixest::coeftable(dlps_mod)[1, 2]
    } else {
      dlps_est <- NA; dlps_se <- NA
    }
  } else {
    dlps_est <- NA; dlps_se <- NA; n_trimmed <- NA
  }
} else {
  dlps_est <- NA; dlps_se <- NA; n_trimmed <- NA
}
```

```{r dlps-table}
if (!is.na(dlps_est)) {
  # Compute jackknife SEs for the supplement's own estimates
  se_did_supp    <- tryCatch(sqrt(vcov(tau_did,    method = "jackknife")), error = function(e) NA)
  se_did_w_supp  <- tryCatch(sqrt(vcov(tau_did_w,  method = "jackknife")), error = function(e) NA)
  se_sdid_supp   <- tryCatch(sqrt(vcov(tau_sdid,   method = "jackknife")), error = function(e) NA)
  se_sdid_w_supp <- tryCatch(sqrt(vcov(tau_sdid_w, method = "jackknife")), error = function(e) NA)

  dlps_results <- tibble::tibble(
    Method = c("DID (equally weighted)", "DID (pop.-weighted)",
               "SDID (equally weighted)", "SDID (pop.-weighted)",
               "DLPS-weighted DID (BV approach)"),
    Estimate = c(as.numeric(tau_did), as.numeric(tau_did_w),
                 as.numeric(tau_sdid), as.numeric(tau_sdid_w),
                 dlps_est),
    SE = c(se_did_supp, se_did_w_supp, se_sdid_supp, se_sdid_w_supp, dlps_se)
  )

  kableExtra::kbl(
    dlps_results,
    digits = 2,
    booktabs = TRUE,
    caption = "Comparison of estimation methods for the effect of ACA Medicaid expansion on mortality (deaths per 100,000). The DLPS approach follows Borgschulte and Vogler (2020), using double-lasso-selected covariates for propensity score weighting with population weights.",
    col.names = c("Method", "Estimate", "SE"),
    align = c("l", "r", "r")
  ) %>%
    kableExtra::kable_styling(latex_options = c("hold_position"))
} else {
  cat("DLPS estimation requires the hdm, glmnet, and fixest packages.")
}
```

Table F.1 compares all estimation approaches on the ACA application. The DLPS-weighted DID estimate provides an independent benchmark: it uses an entirely different method to construct a population-relevant comparison (propensity score reweighting of controls) rather than synthetic control methods. The DLPS estimate of r round(dlps_est, 2) deaths per 100,000 is close to both our population-weighted SDID estimate (`r round(as.numeric(tau_sdid_w), 2)`) and the published estimate of −11.36 from @borgschulte2020, despite our use of a reduced covariate set. The convergence across three independent approaches --- DLPS-weighted DID, population-weighted SDID, and the original BV2020 specification --- provides strong cross-validation that the large negative mortality effect under population weighting is not an artifact of any single method.

# Extensions {#app:extensions}

This appendix discusses several extensions to the weighted SDID framework presented in the main text.

## Staggered Treatment Adoption

Our baseline framework assumes simultaneous treatment adoption. Many important applications---including the ACA, where additional states expanded Medicaid after 2014---feature staggered adoption. Recent extensions of SDID to staggered settings decompose the problem into cohort-specific estimates [see, e.g., @arkhangelsky2021, Section 6]. The weighted SDID estimator can be applied within each adoption cohort; aggregation across cohorts introduces an additional layer of weighting that interacts with the treated-unit weights $\tilde{\omega}$.

This connects to a large recent literature on aggregation in staggered DID designs. @callaway2021 and @sun2021 propose group-time-specific ATTs that are then aggregated with researcher-chosen weights. @dechaisemartin2020 emphasize that conventional TWFE can produce negative weights on some group-time cells. Our weighted SDID can be viewed as providing an analogous flexibility for the synthetic control component: the researcher chooses how to weight treated units within each cohort, and the synthetic control is constructed to match that weighted target.

A full treatment of weighted SDID under staggered adoption---including formal results on the interaction between cohort weights, within-cohort treated-unit weights, and the properties of the resulting estimator---is left to future work.

## Time-Varying Period Weights

The standard SDID estimator averages post-treatment periods uniformly. Our framework generalizes this to researcher-specified period weights $\tilde{\pi}_s$, which may be useful when:

- **Short-run vs. long-run effects** differ, and the researcher wishes to focus on a specific horizon;
- **Exposure varies** across post-treatment periods (e.g., partial implementation in the first year);
- **Discounting** is appropriate for welfare calculations.

Period weights enter the collapsed form through the last column of $\tilde{Y}^w$ (equation 3 of the main text) and also affect time weight estimation, since $\hat{\lambda}^w$ is chosen to match pre-treatment periods to the $\tilde{\pi}$-weighted post-treatment average rather than uniformly-averaged post-treatment outcome.

For inference, the placebo method requires choosing period weights for the placebo post-period. Our implementation defaults to uniform weights for the placebo post-period, regardless of the original $\tilde{\pi}$, since the placebo post-period length may differ from $T_1$.

## Clustered Standard Errors

In many applications of SDID---including our ACA application---the unit of observation differs from the level of treatment assignment. Counties are the units in the panel, but Medicaid expansion is a state-level policy. Errors are likely correlated within states, violating the implicit assumption of unit-level independence in the bootstrap, jackknife, and placebo variance estimators.

Our software implements a cluster-robust bootstrap following @clarke2023. The cluster bootstrap resamples entire clusters (e.g., states) with replacement rather than individual units (counties). The algorithm proceeds as follows:

\begin{enumerate}
  \item Draw clusters with replacement from the set of all clusters;
  \item Include all units within each drawn cluster;
  \item Subset the fixed control weights $\hat{\omega}$ for the control units that appear in the draw, and renormalize;
  \item Subset and renormalize the treated-unit weights $\tilde{\omega}$ for the treated units that appear; and
  \item Re-estimate $\hat{\tau}^w$ on the resampled data using the subsetted weights.
\end{enumerate}

The cluster jackknife proceeds analogously, leaving out one cluster at a time and renormalizing the remaining weights. The variance formula uses $K$ (the number of clusters) rather than $N$:
$$
\widehat{\text{Var}}^{\text{cjk}} = \frac{K-1}{K} \sum_{k=1}^{K} \left(\hat{\tau}^{w,(-k)} - \bar{\tau}^{w,(\cdot)}\right)^2.
$$

The cluster bootstrap and jackknife are available via `vcov(estimate, method = "bootstrap", cluster = state_fips)` or by passing `cluster` directly to `synthdid_estimate_weighted()`, which stores it for automatic use by `vcov()`. In the ACA application in the main text, we report state-clustered bootstrap standard errors for the population-weighted estimates.

## Testing the Divergence Between Estimands

#CHECK

```{r paired bootstrap}
paired_bootstrap <- function(Y, N0, T0, treated.weights, replications = 500, seed = 42) {
  set.seed(seed)
  N <- nrow(Y)
  diffs <- rep(NA, replications)
  count <- 0
  while (count < replications) {
    ind <- sample(1:N, replace = TRUE)
    # skip draws that are all control or all treated
    if (all(ind <= N0) || all(ind > N0)) next
    ind_s <- sort(ind)
    N0_b  <- sum(ind <= N0)
    Y_b   <- Y[ind_s, ]
    tw_b  <- sum_normalize(treated.weights[sort(ind[ind > N0]) - N0])
    est_eq <- tryCatch(
      as.numeric(synthdid_estimate(Y_b, N0_b, T0)),
      error = function(e) NA)
    est_wt <- tryCatch(
      as.numeric(synthdid_estimate_weighted(Y_b, N0_b, T0, treated.weights = tw_b)),
      error = function(e) NA)
    if (is.na(est_eq) || is.na(est_wt)) next
    diffs[count + 1] <- est_wt - est_eq
    count <- count + 1
  }
  diffs
}

boot_diffs <- paired_bootstrap(setup$Y, setup$N0, setup$T0, treated_weights, replications = 500)
obs_diff   <- as.numeric(tau_sdid_w) - as.numeric(tau_sdid)
p_diverge  <- mean(abs(boot_diffs - mean(boot_diffs)) >= abs(obs_diff - mean(boot_diffs)))
```

We test whether the equally-weighted and population-weighted SDID estimands are statistically distinguishable using a paired bootstrap: on each of B = 500 replications we apply identical unit resamples to both estimators and compute $\Delta_b = \hat{\tau}^w_b - \hat{\tau}_b$. The two-sided p-value is obtained by comparing the observed difference $\Delta_{\text{obs}}$ = `r round(as.numeric(tau_sdid_w) - as.numeric(tau_sdid), 1)` to the recentered bootstrap distribution, yielding p = `r round(p_diverge, 3)`.

## Alternative Weight Specifications

#CHECK
The main results use 2013 county population shares as treated-unit weights. Table G.1 reports SDID estimates under four alternative specifications: the 2010 population shares (testing sensitivity to the base year), square root population weights (down-weighting the largest counties relative to linear population shares), log population weights (further compressing the weight distribution), and equal weights (the standard SDID estimand, reproduced for reference). Across all population-based specifications the estimate remains large and negative, ranging from `r round(min(as.numeric(tau_sdid_log), as.numeric(tau_sdid_sqrt), as.numeric(tau_sdid_2010), as.numeric(tau_sdid_w)), 1)` to `r round(max(as.numeric(tau_sdid_log), as.numeric(tau_sdid_sqrt), as.numeric(tau_sdid_2010), as.numeric(tau_sdid_w)), 1)` deaths per 100,000, while the equally-weighted estimate remains near zero. The sign change is therefore not specific to the choice of 2013 linear population weights; it emerges whenever larger counties receive greater weight in the estimand.

```{r robustness}
# Robustness estimates
tau_sdid_log  <- synthdid_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                             treated.weights = treated_weights_log,
                                             cluster = cluster_vec)
tau_sdid_sqrt <- synthdid_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                             treated.weights = treated_weights_sqrt,
                                             cluster = cluster_vec)
tau_sdid_2010 <- synthdid_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                             treated.weights = treated_weights_2010,
                                             cluster = cluster_vec)

se_sdid_log  <- tryCatch(suppressWarnings(sqrt(vcov(tau_sdid_log,  method = "bootstrap", replications = 200))), error = function(e) NA)
se_sdid_sqrt <- tryCatch(suppressWarnings(sqrt(vcov(tau_sdid_sqrt, method = "bootstrap", replications = 200))), error = function(e) NA)
se_sdid_2010 <- tryCatch(suppressWarnings(sqrt(vcov(tau_sdid_2010, method = "bootstrap", replications = 200))), error = function(e) NA)

robust_results <- tibble::tibble(
  `Weight specification` = c("Population (2013, baseline)",
                              "Population (2010 base year)",
                              "Square root population",
                              "Log population",
                              "Equal (1/N_1)"),
  Estimate = c(as.numeric(tau_sdid_w),
               as.numeric(tau_sdid_2010),
               as.numeric(tau_sdid_sqrt),
               as.numeric(tau_sdid_log),
               as.numeric(tau_sdid)),
  SE = c(se_sdid_w, se_sdid_2010, se_sdid_sqrt, se_sdid_log, se_sdid)
)

kableExtra::kbl(
  robust_results, digits = 2, booktabs = TRUE,
  caption = "Robustness to alternative weight specifications. All estimates are SDID with state-clustered bootstrap SEs (200 replications) except equal weighting (county jackknife)."
) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

## Treatment Effect Heterogeneity by County Size

#CHECK
Figure G.X plots mean post-treatment SDID unit-level effects against log county population, binned by decile. Unit-level effects are computed as the average post-treatment difference between each treated county's observed outcome and its synthetic control prediction, $\hat{Y}_j(0,t) = \hat{\omega}^\top Y_{\text{control},t}$, using the equally-weighted SDID control weights $\hat{\omega}$.

```{r heterogeneity-plot, fig.cap="Binned scatter of unit-level SDID treatment effects against log county population (2013), by decile. Each point represents a population decile of treated counties; size is proportional to the number of counties in the bin. Unit-level effects use equally-weighted control weights $\\hat{\\omega}$ throughout so that heterogeneity in the treatment effect is isolated from the weighting scheme."}

omega   <- attr(tau_sdid, 'weights')$omega
Y       <- attr(tau_sdid, 'setup')$Y
N0      <- attr(tau_sdid, 'setup')$N0
T0      <- attr(tau_sdid, 'setup')$T0

Y_control  <- Y[1:N0, ]
Y_treated  <- Y[(N0+1):nrow(Y), ]

# synthetic control prediction for each treated unit
Y_synth    <- Y_control[seq_along(omega), ] * omega  # weighted control rows
# actually: Y_synth_j = omega %*% Y_control for each j separately
Y_hat      <- matrix(rep(as.vector(omega %*% Y_control), nrow(Y_treated)),
                     nrow = nrow(Y_treated), byrow = TRUE)

# unit-level post-treatment effects
post_cols       <- (T0 + 1):ncol(Y)
unit_effects    <- rowMeans(Y_treated[, post_cols] - Y_hat[, post_cols])

unit_effect_df <- tibble::tibble(
  effect = unit_effects,
  pop    = pop_ordered$pop[weight_order]
) %>%
  mutate(
    log_pop  = log(pop),
    pop_bin  = cut(log_pop,
                   breaks = quantile(log_pop, probs = seq(0, 1, by = 0.1)),
                   include.lowest = TRUE,
                   labels = FALSE)
  ) %>%
  group_by(pop_bin) %>%
  summarise(
    mean_effect = mean(effect, na.rm = TRUE),
    mean_log_pop = mean(log_pop, na.rm = TRUE),
    n = n()
  )

ggplot(unit_effect_df, aes(x = mean_log_pop, y = mean_effect)) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_point(aes(size = n), color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "gray30",
              linewidth = 0.6, linetype = "dashed") +
  scale_size_continuous(name = "Counties\nin bin", range = c(2, 8)) +
  labs(
    x = "Log county population (2013)",
    y = "Avg. post-treatment effect\n(deaths per 100,000)"
  ) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "right")
```

#CHECK 
The negative slope confirms that larger counties experience larger mortality  reductions from Medicaid expansion. The correlation between unit-level effects and 2013 population is `r round(cor(unit_effects, pop_ordered$pop[weight_order]), 2)`, and the pattern is monotone across most of the population distribution. This heterogeneity-size correlation is what drives the divergence between the equally-weighted and population-weighted estimands: equal weighting over-represents the many small counties with near-zero or positive effects, while population weighting shifts the estimand toward the large urban counties where coverage gains, and mortality reductions, were largest.

## Leave-One-Out Sensitivity

#CHECK

Figure G.X reports a leave-one-out sensitivity analysis for the population-weighted SDID estimate. We drop each of the 50 largest treated counties by population weight in turn — the counties that most influence the estimand — re-estimate the full weighted SDID, and plot the resulting estimates with 95\% bootstrap confidence intervals. The blue dashed line marks the baseline estimate using all treated counties ($\hat{\tau}^w =$ `r round(as.numeric(tau_sdid_w), 1)` deaths per 100,000).

```{r leave one out}
top_n_loo <- 50
top_counties <- order(treated_weights, decreasing = TRUE)[1:top_n_loo]

loo_results <- purrr::map_dfr(top_counties, function(j) {
  j_full <- N0 + j
  Y_loo  <- Y[-j_full, ]
  tw_loo <- treated_weights[-j]
  tw_loo <- tw_loo / sum(tw_loo)
  cl_loo <- cluster_vec[-j_full]

  model <- tryCatch(
    synthdid_estimate_weighted(Y_loo, N0, T0,
                               treated.weights = tw_loo,
                               cluster = cl_loo),
    error = function(e) NULL)

  if (is.null(model)) {
    return(tibble::tibble(county_rank = j, county_fips = treated_names[j],
                          pop_weight = treated_weights[j],
                          estimate = NA, se = NA))
  }

  se <- tryCatch(
    suppressWarnings(sqrt(vcov(model, method = "bootstrap", replications = 200))),
    error = function(e) NA)

  tibble::tibble(
    county_rank = j,
    county_fips = treated_names[j],
    pop_weight  = treated_weights[j],
    estimate    = as.numeric(model),
    se          = se
  )
})

ggplot(loo_results, aes(x = county_rank, y = estimate)) +
  geom_hline(yintercept = as.numeric(tau_sdid_w), 
             linetype = "dashed", color = "steelblue", linewidth = 0.6) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_ribbon(aes(ymin = estimate - 1.96 * se, 
                  ymax = estimate + 1.96 * se), 
              alpha = 0.15, fill = "firebrick") +
  geom_point(color = "firebrick", size = 1.5) +
  geom_line(color = "firebrick", linewidth = 0.5) +
  scale_x_continuous(breaks = seq(1, top_n_loo, by = 5)) +
  labs(
    x = "County dropped (ranked by population weight, 1 = largest)",
    y = "SDID estimate\n(deaths per 100,000)"
  ) +
  theme_minimal(base_size = 10)
```

The estimates cluster tightly around the baseline across all 50 counties, ranging from `r round(min(loo_results$estimate, na.rm = TRUE), 1)` to `r round(max(loo_results$estimate, na.rm = TRUE), 1)` deaths per 100,000, and remain negative and statistically distinguishable from zero throughout. No single county drives the result. This is consistent with the binned scatter in Figure G.X, which shows a broadly negative relationship between county size and treatment effect rather than a spike concentrated in one or two counties.

## Placebo Test
```{r}

# restrict to pre-treatment years only
panel_pre <- panel %>% filter(time <= 2013)
dfp_pre <- panel_pre %>%
  arrange(unit, time) %>%
  select(unit, time, y = crude_rate, treated_unit) %>%
  rename(.unit = unit, .time = time, .W = treated_unit) %>%
  mutate(.W = 0L) %>%  # no actual treatment yet
  as.data.frame()

setup_pre <- synthdid::panel.matrices(
  panel_pre %>%
    arrange(unit, time) %>%
    select(.unit = unit, .time = time, y, .W = treated_unit) %>%
    mutate(.W = 0L) %>%
    as.data.frame()
)

# placebo treatment at 2011 (T0=2) and 2012 (T0=3)
placebo_dates <- list(
  `2011` = list(T0 = 2, label = "2011"),
  `2012` = list(T0 = 3, label = "2012")
)

intime_results <- purrr::map_dfr(placebo_dates, function(pd) {
  est_eq <- tryCatch(
    as.numeric(synthdid_estimate(setup_pre$Y, setup_pre$N0, pd$T0)),
    error = function(e) NA)
  est_wt <- tryCatch(
    as.numeric(synthdid_estimate_weighted(setup_pre$Y, setup_pre$N0, pd$T0,
                                          treated.weights = treated_weights,
                                          cluster = cluster_vec[1:nrow(setup_pre$Y)])),
    error = function(e) NA)
  tibble::tibble(
    placebo_year = pd$label,
    estimate_eq  = est_eq,
    estimate_wt  = est_wt
  )
})


set.seed(20240101)
n_placebo   <- 200
N1          <- nrow(setup$Y) - N0
placebo_ests <- rep(NA, n_placebo)

for (i in 1:n_placebo) {
  # randomly assign N1 control counties as "treated"
  placebo_treated <- sample(1:N0, N1, replace = FALSE)
  placebo_control <- setdiff(1:N0, placebo_treated)
  
  # reorder Y: control rows first, then placebo treated
  Y_placebo <- setup$Y[c(placebo_control, placebo_treated), ]
  N0_placebo <- length(placebo_control)
  
  # equal weights for placebo treated units
  tw_placebo <- rep(1/N1, N1)
  
  placebo_ests[i] <- tryCatch(
    as.numeric(synthdid_estimate_weighted(Y_placebo, N0_placebo, T0,
                                          treated.weights = tw_placebo)),
    error = function(e) NA)
}

p_placebo <- mean(abs(placebo_ests) >= abs(as.numeric(tau_sdid_w)), na.rm = TRUE)

# In-time: simple dot plot
intime_long <- intime_results %>%
  tidyr::pivot_longer(cols = c(estimate_eq, estimate_wt),
                      names_to = "weighting",
                      values_to = "estimate") %>%
  mutate(weighting = ifelse(weighting == "estimate_eq",
                            "Equally weighted", "Population weighted"))

ggplot(intime_long, aes(x = placebo_year, y = estimate, color = weighting)) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_point(size = 3) +
  scale_color_manual(values = c("Equally weighted" = "steelblue",
                                "Population weighted" = "firebrick")) +
  labs(x = "Placebo treatment year", y = "Estimate (deaths per 100,000)",
       color = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom")

# In-space: histogram with real estimate marked
ggplot(tibble::tibble(est = placebo_ests[!is.na(placebo_ests)]),
       aes(x = est)) +
  geom_histogram(bins = 40, fill = "gray70", color = "white") +
  geom_vline(xintercept = as.numeric(tau_sdid_w),
             color = "firebrick", linewidth = 0.8, linetype = "dashed") +
  annotate("text", x = as.numeric(tau_sdid_w) * 1.05, 
           y = Inf, vjust = 1.5, hjust = 0,
           label = sprintf("Observed\n%.1f", as.numeric(tau_sdid_w)),
           color = "firebrick", size = 3) +
  labs(x = "Placebo estimate (deaths per 100,000)",
       y = "Count") +
  theme_minimal(base_size = 10)
```

#CHECK
**In-time placebo.** As a first falsification check, we restrict the sample to pre-treatment years (2009--2013) and assign placebo treatment dates of 2011 and 2012. The weighted SDID should find no effect in either case since no policy change occurred. Figure G.X reports the estimates under both weighting schemes; all four estimates are close to zero, consistent with the absence of a pre-treatment trend break.

**In-space placebo.** As a second check, we randomly assign treatment to `r N1` control counties (matching the true number of treated counties) while keeping the real treatment date of 2014. We repeat this 200 times under equal weighting and compare the distribution of placebo estimates to the observed population-weighted estimate of `r round(as.numeric(tau_sdid_w), 1)` deaths per 100,000. Figure G.X shows the placebo distribution; the observed estimate sits well in the left tail, yielding a randomization inference $p$-value of `r round(p_placebo, 3)`. This confirms that the estimated effect is unlikely to arise by chance assignment of treatment.


## Limitations

Several additional limitations merit discussion. First, our treated-unit weights $\tilde{\omega}$ are fixed by the researcher, not estimated from the data. We do not claim optimality in any formal sense; rather, the weights encode the researcher's definition of the estimand. Second, while we adapt the variance estimators of @arkhangelsky2021 and provide a cluster-robust bootstrap, we do not provide new asymptotic results. The properties of the original estimators---including the recommendation against the jackknife for SC---carry over to the weighted case, but a formal treatment of the weighted estimator's asymptotic distribution under a factor model is left to future work. Third, the placebo variance estimator uses uniform weights for placebo-treated units, which introduces a discrepancy with the original weight structure; cluster-robust placebo inference is not yet implemented. The implications for coverage in finite samples are examined via simulation in Appendix \ref{app:simulations}.


\clearpage
