---
title: "Weighted Synthetic Difference-in-Differences"
author: "John Iselin\\thanks{The Budget Lab at Yale. Corresponding author: john.iselin@yale.edu} \\and Erica Ryan\\thanks{Amazon}"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  pdf_document:
    latex_engine: pdflatex
    number_sections: true
    keep_tex: true
    fig_caption: true
    citation_package: natbib
    includes:
      in_header: header.tex
bibliography: references.bib
biblio-style: elsarticle-harv
fontsize: 11pt
geometry: margin=1in
linestretch: 1.4
abstract: |
 We extend synthetic difference-in-differences (SDID) to allow researcher-specified treated-unit weights. This extension, implemented via a single modification of the collapsed form of @arkhangelsky2021, allows researchers to specify the estimand of interest. We adapt the bootstrap, jackknife, and placebo variance estimators in two ways: allowing treated-unit weights and adding cluster-robust versions. We further establish consistency and asymptotic normality of the weighted estimator under a condition requiring the effective number of treated units to grow. Because that condition is strained in many applications, including ours, we recommend cluster-robust bootstrap inference, and we propose a paired-bootstrap test of the gap between estimands as an ex-ante practitioner diagnostic. The estimator extends the \texttt{synthdid} R package. We demonstrate the value of this extension by studying the county-level mortality effect of the 2014 ACA Medicaid expansion. Replacing the equal treated-unit weighting of SDID with population weights moves the estimate from a precise null to a reduction of roughly 17 deaths per 100,000. The divergence is driven entirely by the choice of estimand, not of estimator, and reflects treatment effects concentrated in the largest urban counties. In this setting, however, the population-weighted estimand fails an in-time placebo test, reflecting a differential pre-trend in the largest counties that convex, level-matched designs cannot remove. We argue this is a feature of the ACA county setting---it equally afflicts the published propensity-score benchmark---rather than of weighting per se, and that credibly identifying a population-weighted effect calls for the triple-difference and administrative-data designs of the recent mortality literature.

  \medskip\noindent
  \textbf{Keywords:} Synthetic difference-in-differences, causal inference, panel data, heterogeneous treatment effects, Medicaid expansion

  \medskip\noindent
  \textbf{JEL Codes:} C21, C23, I13
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4,
  fig.align = "center",
  out.width = "90%",
  cache = TRUE
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)

# This build only reads and formats the committed results/ tables (see the
# load-results chunk); it computes nothing from the synthdid functions, so the
# package/source is not loaded here.
set.seed(20240101)

# Shared series aesthetics for the two-series (equal vs. population) figures,
# designed for black-and-white readability via REDUNDANT encoding: hue +
# luminance (Okabe-Ito blue/orange differ in luminance, so they stay distinct in
# grayscale) + point shape + line type. Reference these everywhere instead of
# re-specifying scales inline, so the encoding is consistent across figures.
w_pal <- c("Equally weighted" = "#0072B2", "Population weighted" = "#E69F00")
w_lty <- c("Equally weighted" = "solid",   "Population weighted" = "dashed")
w_shp <- c("Equally weighted" = 16,        "Population weighted" = 17)
```

\setcounter{secnumdepth}{3}

# Introduction

The synthetic difference-in-differences (SDID) estimator of @arkhangelsky2021 has become a workhorse for causal inference in panel settings, combining the strengths of the synthetic control method (SCM) [@abadie2010] and difference-in-differences by using data-driven unit weights $\hat{\omega}$ and time weights $\hat{\lambda}$ while permitting level differences across units. When multiple units are treated, the standard SDID implementation averages them equally: each treated unit contributes $1/N_1$ to both the target estimand and the construction of the synthetic control. This convention likely reflects the origins of the estimator in the SCM, which was developed for settings with a single treated unit [@abadie2010].

This way of accounting for multiple treated units means that when treatment effects are heterogeneous and correlated with unit size, the equally-weighted estimand can diverge sharply from estimands that researchers actually care about: for example, per-capita treatment effects, average effects on the population exposed to treatment, or effects on subpopulations of policy interest. A growing literature has emphasized that aggregation choices in panel estimators are not innocuous: @gibbons2019 show that two-way fixed-effects regressions implicitly weight observations in ways that may diverge from any natural causal estimand; @sloczynski2022 extends this critique to OLS with heterogeneous effects; @callaway2021 and @dechaisemartin2020 develop frameworks for staggered-adoption settings in which conventional aggregations can produce negatively-weighted combinations of unit-specific effects. 

We make three contributions. First, we develop a weighted SDID estimator that replaces the uniform $1/N_1$ averaging with researcher-chosen weights $\tilde{\omega}_j$ in the collapsed form of the outcome matrix. The modification is parsimonious (a single change to the optimization target) and propagates naturally through the data-driven control and time weights $\hat{\omega}^w, \hat{\lambda}^w$. Standard SDID is recovered when $\tilde{\omega}_j = 1/N_1$, and weighted DID and weighted SCM emerge as boundary cases. Second, we establish consistency and asymptotic normality of the weighted estimator under conditions that inherit from the original SDID factor-model framework, supplemented by a Lindeberg-type requirement that the *effective* number of treated units, the Kish $N_1^{\text{eff}} = (\sum_j \tilde\omega_j^2)^{-1}$, grow with $N_1$. We document that this condition loses its practical bite in many policy-relevant applications, including ours, and motivate alternative inference procedures as the appropriate finite-sample tool. Third, we adapt all three variance estimators of @arkhangelsky2021 (bootstrap, jackknife, and placebo) to the weighted case with appropriate weight renormalization, and extend the bootstrap to handle clustered treatment assignment following @clarke2023. The full set of estimators is implemented as an extension to the \texttt{synthdid} R package.\footnote{Available at \url{https://github.com/johniselin-econ/synthdid_weights}. Replication code and a companion supplementary appendix with proofs and additional robustness checks are also provided.}

We apply the method to the county-level mortality effects of the ACA's 2014 Medicaid expansion, following the design of @borgschulte2020. They use double-lasso propensity score (DLPS) weighting to address pre-existing differences in mortality trends between expansion and non-expansion counties---a setting where SDID's data-driven reweighting is a natural alternative. Treated counties range from under 10,000 to over 10 million residents, making the weighting choice consequential. The equally-weighted (i.e., standard) SDID estimate is statistically indistinguishable from zero. The population-weighted SDID estimate is approximately $-17.5$ deaths per 100,000, significant at the 1\% level and of comparable magnitude to the $-11.36$ that @borgschulte2020 obtain via DLPS weighting. The divergence from the equally-weighted estimate is driven entirely by the weighting choice, not the estimator: a parallel weighted DID exercise produces qualitatively identical results. A binned scatter of unit-level treatment effects against log county population shows that effects in small and mid-sized counties oscillate around zero while the largest urban counties exhibit reductions exceeding $25$ deaths per 100,000---the heterogeneity-size gradient that drives the result. The estimate is robust to alternative population-weight specifications (2010 vs.\ 2013 base years, square-root population), to leave-one-out exclusion of large counties, and to an in-space placebo test. An in-time placebo, however, does not pass for the population-weighted target: it detects a pre-existing differential mortality decline in the largest counties that SDID's convex, simplex-constrained reweighting cannot fully remove. The same falsification fails for a parallel weighted DID and for the level-matched propensity-score design of @borgschulte2020, while the equally-weighted target---in which the largest counties are a handful among equals---is essentially unexposed and its placebos pass. We read this as a property of the identification demands of the ACA county setting rather than a defect of the weighting, and we do not report a trend-corrected population-weighted magnitude for a design that cannot separate a differential trend from genuine early-coverage gains. The divergence between estimands nonetheless survives intact: the estimate is *not* robust to compressing the weight distribution toward equality (log population shares are sufficient to eliminate the effect), confirming that the gap between estimands is real and quantitatively meaningful whatever the true population-weighted level.

Our application also speaks to a rapidly growing literature estimating the mortality effects of the ACA Medicaid expansions. @miller2021 and, most recently, @wyse2025 use large linked administrative datasets to estimate population-level mortality reductions; @wyse2025, using the near-universe of 37 million low-income adults, report a 2.5 percent reduction in the population mortality hazard (a 21 percent reduction for enrollees). Their headline is a *population-weighted* estimand: because identification is at the state level and observations are weighted to the population, a handful of large expansion states dominate the treated mass---the same concentration that drives the divergence we study, here across states rather than counties. Their identification rests on parallel trends between expansion and non-expansion states, defended with event-study pre-trends and a triple-difference that nets out trends common to a higher-income comparison group. Notably, their triple difference nets out trends common to a higher-income comparison group and largely survives its trend check, whereas our single-difference county design does not; this contrast underscores our central point, that whether population weighting merely re-selects an estimand or also imports a differential-trend identification burden is an empirical question specific to the application, resolved only by a formal placebo rather than by inspection of event-study coefficients. The design that answers it is orthogonal to, and composable with, the weighting choice: the weighted estimator selects the estimand, and a triple-difference or administrative-data design supplies the identification that the estimand demands.

Section 2 introduces the weighted estimator, develops its asymptotic properties, and discusses inference; Section 3 applies the method to the ACA Medicaid expansion; Section 4 reports Monte Carlo evidence on finite-sample bias and confidence-interval coverage; Section 5 concludes.


# The Weighted SDID Estimator

## Setup

Consider a balanced panel with $N$ units observed over $T$ periods. The first $N_0$ units are untreated controls; the remaining $N_1 = N - N_0$ are treated beginning at period $T_0 + 1$, with $T_1 = T - T_0$ post-treatment periods. Define the treatment indicator $W_{it} = \mathbf{1}\{i > N_0\} \cdot \mathbf{1}\{t > T_0\}$ and observed outcomes $Y_{it} = Y_{it}(0) + \tau_{it} W_{it}$, where $\tau_{it}$ is the unit-time treatment effect.

Following @arkhangelsky2021, the standard SDID estimator solves
\begin{equation} \label{eq:sdid-regression}
(\hat{\tau}^{\text{sdid}}, \hat{\mu}, \hat{\alpha}, \hat{\beta}) = \underset{\tau, \mu, \alpha_i, \beta_t}{\arg\min} \sum_{i=1}^{N} \sum_{t=1}^{T} \left( Y_{it} - \mu - \alpha_i - \beta_t - \tau \cdot W_{it} \right)^2 \hat{\omega}_i \hat{\lambda}_t,
\end{equation}
where $\hat{\omega}_i$ and $\hat{\lambda}_t$ are data-driven unit and time weights. This admits a closed-form solution with a *collapsed form* $\tilde{Y}$, an $(N_0+1) \times (T_0+1)$ matrix in which the last row averages treated units and the last column averages post-treatment periods:
\begin{equation} \label{eq:collapsed}
\tilde{Y}_{N_0+1, t} = \frac{1}{N_1} \sum_{j=1}^{N_1} Y_{N_0+j,t}, \quad
\tilde{Y}_{i, T_0+1} = \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it}.
\end{equation}
Control unit weights $\hat{\omega} \in \Delta^{N_0}$ and time weights $\hat{\lambda} \in \Delta^{T_0}$ solve the penalized matching programs
\begin{align}
(\hat\omega_0, \hat\omega) &= \underset{\omega_0 \in \mathbb{R},\; \omega \in \Delta^{N_0}}{\arg\min} \; \sum_{t=1}^{T_0}\Big(\omega_0 + \sum_{i=1}^{N_0}\omega_i Y_{it} - \tilde Y_{N_0+1, t}\Big)^2 + \zeta^2 T_0 \|\omega\|_2^2, \label{eq:omega-program}\\
(\hat\lambda_0, \hat\lambda) &= \underset{\lambda_0 \in \mathbb{R},\; \lambda \in \Delta^{T_0}}{\arg\min} \; \sum_{i=1}^{N_0}\Big(\lambda_0 + \sum_{s=1}^{T_0}\lambda_s Y_{is} - \tilde Y_{i, T_0+1}\Big)^2, \label{eq:lambda-program}
\end{align}
where $\zeta$ is the regularization level of @arkhangelsky2021: the intercepts permit level differences between the treated average and its synthetic control, and the ridge penalty spreads $\hat\omega$ across control units. Both programs are solved by Frank-Wolfe iterations on the simplex. The treatment effect is the bilinear form
\begin{equation} \label{eq:tau-sdid}
\hat{\tau}^{\text{sdid}} = \begin{pmatrix} -\hat{\omega} \\ \frac{1}{N_1}\mathbf{1}_{N_1} \end{pmatrix}^\top Y \begin{pmatrix} -\hat{\lambda} \\ \frac{1}{T_1}\mathbf{1}_{T_1} \end{pmatrix}.
\end{equation}

The key observation that motivates our extension is that the equal weights $1/N_1$ appear in two distinct places: in the collapsed form \eqref{eq:collapsed} that determines *which* synthetic control is constructed, and in the treated-row block of \eqref{eq:tau-sdid} that determines *which* treated average is being matched. In standard SDID, both are uniform by convention. Researchers who care about a non-uniform target estimand (per-capita effects, effects weighted by exposure or population) should reweight both.

## The Weighted Modification

Let $\tilde{\omega} = (\tilde{\omega}_1, \ldots, \tilde{\omega}_{N_1})$ be researcher-specified non-negative weights on treated units, normalized so $\sum_j \tilde{\omega}_j = 1$. The target estimand is
\begin{equation} \label{eq:estimand}
\tau^{(\tilde{\omega})} = \sum_{j=1}^{N_1} \tilde{\omega}_j \, \bar{\tau}_j, \quad \bar{\tau}_j = T_1^{-1}\sum_{t=T_0+1}^{T} \tau_{N_0+j, t}.
\end{equation}
We modify SDID by replacing the uniform average in the treated row of the collapsed form with the weighted average:
\begin{equation} \label{eq:collapsed-w}
\tilde{Y}^w_{N_0+1, t} = \sum_{j=1}^{N_1} \tilde{\omega}_j Y_{N_0+j,t}, \quad
\tilde{Y}^w_{i, T_0+1} = \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it}.
\end{equation}
Control weights $\hat{\omega}^w$ and time weights $\hat{\lambda}^w$ are estimated from $\tilde{Y}^w$ via the same Frank-Wolfe algorithm (no modification to the solver is needed) and the weighted treatment effect is
\begin{equation} \label{eq:tau-w}
\hat{\tau}^{w} = \begin{pmatrix} -\hat{\omega}^w \\ \tilde{\omega} \end{pmatrix}^\top Y \begin{pmatrix} -\hat{\lambda}^w \\ \frac{1}{T_1}\mathbf{1}_{T_1} \end{pmatrix}.
\end{equation}
The modification is concentrated in two places: the collapsed form that feeds the weight solver, and the treated-unit block of the bilinear form. Equation \eqref{eq:tau-w} preserves the double-difference structure and delivers two boundary cases. Setting $\tilde{\omega}_j = 1/N_1$ for all $j$ recovers $\hat{\tau}^{\text{sdid}}$ exactly: the collapsed form \eqref{eq:collapsed-w} reduces to \eqref{eq:collapsed}, the Frank-Wolfe solutions coincide, and \eqref{eq:tau-w} reduces to \eqref{eq:tau-sdid}. The weighted analogues of the constituent estimators emerge at the remaining boundaries: fixing $\hat{\omega} = \mathbf{1}_{N_0}/N_0$ and $\hat{\lambda} = \mathbf{1}_{T_0}/T_0$ produces a weighted DID estimator, while fixing $\hat{\lambda} = \mathbf{0}$ produces a weighted synthetic control estimator.

The interpretation of \eqref{eq:tau-w} as a sample analogue of \eqref{eq:estimand} is that the synthetic control trajectory is constructed to match the weighted treated trajectory in pre-treatment periods, so that the post-treatment double-difference recovers the weighted average of unit-level treatment effects.

\paragraph{A two-unit example.} The choice of $\tilde\omega$ matters exactly when treatment effects covary with the weights. From \eqref{eq:estimand}, the gap between the weighted and equally-weighted estimands is
$$\tau^{(\tilde\omega)} - \tau^{(1/N_1)} = \sum_{j=1}^{N_1}\left(\tilde\omega_j - \frac{1}{N_1}\right)\bar\tau_j = \sum_{j=1}^{N_1}\left(\tilde\omega_j - \frac{1}{N_1}\right)\left(\bar\tau_j - \bar{\bar\tau}\,\right),$$
where $\bar{\bar\tau} = N_1^{-1}\sum_j \bar\tau_j$ (the second equality holds because the weight deviations sum to zero)---an empirical covariance between weights and unit-level effects. Table 1 walks through a concrete case: $N_1 = 2$, with one large treated unit ($\tilde\omega_1 = 0.9$, $\bar\tau_1 = -15$) and one small unit ($\tilde\omega_2 = 0.1$, $\bar\tau_2 = +10$). The unit-level effects are held fixed across the table's three weighting schemes; only the weights change, and each estimand is the sum of the weight $\times$ effect contributions in its column pair. Population weighting yields $0.9(-15) + 0.1(10) = -12.5$ against the equally-weighted $(-15+10)/2 = -2.5$; the gap of $-10$ is exactly the covariance term above. Under the assumptions of Section 2.3, each estimator is consistent for its own target, so the gap is not a bias of either: the two numbers answer different questions. Heterogeneity alone is not sufficient. Reversing the weights (the final column pair) moves the weighted estimand to $+7.5$ without touching either unit's effect, and homogeneous effects ($\bar\tau_1 = \bar\tau_2 = -2.5$) would make all three schemes return $-2.5$. Both heterogeneous effects *and* weight--effect correlation are required, which is what the binned scatter of Section 3.3 diagnoses in our application.

```{r two-unit-table}
two_unit <- tibble::tibble(
  unit = c("1 (large)", "2 (small)", "Estimand"),
  tau  = c("$-15$",  "$+10$",  ""),
  eq_w = c("$0.5$",  "$0.5$",  ""),
  eq_c = c("$-7.5$", "$+5.0$", "$-2.5$"),
  pw_w = c("$0.9$",  "$0.1$",  ""),
  pw_c = c("$-13.5$", "$+1.0$", "$-12.5$"),
  rv_w = c("$0.1$",  "$0.9$",  ""),
  rv_c = c("$-1.5$", "$+9.0$", "$+7.5$")
)

kableExtra::kbl(
  two_unit, booktabs = TRUE, escape = FALSE,
  caption = "The two-unit example: three estimands from the same unit-level effects.",
  col.names = c("Unit $j$", "$\\bar\\tau_j$",
                "$\\tilde\\omega_j$", "$\\tilde\\omega_j \\bar\\tau_j$",
                "$\\tilde\\omega_j$", "$\\tilde\\omega_j \\bar\\tau_j$",
                "$\\tilde\\omega_j$", "$\\tilde\\omega_j \\bar\\tau_j$"),
  align = c("l", rep("r", 7))
) %>%
  kableExtra::add_header_above(
    c(" " = 2, "Equal" = 2, "Population" = 2, "Reversed" = 2)) %>%
  kableExtra::row_spec(2, hline_after = TRUE) %>%
  kableExtra::kable_styling(latex_options = "hold_position") %>%
  kableExtra::footnote(
    general = paste(
      "Unit-level average effects $\\\\bar\\\\tau_j$ are identical in every column pair; only the",
      "treated-unit weights $\\\\tilde\\\\omega$ change. Each estimand $\\\\tau^{(\\\\tilde\\\\omega)} = \\\\sum_j \\\\tilde\\\\omega_j \\\\bar\\\\tau_j$",
      "is the sum of the contributions above it. The gap between any two estimands equals the",
      "covariance between the weight deviations and the unit-level effects; with homogeneous",
      "effects ($\\\\bar\\\\tau_1 = \\\\bar\\\\tau_2$) all three schemes would agree."),
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```

## Asymptotic Theory

We establish consistency and asymptotic normality of $\hat{\tau}^w$ under conditions that inherit from the factor-model framework of @arkhangelsky2021, supplemented by regularity conditions on the treated-unit weights. The data-generating-process assumptions, stated formally as Assumptions A1--A5 in the supplementary appendix, are: (A1) outcomes follow a low-rank-plus-noise structure $Y_{it} = L_{it} + \tau_{it} W_{it} + \epsilon_{it}$ with $L_{it} = \mu_{it} + \alpha_i + \beta_t$ for some low-rank $\mu$; (A2) sub-Gaussian errors, independent across units and weakly dependent over time; (A3) factor-structure regularity; (A4) sample sizes $N_0, T_0 \to \infty$ with $\log N_0 = o(T_0)$ and $\log T_0 = o(N_0)$; (A5) existence of oracle weights $(\omega^*, \lambda^*)$ on the intercept-augmented simplex that match the weighted treated factor structure. We also impose two conditions on the treated-unit weights. Condition (W1) requires the weights to form a probability vector, $\tilde{\omega}_j \geq 0$ and $\sum_j \tilde{\omega}_j = 1$. Condition (W2) requires that $\max_j \tilde{\omega}_j \to 0$ as $N_1 \to \infty$ or, equivalently, that the Kish effective sample size $N_1^{\text{eff}} := \left(\sum_j \tilde{\omega}_j^2\right)^{-1}$ diverge.

Condition (W2) rules out a single dominant treated unit and is the key requirement for the central-limit machinery to apply to the weighted treated average. It is automatically satisfied by uniform weights ($N_1^{\text{eff}} = N_1$). For population weights, it is satisfied when treated units' size shares decay at any rate (e.g., a Pareto distribution with finite second moment), but it can fail when one or two units carry a non-vanishing share of total weight in the limit. We return to this point below.

\medskip
\noindent\textbf{Proposition 1 (Consistency).} *Under Assumptions A1--A5 of @arkhangelsky2021 and (W1)--(W2),*
$$\hat{\tau}^w \xrightarrow{p} \tau^{(\tilde{\omega})} \quad \text{as } N_0, N_1, T_0 \to \infty.$$

\medskip
\noindent\textit{Proof sketch.} Following the decomposition argument of @arkhangelsky2021, write
$$\hat{\tau}^w - \tau^{(\tilde{\omega})} = \underbrace{\sum_{j} \tilde{\omega}_j \xi_j}_{\text{treated noise (i)}} + \underbrace{\Delta_{\text{ctrl}}}_{\text{control noise (ii)}} + \underbrace{\Delta_{\text{est}}}_{\text{weight estimation (iii)}},$$
where $\xi_j$ is the post-minus-pre noise *contrast* for treated unit $j$ (its post-period noise average minus the oracle-$\lambda^*$-weighted pre-period noise sum), with bounded variance under (A2); the contrast form absorbs within-unit serial covariance between post- and pre-period noise. For (i): $\text{Var}\!\left(\sum_j \tilde\omega_j \xi_j\right) \leq C \sum_j \tilde\omega_j^2 = C / N_1^{\text{eff}} \to 0$ by (W2), so (i) is $o_p(1)$ by Markov's inequality. For (ii): the analogous contrast at the oracle control weights has variance bounded by $C\|\omega^*\|_2^2 \to 0$ under (A5). For (iii): the weighted target column of the collapsed form remains a fixed convex combination of treated trajectories with an oracle representation in the control block (A5), so the weight-estimation error bounds of @arkhangelsky2021 apply with $\tilde\omega$ in place of $1/N_1$, and the surviving cross term, a product of loading-space estimation errors, vanishes under (A3)--(A4). Combining (i)--(iii) gives consistency. $\square$

\medskip
\noindent\textbf{Proposition 2 (Asymptotic normality).} *Under A1--A5, (W1)--(W2), the Lindeberg condition (W3): $\max_j \tilde{\omega}_j^2 / \sum_k \tilde{\omega}_k^2 \to 0$, and a negligibility condition (R) requiring the control-noise and weight-estimation terms to vanish at the $\sqrt{N_1^{\text{eff}}}$ scale,*
$$\sqrt{N_1^{\text{eff}}} \cdot \left( \hat{\tau}^w - \tau^{(\tilde{\omega})} \right) \xrightarrow{d} \mathcal{N}\!\left(0, \sigma_\xi^2\right),$$
*where $\sigma_\xi^2$ is the limiting variance of the standardized weighted treated-noise contrast.*

\medskip
\noindent\textit{Proof sketch.} Component (i) of the decomposition above is a Lindeberg-CLT-eligible weighted sum: $\sqrt{N_1^{\text{eff}}} \sum_j \tilde\omega_j \xi_j$ converges in distribution to $\mathcal{N}(0, \sigma_\xi^2)$ provided no single weight dominates the variance contribution---which is precisely condition (W3). Because $\xi_j$ is a post-minus-pre contrast, $\sigma_\xi^2$ absorbs the within-unit serial covariance between post- and pre-period noise rather than assuming it away. Components (ii) and (iii) are $o_p(1)$ at the $\sqrt{N_1^{\text{eff}}}$ scale under (R); primitive sufficient conditions, inherited from the rate conditions behind @arkhangelsky2021's Theorem 1, are discussed in the supplementary appendix, which also characterizes when (R) is plausible. In finite samples, the bootstrap-based intervals we recommend capture the contributions of (ii)--(iii) regardless of whether (R) holds exactly. $\square$

\medskip
\noindent\textbf{Proposition 3 (Cluster-robust variance).} *Suppose treatment is assigned at the cluster level, with $K$ clusters indexed $k = 1, \ldots, K$, $K \to \infty$ and $\max_k \tilde\omega^{(k)} \to 0$ where $\tilde\omega^{(k)} = \sum_{j \in k} \tilde\omega_j$ is the cluster-aggregated weight, and suppose errors are independent across clusters but arbitrarily dependent within them. Then the cluster bootstrap variance estimator
$$\widehat{V}_w^{\text{cb}} = \frac{1}{B-1}\sum_{b=1}^{B} \left(\hat{\tau}^{w,(b)} - \bar{\tau}^{w}\right)^2$$
*is consistent for the cluster-robust asymptotic variance, which augments $\sigma_\xi^2$ with within-cluster covariance terms, where $\hat\tau^{w,(b)}$ is computed from a cluster-resampled bootstrap draw with weight renormalization $\tilde\omega_j^{(b)} = \tilde\omega_j n_j^{(b)} / \sum_{\ell=1}^{N_1} \tilde\omega_\ell n_\ell^{(b)}$.*

\medskip
\noindent\textit{Proof sketch.} The central limit argument of Proposition 2 is applied one level up, to cluster-aggregated noise contrasts whose variances absorb within-cluster covariances; standard cluster-bootstrap arguments [@cameron2008; @clarke2023] then apply to the bilinear form \eqref{eq:tau-w} provided the cluster-level weights satisfy the no-dominant-cluster condition---the Lindeberg analogue at the cluster level. The renormalization step preserves the convex-combination structure of $\tilde\omega$ within each draw. No bound on cluster sizes is required; what matters is that no cluster's weight share dominates. $\square$

The companion supplementary appendix gives full proofs of all three propositions. Our purpose here is to characterize the regime in which inference based on \eqref{eq:tau-w} has the standard frequentist interpretation.

\paragraph{When the asymptotics lose their bite.} Condition (W2), and its cluster-level analogue in Proposition 3, is not innocuous in applications. When researchers use population weights in panels with one or two dominant units (e.g., California Proposition 99, where one large state is treated; or our ACA application, where Los Angeles County alone carries roughly 6\% of the population-weighted treated mass), the effective sample size cannot exceed the inverse squared share of the largest unit and in practice sits far below $N_1$, however many small units the panel adds. A sequence of such weights may still satisfy (W2) formally, but the convergence is too slow for the asymptotic approximation to be reliable at realistic sample sizes. In these settings, plug-in asymptotic standard errors will understate uncertainty, and one must rely either on cluster-robust bootstrap inference at a level of aggregation where Proposition 3's no-dominant-cluster condition is plausible, or on randomization-inference and placebo-distribution tests that condition on the actual weight structure. We pursue both routes in our application.

## Inference

We adapt the three variance estimators of @arkhangelsky2021 (bootstrap, jackknife, and placebo) to the weighted case. The key complication in resampling is the renormalization of $\tilde{\omega}$: when treated units are dropped or duplicated, the weights must be renormalized to remain a probability vector.

\paragraph{Bootstrap.} In bootstrap draw $b$, treated unit $j$ appears $n_j^{(b)}$ times, and the bootstrap-adjusted weights are $\tilde{\omega}_j^{(b)} = \tilde{\omega}_j n_j^{(b)} / \sum_k \tilde{\omega}_k n_k^{(b)}$. Control weights $\hat{\omega}$ are similarly renormalized. The cluster bootstrap of @clarke2023 generalizes this to the case where clusters of units are resampled together; we extend the renormalization to act at both unit and cluster levels.

\paragraph{Jackknife.} The leave-one-out jackknife renormalizes the remaining weights when a treated unit is dropped: $\tilde{\omega}_k^{(-j)} = \tilde{\omega}_k / \sum_{\ell \neq j} \tilde{\omega}_\ell$ for $k \neq j$. We use the *fixed-weight* jackknife, holding $\hat{\omega}^w$ and $\hat{\lambda}^w$ at their full-sample values. Following @arkhangelsky2021, the jackknife is not recommended for the SC variant due to non-differentiability at the boundary of the simplex.

\paragraph{Placebo.} The placebo method assigns $N_1$ control units as pseudo-treated and estimates the treatment effect on them. Since the original $\tilde{\omega}$ do not naturally map to control units, our default uses uniform weights for pseudo-treated units, with software options for size-matched and permuted weighting schemes. The cluster-robust placebo extension is left to future work.

\paragraph{Diagnosing estimand divergence.} Because $\hat\tau^w$ and $\hat\tau^{\text{sdid}}$ estimate different estimands, we recommend that practitioners estimate both and test whether the gap between them exceeds sampling noise. The paired bootstrap does this directly: apply identical unit (or cluster) resamples to both estimators in each replication, compute $\Delta_b = \hat\tau^w_b - \hat\tau_b$, and compare the observed gap to the recentered bootstrap distribution of $\Delta_b$. This is a Hausman-style comparison. A significant gap indicates treatment-effect heterogeneity correlated with the weights, so the estimand choice of Section 2.2 is consequential and should be made (and defended) explicitly; an insignificant gap means the weighting choice is innocuous and the equally-weighted default loses nothing. We view this as the natural ex-ante diagnostic for whether weighted SDID is needed in a given application, and we apply it in Section 3.2.

The supplementary appendix states each estimator in full; Section 4 reports a Monte Carlo evaluation of finite-sample bias and coverage.

## Staggered Adoption

The weighted modification is developed above for a single adoption date, matching our application. It extends to staggered adoption by the same logic that motivates the estimand. Within each adoption cohort $g$, the collapsed form of Section 2.2 is applied with cohort-restricted treated weights $\tilde\omega^{(g)}$ and a donor pool of not-yet-treated units, yielding a cohort estimate $\hat\tau^{(\tilde\omega),g}$; these aggregate to $\hat\tau^{(\tilde\omega),\text{stag}} = \sum_g W_g\,\hat\tau^{(\tilde\omega),g}$ with researcher-chosen cohort weights $W_g = \sum_{j \in g}\tilde\omega_j$. Standard staggered SDID [@clarke2023] is recovered at $\tilde\omega_j = 1/N_1$ with treated-period cohort weights, and a single cohort returns the estimator above. The identification burden remains estimand-specific and is now also cohort-specific: the oracle-weights assumption of Section 2.3 must hold for each cohort's weighted composite, and the effective-sample condition binds cohort by cohort, so a cohort with few treated units---or one dominant unit---governs the precision of the aggregate. We state the estimand, its reduction properties, and inference formally in the supplementary appendix; the software implements it as \texttt{synthdid\_estimate\_staggered()}. We do not pursue a staggered application in this paper.


# Application: The ACA Medicaid Expansion

We apply the weighted SDID estimator to the county-level mortality effects of the ACA's 2014 Medicaid expansion, building on the design of @borgschulte2020. The ACA application is well-suited to demonstrating the contribution of the weighted estimator: counties differ in population by three orders of magnitude, treatment was assigned at the state level (so within-state correlation matters for inference), and there is plausible heterogeneity in mortality response across the urban--rural gradient given documented differences in baseline uninsurance rates [@sommers2016; @miller2021].

## Setting and Data

```{r load-results}
## All heavy computation is precomputed by scripts/analyze_application.R and
## committed to results/ (see results/_manifest.csv for the seed, git SHA, and
## replication counts). This build only reads and formats those tables, so it
## renders in seconds. The objects reconstructed below carry the names the rest
## of the document already references.
rd <- function(f) readr::read_csv(file.path("..", "results", f), show_col_types = FALSE)
.est     <- rd("app_estimates.csv")
.sc      <- rd("app_scalars.csv"); sc <- setNames(.sc$value, .sc$name)
ev       <- rd("event_studies.csv")
het      <- rd("heterogeneity.csv")
rob      <- rd("robustness.csv")
headline <- rd("headline_comparison.csv")
pintime  <- rd("placebo_intime.csv")
pdist    <- rd("placebo_distribution.csv")

## Descriptive scalars
n_counties <- as.integer(sc[["n_counties"]]); n_treated <- as.integer(sc[["n_treated"]])
n_control  <- as.integer(sc[["n_control"]])
year_range <- c(sc[["year_min"]], sc[["year_max"]]); n_years <- as.integer(sc[["n_years"]])
T0 <- as.integer(sc[["T0"]]); T1 <- as.integer(sc[["T1"]])
treated_pop <- list(total = sc[["treated_pop_total"]], median = sc[["treated_pop_median"]],
                    p10 = sc[["treated_pop_p10"]], p90 = sc[["treated_pop_p90"]])
hhi <- sc[["hhi"]]; n1_eff <- sc[["n1_eff"]]; N1 <- as.integer(sc[["N1"]])
top1_share <- sc[["top1_share"]]; top10_share <- sc[["top10_share"]]
n_states <- as.integer(sc[["n_states"]])
## Treated-jurisdiction count and California's share of treated population
## mass, computed from the heterogeneity file (one row per treated county)
.het_st <- floor(as.numeric(het$fips) / 1000)
n_states_treated <- length(unique(.het_st))
ca_share_treated <- sum(het$pop[.het_st == 6]) / sum(het$pop)
obs_diff <- sc[["obs_diff"]]; p_diverge <- sc[["p_diverge"]]; p_placebo <- sc[["p_placebo"]]
p_diverge_text <- ifelse(p_diverge < 0.001, "$p < 0.001$", paste0("$p = ", round(p_diverge, 3), "$"))

## Estimates and SEs as named vectors. Point estimates are deterministic; the
## state-clustered (se_cluster) and county-level (se_unit) bootstrap SEs are
## from the committed run. tau_sdid is the *standard* synthdid_estimate (the
## robustness "equal" row), which the heterogeneity decomposition averages to
## by construction; the Table-2 "equally weighted" cell (sdid_eq) is the
## weighted-uniform variant and differs by ~2e-5 (Proposition 0).
est_val    <- setNames(.est$estimate,   .est$spec)
se_unit    <- setNames(.est$se_unit,    .est$spec)
se_cluster <- setNames(.est$se_cluster, .est$spec)
estimates_t1 <- as.list(est_val)          # cell()/as.numeric() work on plain numbers
tau_did <- est_val[["did_eq"]]
tau_sdid <- rob$estimate[rob$spec == "equal"]
tau_did_w <- est_val[["did_w"]]; tau_sdid_w <- est_val[["sdid_w"]]
tau_did_eq_x <- est_val[["did_eq_x"]]; tau_sdid_eq_x <- est_val[["sdid_eq_x"]]
tau_did_w_x  <- est_val[["did_w_x"]];  tau_sdid_w_x  <- est_val[["sdid_w_x"]]
se_sdid <- se_cluster[["sdid_eq"]]; se_did <- se_cluster[["did_eq"]]
se_sdid_w <- se_cluster[["sdid_w"]]; se_did_w <- se_cluster[["did_w"]]
```

The sample is a balanced county-year panel covering `r year_range[1]`--`r year_range[2]` with `r format(n_counties, big.mark = ",")` counties (`r format(n_treated, big.mark = ",")` treated, `r format(n_control, big.mark = ",")` control). The outcome is the all-cause mortality rate per 100,000 among adults aged 20--64, drawn from CDC WONDER's Underlying Cause of Death file [@cdcwonder]. We use these public-use tabulations rather than the restricted-access National Vital Statistics System microdata used by @borgschulte2020; both derive from the same NCHS underlying records. The treated counties are those in the `r n_states_treated` jurisdictions (`r n_states_treated - 1` states and the District of Columbia) that expanded Medicaid coverage to childless adults effective January 1, 2014 under the ACA. Following @borgschulte2020, the six states that expanded *after* January 2014 (Alaska, Indiana, Louisiana, Montana, New Hampshire, and Pennsylvania) are excluded from the panel entirely: over the sample period they are neither cleanly treated nor valid never-treated controls. The remaining `r n_states` jurisdictions (44 states plus the District of Columbia) constitute the clusters for state-level inference. Treated-unit weights $\tilde{\omega}_j$ are each county's share of total treated-county population as of 2013. The population distribution is highly skewed: the 90th percentile county (`r format(round(treated_pop$p90), big.mark = ",")` residents) is an order of magnitude larger than the 10th percentile (`r format(round(treated_pop$p10), big.mark = ",")`).

The skewness of the weight distribution is consequential for inference. The Herfindahl index of the population weights is $H = `r round(hhi, 4)`$, implying an effective number of treated units $N_1^{\text{eff}} = `r round(n1_eff, 0)`$ relative to the raw $N_1 = `r N1`$. A single county (Los Angeles County) accounts for roughly `r round(top1_share * 100, 1)`\% of total population-weighted treated mass; the ten largest counties account for `r round(top10_share * 100, 0)`\%. This concentration leaves the effective-sample-size condition (W2) of Proposition 1 with little practical bite, so we rely on cluster-robust bootstrap inference at the state level, where treatment is assigned. The no-dominant-cluster condition of Proposition 3 is more plausible across the `r n_states_treated` treated state clusters, though not comfortably satisfied, California alone accounts for roughly `r round(ca_share_treated * 100)`\% of treated population mass (Supplementary Appendix B.5), which is why we complement the cluster-bootstrap intervals with randomization inference via the in-space placebo (Section 3.6).

\paragraph{Spillovers and SUTVA.} Identification treats non-expansion counties as unaffected by neighboring states' expansions. Cross-border spillovers could violate this: newly insured patients can use providers across state lines, hospital systems near borders serve mixed patient pools, and expansion may have induced enrollment among already-eligible individuals elsewhere [@frean2017; @sommers2016]. Two considerations limit the threat to our conclusions. First, the plausible direction of health-improving spillovers onto control counties *attenuates* the estimated treated--control contrast; spillovers of this kind cannot manufacture the negative population-weighted estimate, which would if anything be larger in their absence. Second, our central result (the divergence between the equally-weighted and population-weighted estimands) is a comparison within the treated group: both estimands rest on the same control pool and the same estimator class, so contamination of control counties moves the two estimates together rather than apart. Spillovers *within* treated states (say, from urban treatment intensity to neighboring rural counties) would color the interpretation of the size gradient in Figure 2, but not the existence of the divergence.

## Main Results

```{r main-results-table}
stars <- function(est, se) {
  if (is.na(est) || is.na(se)) return("")
  p <- 2 * pnorm(-abs(est / se))
  dplyr::case_when(p < 0.01 ~ "***", p < 0.05 ~ "**", p < 0.10 ~ "*", TRUE ~ "")
}
fmt <- function(x) formatC(x, digits = 2, format = "f")

cell <- function(key) {
  est <- as.numeric(estimates_t1[[key]])
  paste0(fmt(est), stars(est, se_cluster[[key]]))
}
row_keys_eq <- c("did_eq", "did_eq_x", "sdid_eq", "sdid_eq_x")
row_keys_w  <- c("did_w",  "did_w_x",  "sdid_w",  "sdid_w_x")

results <- tibble::tibble(
  Estimator    = c("DID", "DID", "SDID", "SDID"),
  Covariates   = c("No", "Yes", "No", "Yes"),
  eq_est       = sapply(row_keys_eq, cell),
  eq_se_boot   = fmt(se_unit[row_keys_eq]),
  eq_se_clust  = fmt(se_cluster[row_keys_eq]),
  w_est        = sapply(row_keys_w, cell),
  w_se_boot    = fmt(se_unit[row_keys_w]),
  w_se_clust   = fmt(se_cluster[row_keys_w])
)

kableExtra::kbl(
  results, digits = NULL, booktabs = TRUE,
  caption = "Effect of the 2014 Medicaid expansion on county-level mortality.",
  col.names = c("Estimator", "Covariates", "Estimate", "SE (boot)", "SE (cluster)",
                "Estimate", "SE (boot)", "SE (cluster)"),
  align = c("l", "c", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 2, "Equally weighted" = 3, "Population weighted" = 3)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kableExtra::footnote(
    general = sprintf(paste(
      "Outcome: all-cause mortality per 100,000, ages 20--64 (CDC WONDER). Sample: %s counties, %s--%s.",
      "Each cell reports the same estimator under two standard errors: SE (boot) is the county-level bootstrap;",
      "SE (cluster) is the state-clustered bootstrap (%s states), both with 500 replications.",
      "Covariate rows adjust for the six time-varying county controls of Borgschulte and Vogler (2020):",
      "percent white, percent aged 55--64, log population aged 20--64, log population aged 35--44,",
      "log female population aged 20--64, and the unemployment rate; covariates for three FIPS-change",
      "counties (15005, 46113, 51019) are imputed with state-year means so all rows share the same sample.",
      "Significance stars use the state-clustered bootstrap SE: * p$<$0.10, ** p$<$0.05, *** p$<$0.01.",
      "Synthetic control results are reported and diagnosed in Supplementary Appendix E.1."),
      format(n_counties, big.mark = ","), year_range[1], year_range[2], n_states),
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```

Table 2 reports the main results. Across both estimators, with and without covariate adjustment, the equally-weighted and population-weighted estimates diverge sharply: the equally-weighted estimates are statistically indistinguishable from zero while their population-weighted counterparts are large and negative. The equally-weighted SDID estimate is `r fmt(as.numeric(tau_sdid))` deaths per 100,000 (state-clustered bootstrap SE `r fmt(se_sdid)`); the population-weighted SDID estimate is `r fmt(as.numeric(tau_sdid_w))` (SE `r fmt(se_sdid_w)`), significant at the 1\% level. The DID estimates show the same pattern. Adjusting for the six time-varying county covariates used by @borgschulte2020 leaves both conclusions essentially unchanged (`r fmt(as.numeric(tau_sdid_eq_x))` equally weighted, `r fmt(as.numeric(tau_sdid_w_x))` population weighted for SDID), confirming that the divergence is a property of the weighting choice rather than of omitted time-varying confounders. For every specification we report both the county-level bootstrap and the state-clustered bootstrap standard error; clustering at the state level (the level at which treatment is assigned) increases the standard errors but does not change any qualitative conclusion. We exclude synthetic control from Table 2: its estimates are an extreme outlier relative to the DID--SDID family in this application, and we report them, together with the standard battery of post-estimation diagnostics that explains their magnitude, in Supplementary Appendix E.1.

To confirm that the divergence between the equally-weighted and population-weighted SDID estimates is not attributable to sampling variability, we apply the paired-bootstrap divergence diagnostic recommended in Section 2.4. On each of $B = 500$ bootstrap replications we apply identical unit resamples to both estimators and compute $\Delta_b = \hat{\tau}^w_b - \hat{\tau}_b$. Recentering the bootstrap distribution to the null of no difference, the two-sided $p$-value for the observed gap of $|\Delta_{\text{obs}}| = `r abs(round(as.numeric(tau_sdid_w) - as.numeric(tau_sdid), 1))`$ deaths per 100,000 is `r p_diverge_text`. The two estimands are sharply distinguishable---in the recommended workflow, exactly the situation in which the researcher must take an explicit stand on the target estimand before reading off a policy conclusion.

```{r event-study-both, fig.cap="Event-study decomposition under equal and population weighting.\\newline \\footnotesize\\textit{Notes:} DID (left) and SDID (right). Points are year-specific effects under equal (blue circles) and population (orange triangles) weighting of treated counties; shaded bands are 95\\% state-clustered bootstrap confidence intervals (500 replications). The vertical dashed line marks treatment onset (2014).", fig.height=4, fig.width=8, cache=TRUE}
es_df <- ev %>% mutate(Estimator = factor(Estimator, levels = c("DID", "SDID")))

rmspe_did_pre  <- es_df %>% filter(Estimator == "DID",  Weighting == "Equally weighted", year < 2014) %>% summarise(rmspe = sqrt(mean(estimate^2))) %>% pull(rmspe)
rmspe_sdid_pre <- es_df %>% filter(Estimator == "SDID", Weighting == "Equally weighted", year < 2014) %>% summarise(rmspe = sqrt(mean(estimate^2))) %>% pull(rmspe)

ggplot(es_df, aes(x = year, y = estimate, color = Weighting, fill = Weighting)) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.15, color = NA) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 2013.5, linetype = "dashed", color = "gray30") +
  geom_line(aes(linetype = Weighting), linewidth = 0.5) +
  geom_point(aes(shape = Weighting), size = 1.3) +
  facet_wrap(~ Estimator) +
  scale_color_manual(values = w_pal) +
  scale_fill_manual(values  = w_pal) +
  scale_shape_manual(values = w_shp) +
  scale_linetype_manual(values = w_lty) +
  labs(x = NULL, y = "Effect on mortality\n(deaths per 100,000)",
       color = NULL, fill = NULL, shape = NULL, linetype = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom")
```

Figure 1 displays the event-study decomposition. Under equal weighting (blue circles), pre-treatment coefficients cluster near zero for both estimators. Under population weighting (orange triangles), the DID panel shows non-trivial pre-treatment divergence between treated and control groups, raising parallel-trends concerns when larger counties are weighted more heavily---precisely the concern that motivated @borgschulte2020's DLPS approach. The SDID estimator addresses the pre-treatment fit problem through a different route: its data-driven time weights $\hat{\lambda}^w$ down-weight periods in which treated and synthetic-control trajectories diverge, yielding tighter pre-treatment coefficients in the right panel (RMSPE: `r round(rmspe_sdid_pre, 2)` for SDID vs.\ `r round(rmspe_did_pre, 2)` for DID, both equally weighted). Post-treatment, both estimators show the same qualitative pattern: under population weighting the coefficients are consistently negative ($-14$ to $-24$ deaths per 100,000); under equal weighting they fluctuate near zero. The consistency across estimators isolates the *weighting choice*, not the *estimator choice*, as the source of the divergence.

## Mechanism: Heterogeneity by County Size

```{r heterogeneity-plot, fig.cap="Treatment-effect heterogeneity by county population.\\newline \\footnotesize\\textit{Notes:} Binned scatter of unit-level SDID treatment effects against log county population (2013), in 20 quantile bins of log population. Point size is proportional to the total 2013 population of counties in the bin (so larger-population deciles render as larger points even though each bin holds the same number of counties). Unit-level effects use the equally-weighted SDID control weights $\\hat{\\omega}$ and time weights $\\hat{\\lambda}$ (the post-period gap relative to the synthetic control minus the $\\hat{\\lambda}$-weighted pre-period gap), isolating heterogeneity in the treatment effect from heterogeneity in the weighting scheme; the equally-weighted average of the unit-level effects recovers the SDID estimate exactly. The dashed line is a population-weighted OLS fit through the bin means.", fig.height=3.8}
## Per-county SDID unit-level effects are precomputed in
## results/heterogeneity.csv: the post-period gap relative to the synthetic
## control minus the lambda-weighted pre-period gap, using the equally-weighted
## control/time weights. By construction mean(effect) equals the SDID estimate.
unit_effects <- het$effect
stopifnot(abs(mean(unit_effects) - as.numeric(tau_sdid)) < 1e-6)

n_bins <- 20

unit_effect_df <- tibble::tibble(
  effect = het$effect, pop = het$pop
) %>%
  mutate(log_pop = log(pop),
         pop_bin = dplyr::ntile(log_pop, n_bins)) %>%
  group_by(pop_bin) %>%
  summarise(mean_effect  = mean(effect, na.rm = TRUE),
            mean_log_pop = mean(log_pop, na.rm = TRUE),
            bin_pop      = sum(pop, na.rm = TRUE),
            n            = dplyr::n(),
            .groups      = "drop")

bin_cor         <- cov.wt(unit_effect_df[, c("mean_log_pop", "mean_effect")],
                          wt = unit_effect_df$bin_pop, cor = TRUE)$cor[1, 2]
ind_cor         <- cor(het$effect, log(het$pop))
smallest_effect <- unit_effect_df$mean_effect[1]
largest_effect  <- unit_effect_df$mean_effect[nrow(unit_effect_df)]
smallest_logpop <- unit_effect_df$mean_log_pop[1]
largest_logpop  <- unit_effect_df$mean_log_pop[nrow(unit_effect_df)]
## Population-weighted mean of the unit-level effects: approximately recovers
## the weighted SDID estimate (omega-hat/lambda-hat are held at their
## equally-weighted values, so the recovery is not mechanical)
popw_unit_mean  <- sum(het$effect * het$pop) / sum(het$pop)

ggplot(unit_effect_df, aes(x = mean_log_pop, y = mean_effect)) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_point(aes(size = bin_pop), color = "steelblue", alpha = 0.85) +
  geom_smooth(aes(weight = bin_pop), method = "lm", se = TRUE,
              color = "gray30", linewidth = 0.6, linetype = "dashed") +
  scale_size_area(name = "Total population\nin bin (millions)",
                  max_size = 9,
                  labels = function(x) round(x / 1e6, 1)) +
  labs(x = "Log county population (2013)",
       y = "Avg. post-treatment effect\n(deaths per 100,000)") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "right")
```

The divergence between estimands is not a statistical artifact: it reflects substantive treatment-effect heterogeneity that is correlated with county size. Figure 2 plots mean SDID unit-level effects against log county population in 20 quantile bins of log population, with point area scaled to the total 2013 population in each bin so the visual emphasis matches the population-weighted estimand. Unit-level effects are computed with the *equally-weighted* SDID weights (the post-period gap between each treated county and the synthetic control $\hat{\omega}^\top Y_{\text{control},t}$, net of the $\hat{\lambda}$-weighted pre-period gap), so the heterogeneity displayed is in the treatment effect itself, not in the weighting scheme, and the equally-weighted average of the unit-level effects reproduces the SDID estimate by construction. Across the bottom three population quartiles, mean effects oscillate around zero (between $`r round(min(unit_effect_df$mean_effect[1:15]), 0)`$ and $+`r round(max(unit_effect_df$mean_effect[1:15]), 0)`$ deaths per 100,000) with no systematic pattern. The mortality response is concentrated in the top population bins: mean effects decline monotonically over the four largest-population bins, reaching $`r round(largest_effect, 0)`$ deaths per 100,000 in the largest bin (log population $\approx$ `r round(largest_logpop, 1)`). The population-weighted bin-level correlation between mean log population and mean treatment effect is `r round(bin_cor, 2)`; the unweighted county-level correlation is more modest (`r round(ind_cor, 2)`), reflecting noise in county-level synthetic control predictions for small counties where mortality rates are volatile. As an internal-consistency check, the *population-weighted* average of these same unit-level effects is `r round(popw_unit_mean, 1)` deaths per 100,000---nearly identical to the weighted SDID estimate of `r round(as.numeric(tau_sdid_w), 1)`, even though the unit-level effects hold $\hat{\omega}$ and $\hat{\lambda}$ at their equally-weighted values. The divergence between the two estimands is driven by *which counties* the estimand averages over, not by the re-optimization of the synthetic control.

```{r urban-binscatter, dependson="heterogeneity-plot", fig.cap="Treatment-effect heterogeneity by urban-population share.\\newline \\footnotesize\\textit{Notes:} Binned scatter of unit-level SDID treatment effects against the urban-population share of each county, in 20 quantile bins of pct.\\ urban (2010 Decennial Census, P002002 / P002001). Point area is proportional to the total 2013 population of counties in the bin; the dashed line is a population-weighted OLS fit through the bin means. Urban share is fixed at its 2010 (pre-treatment) value, so the gradient reflects pre-existing county composition rather than any post-2014 reclassification.", fig.height=3.4}
urban_share <- readr::read_csv("data/urban_share.csv", show_col_types = FALSE)

unit_effect_urban <- het %>%
  dplyr::transmute(fips = as.numeric(fips), effect, pop) %>%
  dplyr::left_join(urban_share, by = "fips") %>%
  dplyr::filter(!is.na(pct_urban))

n_urban_dropped <- nrow(het) - nrow(unit_effect_urban)

urban_bin_df <- unit_effect_urban %>%
  ## ntile() rather than cut(quantile(...)): the suppression-recovered small
  ## counties make pct_urban = 0 a mass point, so raw quantile breaks repeat
  mutate(pop_bin = dplyr::ntile(pct_urban, n_bins)) %>%
  group_by(pop_bin) %>%
  summarise(mean_effect = mean(effect, na.rm = TRUE),
            mean_urban  = mean(pct_urban, na.rm = TRUE),
            bin_pop     = sum(pop, na.rm = TRUE),
            n           = dplyr::n(),
            .groups     = "drop")

urban_bin_cor <- cov.wt(urban_bin_df[, c("mean_urban", "mean_effect")],
                        wt = urban_bin_df$bin_pop, cor = TRUE)$cor[1, 2]
urban_ind_cor <- cor(unit_effect_urban$pct_urban, unit_effect_urban$effect)
urban_lo_eff  <- urban_bin_df$mean_effect[1]
urban_hi_eff  <- urban_bin_df$mean_effect[nrow(urban_bin_df)]
urban_lo_pct  <- urban_bin_df$mean_urban[1]
urban_hi_pct  <- urban_bin_df$mean_urban[nrow(urban_bin_df)]

ggplot(urban_bin_df, aes(x = mean_urban, y = mean_effect)) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_point(aes(size = bin_pop), color = "darkgreen", alpha = 0.85) +
  geom_smooth(aes(weight = bin_pop), method = "lm", se = TRUE,
              color = "gray30", linewidth = 0.6, linetype = "dashed") +
  scale_size_area(name = "Total population\nin bin (millions)",
                  max_size = 9,
                  labels = function(x) round(x / 1e6, 1)) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Urban-population share, 2010",
       y = "Avg. post-treatment effect\n(deaths per 100,000)") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "right")
```

Figure 3 repeats the exercise using each county's 2010 urban-population share, fixed at its pre-treatment value to avoid contamination from any post-2014 reclassification. The pattern is the same as in Figure 2 but maps onto a more directly interpretable substantive dimension: the most-rural bin of counties (the bottom twentieth by urban share, $\approx$ `r scales::percent(urban_lo_pct, accuracy = 1)` urban) shows mean effects of $`r ifelse(urban_lo_eff >= 0, "+", "")``r round(urban_lo_eff, 0)`$ deaths per 100,000 (a noisy positive point estimate, the urban-share counterpart of the around-zero oscillation among small-county bins in Figure 2, not a precisely measured harm), while the most-urban bin ($\approx$ `r scales::percent(urban_hi_pct, accuracy = 1)` urban) shows reductions of $`r abs(round(urban_hi_eff, 0))`$. The population-weighted bin-level correlation is `r round(urban_bin_cor, 2)` (unweighted county-level: `r round(urban_ind_cor, 2)`). Urban share and log population are mechanically correlated, so this view does not provide an independent test, but it makes the substantive content of the heterogeneity gradient more transparent: the population-weighted estimand is dominated by counties that are both populous *and* highly urbanized---precisely the counties for which @sommers2016 and @miller2021 document larger pre-expansion uninsurance rates and more health-care infrastructure capable of absorbing the coverage shock.

Whatever its substantive interpretation, the gradient mechanically generates the divergence between the equally-weighted and population-weighted estimands: equal weighting averages over the many small counties whose estimated effects oscillate around zero; population weighting shifts the estimand toward the few large counties where effects are large and negative.

## Robustness to Alternative Weight Specifications

```{r robustness-table, cache=TRUE}
robust_results <- tibble::tibble(
  `Weight specification` = rob$label,
  Estimate = rob$estimate,
  SE = rob$se,
  `$N_1^{\\text{eff}}$` = rob$n1_eff
)
tau_sdid_2010_app <- rob$estimate[rob$spec == "pop_2010"]  # referenced in the prose below

kableExtra::kbl(
  robust_results, digits = 2, booktabs = TRUE, escape = FALSE,
  caption = "Robustness to alternative treated-unit weight specifications."
) %>%
  kableExtra::kable_styling(latex_options = "hold_position") %>%
  kableExtra::footnote(
    general = "SDID estimates with state-clustered bootstrap SEs (500 replications) for every row, including equal weighting. $N_1^{\\\\text{eff}}$ is the Kish effective sample size $(\\\\sum_j \\\\tilde\\\\omega_j^2)^{-1}$. The baseline weights are each county's share of total treated-county population in 2013; alternative rows change the base year or compress the weight distribution toward equality.",
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```

Table 3 examines the sensitivity of the population-weighted SDID estimate to alternative weight specifications. Replacing 2013 population with 2010 population shares yields an estimate of similar magnitude (`r round(as.numeric(tau_sdid_2010_app), 1)` deaths per 100,000), confirming that the result does not depend on the choice of base year. As the weight distribution is compressed (square-root population, log population, equal weighting), the estimate moves monotonically toward zero. The effective sample size grows correspondingly, from $N_1^{\text{eff}} \approx `r round(n1_eff, 0)`$ under linear population weights to $N_1 = `r N1`$ under equal weights. The substantive lesson is that the large negative effect requires giving substantial weight to the largest counties: log compression alone is sufficient to eliminate it. This is not an artifact of the specific 2013 linear-population specification, but it does pinpoint *which* aggregation choices produce the result.

## Comparison with the Published Estimate

```{r headline-comparison-table}
headline_results <- tibble::tibble(
  Method = c("Standard SDID (equally weighted)",
             "Weighted SDID (population)",
             "Weighted DID (population)",
             "Borgschulte \\& Vogler (2020), Table 2 col.\\ 2"),
  Estimate = c(as.numeric(tau_sdid), as.numeric(tau_sdid_w),
               as.numeric(tau_did_w), -11.36),
  # BV2020 SE 3.59 from Table 2, col 2 (state-clustered, 45 clusters)
  SE = c(se_sdid, se_sdid_w, se_did_w, 3.59)
)

kableExtra::kbl(
  headline_results, digits = 2, booktabs = TRUE, escape = FALSE,
  caption = "Comparison with the published estimate.",
  col.names = c("Method", "Estimate", "SE")
) %>%
  kableExtra::kable_styling(latex_options = c("hold_position")) %>%
  kableExtra::footnote(
    general = "The first three rows are estimated on our balanced county-year panel with state-clustered bootstrap SEs; the fourth row is the headline DLPS-weighted DID estimate published by Borgschulte and Vogler (2020) (Table 2, column 2; state-clustered SE, 45 clusters). Supplementary Appendix E.5 documents our public-data DLPS reproduction, which recovers the structure of the BV results, including their heterogeneity by baseline uninsurance, on a near-identical county sample.",
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```

Table 4 places our population-weighted SDID estimate alongside the headline estimate from @borgschulte2020, who weight a propensity-score-adjusted DID by working-age county population using a double-lasso variable selection procedure on a broad set of demographic, economic, and political characteristics. Our population-weighted SDID estimate (`r round(as.numeric(tau_sdid_w), 2)` deaths per 100,000) is in the same neighborhood as BV's published $-11.36$, despite differences in identification strategy (data-driven control matching plus pre-treatment time-weight selection vs.\ a propensity score for treatment selection) and in data source (CDC WONDER public tabulations vs.\ restricted NVSS microdata). The corresponding equally-weighted SDID estimate is statistically indistinguishable from zero, which shows that the choice of treated-unit weights, not the choice of estimator, accounts for nearly all of the gap. Supplementary Appendix E.5 reports our public-data reproduction of the BV design itself: recovering CDC WONDER's suppressed small-count cells by allocating exact state-year residuals brings the estimation sample within two counties of BV's, and the reproduction recovers their qualitative results, including the concentration of the mortality response in counties with high baseline uninsurance. Because the BV design is itself a population-weighted, level-matched comparison, it rests on the same parallel-trends assumption that our in-time placebo (Section 3.6) stresses; the proximity of the two estimates is therefore a comparison of like designs sharing a common identifying assumption, not an independent validation of either.

## Placebo Tests

```{r placebo-intime-app}
## In-time placebo estimates/SEs precomputed in results/placebo_intime.csv.
intime_results <- pintime
```

```{r placebo-inspace-app}
## In-space placebo distribution precomputed in results/placebo_distribution.csv;
## p_placebo (randomization-inference p-value) is read in load-results.
placebo_ests <- pdist$est
```

```{r placebo-figure, fig.cap="Placebo tests.\\newline \\footnotesize\\textit{Notes:} Left: in-time placebo with simulated treatment dates of 2011 and 2012, restricting to pre-treatment data (2009--2013); circles are equally-weighted and triangles population-weighted SDID estimates, with whiskers marking 95\\% state-clustered bootstrap confidence intervals (500 replications). Right: in-space placebo distribution from 500 random re-assignments of treatment to control counties, with the dashed vertical line marking the observed population-weighted estimate.", fig.height=3.6, fig.width=8}
intime_long <- bind_rows(
  intime_results %>%
    transmute(placebo_year, weighting = "Equally weighted",
              estimate = estimate_eq, se = se_eq),
  intime_results %>%
    transmute(placebo_year, weighting = "Population weighted",
              estimate = estimate_wt, se = se_wt)
)

pd <- position_dodge(width = 0.35)
p_intime <- ggplot(intime_long,
                   aes(x = placebo_year, y = estimate,
                       color = weighting, shape = weighting)) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_errorbar(aes(ymin = estimate - 1.96 * se, ymax = estimate + 1.96 * se,
                    linetype = weighting),
                width = 0.12, linewidth = 0.5, position = pd) +
  geom_point(size = 3, position = pd) +
  scale_color_manual(values = w_pal) +
  scale_shape_manual(values = w_shp) +
  scale_linetype_manual(values = w_lty) +
  scale_x_continuous(breaks = c(2011, 2012)) +   # integer years, no decimal ticks
  labs(x = "Placebo year", y = "Estimate (deaths per 100,000)",
       color = NULL, shape = NULL, linetype = NULL, title = "In-time") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom")

p_inspace <- ggplot(tibble::tibble(est = placebo_ests[!is.na(placebo_ests)]), aes(x = est)) +
  geom_histogram(bins = 40, fill = "gray70", color = "white") +
  geom_vline(xintercept = as.numeric(tau_sdid_w), color = "black", linewidth = 0.9, linetype = "dashed") +
  annotate("text", x = as.numeric(tau_sdid_w) * 1.05, y = Inf, vjust = 1.5, hjust = 0,
           label = sprintf("Observed\n%.1f", as.numeric(tau_sdid_w)), color = "black", size = 3) +
  labs(x = "Placebo estimate (deaths per 100,000)", y = "Count", title = "In-space") +
  theme_minimal(base_size = 10)

if (requireNamespace("patchwork", quietly = TRUE)) {
  patchwork::wrap_plots(p_intime, p_inspace, ncol = 2)
} else {
  print(p_intime); print(p_inspace)
}
```

We conduct two complementary falsification exercises. The *in-time* placebo restricts the sample to pre-treatment years (2009--2013) and assigns simulated treatment dates of 2011 and 2012. Absent a policy change, the estimator should return zero. Figure 4 (left) shows that the *equally-weighted* placebo estimates pass this test: both are statistically indistinguishable from zero. The *population-weighted* estimates do not. They are `r round(intime_results$estimate_wt[1], 1)` (2011) and `r round(intime_results$estimate_wt[2], 1)` (2012) deaths per 100,000, with 95\% state-clustered bootstrap intervals that exclude zero. A significant placebo under the estimand we favor deserves to be taken seriously as evidence of bias, not explained away, so we work through what it implies for the headline estimate.

The mechanism is visible in the population-weighted event study (Figure 1): the largest treated counties were already declining relative to their synthetic control over the pre-period, with the pre-treatment gap moving from positive in 2009 to significantly negative by 2013. SDID has two tools against such a differential trend, and both are limited here. The unit weights $\hat\omega^w$ can select control counties on similar trajectories, but only within the convex hull of control trajectories. The time weights $\hat\lambda^w$ do help---the pre-treatment divergence is markedly smaller under SDID than under DID in Figure 1---but they live on the simplex: they can down-weight divergent pre-periods, not extrapolate beyond the pre-period range. A differential trend of $\delta$ per year that no convex combination of controls matches therefore survives the double difference in proportion to the time gap between the post-period average and the $\hat\lambda^w$-weighted pre-period center; that gap is mechanically positive and, with $T_0 = 5$ and a post-period running through 2017, is at least two and a half years. In the factor-model language of Section 2.3, this is a violation of the oracle-weights assumption (A5) that is specific to the *population-weighted* target: the trend component loads on the largest counties, while the equally-weighted target---in which those counties are `r N1` among equals---is essentially unexposed to it, which is why the equally-weighted placebos pass.

```{r omega-size-table}
osb <- rd("omega_size_bins.csv")
osb <- osb[match(c("target_popw", "omega_popw", "omega_eq"), osb$series), ]
shr <- function(x) sprintf("%.1f", 100 * x)

omega_tbl <- tibble::tibble(
  Series = c("Treated target $\\tilde\\omega$ (population weights)",
             "Control weights $\\hat\\omega^w$ (population-weighted target)",
             "Control weights $\\hat\\omega$ (equally-weighted target)"),
  b1 = shr(osb$b1), b2 = shr(osb$b2), b3 = shr(osb$b3),
  b4 = shr(osb$b4), b5 = shr(osb$b5),
  avg = format(round(osb$wavg_pop), big.mark = ",")
)

kableExtra::kbl(
  omega_tbl, booktabs = TRUE, escape = FALSE,
  caption = "Size composition of the estimated control weights and the treated target.",
  col.names = c("", "1 (smallest)", "2", "3", "4", "5 (largest)", "Avg.\\ county pop."),
  align = c("l", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 1, "Share of weight mass by county-size bin (\\\\%)" = 5, " " = 1),
                               escape = FALSE) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kableExtra::footnote(
    general = sprintf(paste(
      "Bins are pooled quantiles of 2013 county population across all %s counties",
      "(breaks at the 25th/50th/75th/90th percentiles: %s / %s / %s / %s residents;",
      "bin 5 is the top decile). The final column is the weight-averaged 2013 county",
      "population of each series. Control-weight rows are the Frank-Wolfe solutions",
      "$\\\\hat\\\\omega$ from the two SDID estimates of Table 2."),
      format(n_counties, big.mark = ","),
      format(round(osb$brk1[1]), big.mark = ","), format(round(osb$brk2[1]), big.mark = ","),
      format(round(osb$brk3[1]), big.mark = ","), format(round(osb$brk4[1]), big.mark = ",")),
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```

Table 5 locates the failure inside the estimated weights. The population-weighted treated target places `r shr(osb$b5[1])`\% of its mass in the top size decile---the weighted-average treated resident lives in a county of roughly `r format(round(osb$wavg_pop[1], -3), big.mark = ",")` people---while the estimated control weights place `r shr(osb$b5[2])`\% there, indistinguishable from the `r sprintf("%.1f", 100 * sum(osb$b5[3]))`\% under the equally-weighted target and from what a uniform spread over the control pool would produce. More telling still, the size composition of $\hat\omega^w$ is essentially *invariant to the target*: re-solving the weight program against the population-weighted trajectory moves every bin share by less than a tenth of a percentage point. The reason is that level-plus-intercept matching over a five-year pre-period is size-blind. With the intercept absorbing the level gap, a ridge-diffused combination of small counties fits the pre-period *shape* of the big-county trajectory as well as large controls would, and the regularization prefers the diffuse solution. Because secular mortality trends vary systematically with county size, the size-blind donor portfolio imports the wrong trend into the post-period---the mechanism behind the placebo failure above, and the reason the weighted DID fails identically: what is hard-coded in the DID (a population-weighted treated average against a size-unrepresentative control average) emerges in SDID from the estimation itself.

What does the failing placebo imply for the headline? We read it as a statement about the ACA county design, not about the weighting. The population-weighted target loads on a secular mortality trend that varies systematically with county size, and no convex, level-matched estimator on this panel fully removes it: SDID's time weights live on the simplex and cannot extrapolate a differential trend beyond the pre-period range, so the same in-time falsification fails for the parallel weighted DID, and any level-matched design that targets the population-weighted composite---including the propensity-score benchmark of @borgschulte2020---rests on the parallel-trends assumption this placebo stresses. The equally-weighted target, in which the largest counties are `r N1` among equals, is essentially unexposed and its placebos pass, which is why the divergence between estimands survives even though the population-weighted level does not. We therefore do not report a trend-corrected population-weighted point estimate for this design. Five pre-periods cannot separate a size-correlated differential trend from genuine early-coverage gains in the large jurisdictions that dominate the estimand---several expanded coverage to low-income adults before 2014 under ACA-related waivers, most prominently California's county-run Low Income Health Program (2011--2013) [@frean2017]---and any linear trend correction would substitute one strong assumption for another. Credibly identifying a population-weighted mortality effect in this setting instead calls for designs built for it: the triple-difference and administrative-data approaches of @miller2021 and @wyse2025, whose population-weighted, state-identified estimands survive the trend checks this county design cannot.

The *in-space* placebo randomly re-assigns treatment to `r N1` control counties (matching the true number of treated counties) while keeping the real treatment date of 2014 and using equal pseudo-treated weights. Figure 4 (right) shows the resulting distribution of placebo estimates from 500 random draws; the observed population-weighted estimate (`r round(as.numeric(tau_sdid_w), 1)`) sits well in the left tail: `r if (p_placebo == 0) "no draw is as extreme, a randomization-inference $p$-value below $1/501 \\approx 0.002$" else paste0("the randomization-inference $p$-value is $", round(p_placebo, 3), "$")`. Given the in-time results, two limitations of this test are worth stating. The placebo draws are equally weighted, so their dispersion likely understates the sampling variability of a population-weighted statistic whose effective sample size is an order of magnitude smaller; and because pseudo-treated units are drawn from the control pool, the test has no power against the treated-specific differential trend diagnosed above. The in-space placebo is therefore best read as ruling out that an effect of this size arises from noise in the control pool, not as absolving the population-weighted estimate of trend bias.


# Monte Carlo Evidence

We evaluate the finite-sample performance of the weighted SDID estimator and its variance estimators via simulation. Two questions are central. First, is the weighted estimator unbiased for its target $\tau^{(\tilde{\omega})}$ under heterogeneous treatment effects? Second, do the bootstrap, jackknife, and placebo variance estimators achieve nominal coverage when treatment effects are correlated with unit size?

## Design

The data-generating process follows the factor model of @arkhangelsky2021. For $N = N_0 + N_1$ units observed over $T = T_0 + T_1$ periods,
$$Y_{it} = L_{it} + \tau_{it} W_{it} + \sigma \epsilon_{it},$$
where $L = UV^\top + \alpha\mathbf{1}^\top + \mathbf{1}\beta^\top$ is rank-2 with additive fixed effects, $U \in \mathbb{R}^{N \times 2}$ and $V \in \mathbb{R}^{T \times 2}$ are factor loadings drawn from Poisson distributions, $\alpha_i = 10 i / N$, $\beta_t = 10 t / T$, and $\epsilon_{it}$ is drawn from an AR(1) process with $\rho = 0.7$. Each treated unit $j$ is assigned a "size" $s_j$ from a log-normal distribution scaled to achieve a target Herfindahl index for the population weights $\tilde{\omega}_j = s_j / \sum_k s_k$, and the treatment effect is
$$\tau_{ij} = \bar\tau + \gamma\left(\frac{s_j}{\bar s} - 1\right),$$
with $\bar\tau = 1$, $\bar s = N_1^{-1}\sum_j s_j$, and $\gamma \geq 0$ controlling the size-effect correlation. We fix $N_0 = 100$, $T_0 = 40$, $T_1 = 10$, $\sigma = 0.5$, and vary $N_1 \in \{5, 10, 20\}$, $\gamma \in \{0, 0.5, 1.0\}$ (homogeneous, moderate, strong heterogeneity), and the weight concentration (low, HHI $= 1.5/N_1$; high, HHI $= 3/N_1$), yielding 18 grid configurations evaluated over 100 simulations each with bootstrap and placebo variance estimators using 50 replications. A nineteenth, ACA-calibrated configuration gives the application's standard errors direct simulation backing: it reproduces the application's panel dimensions ($N_0 = 1{,}643$, $N_1 = 1{,}181$, $T_0 = 5$, $T_1 = 4$) and sets the treated-unit sizes to the application's actual 2013 county populations, so every simulation has exactly the application's weight concentration (HHI $= 0.0098$, $N_1^{\text{eff}} \approx 102$), with $\gamma = 1$, the strong size-effect-correlation regime the application's county-size gradient suggests. The simulation sweep is parallelized across configurations and simulation draws (`scripts/run_mc_simulations.R`), with seeding that reproduces the equivalent sequential loop exactly; implementation details appear in Supplementary Appendix D.

To connect the Monte Carlo to the identification concern raised by the application, we add a *trend cell* to the design. Holding the treatment effect fixed, we add a size-correlated linear trend to the factor structure of the treated units, $L_{it} \to L_{it} + \gamma_{\text{trend}}\,(s_j/\bar s - 1)\,(t/T)$, so that the largest units---those that dominate the population weights---drift relative to the controls over the panel (normalized time $t/T$ keeps the trend's accumulation on the same scale as the treatment effect). This violates the oracle-weights assumption for the population-weighted target while leaving the equally-weighted target essentially unexposed, exactly the asymmetry diagnosed in Section 3.6. The estimand itself is unchanged: the trend perturbs $L$, not $\tau$. We vary $\gamma_{\text{trend}} \in \{0, 0.5, 1\}$ at $N_1 = 20$, $\gamma = 0.5$, and moderate weight concentration, with 200 simulations per cell (`scripts/run_mc_trend.R`).

```{r mc-run, cache=FALSE}
## The sweep is computed outside knitr by scripts/run_mc_simulations.R
## (parallelized; identical seeding to a sequential loop) and written to
## data/mc_results.csv.
mc_available <- file.exists(file.path("data", "mc_results.csv"))
mc_df <- if (mc_available) {
  readr::read_csv(file.path("data", "mc_results.csv"), show_col_types = FALSE)
} else NULL
## The ACA-calibrated cell (config 19, N1 = 1181) is computed by the same
## script; tables and prose accommodate builds where it has not yet been run.
mc_has_aca <- mc_available && any(mc_df$N1 > 100)
```

## Results

`r if (!mc_available) "*Monte Carlo results are generated by* \\texttt{scripts/run\\_mc\\_simulations.R}; *the results file is not present in this build, so the Monte Carlo tables are omitted.*"`

Tables \ref{tab:mc-bias} and \ref{tab:mc-coverage} report bias, RMSE, and 95\% confidence-interval coverage across the simulation configurations. Three patterns are central to the inference recommendations of Section 2. First, both the equally-weighted and population-weighted SDID estimators are approximately unbiased for their respective targets, with RMSE differences favoring the equally-weighted variant for the obvious reason that it has a larger effective sample. Second, the bootstrap and placebo variance estimators achieve approximately nominal 95\% coverage, with undercoverage at high HHI for the placebo, which uses uniform pseudo-treated weights, as discussed in Section 2.4. Third, the jackknife is near-nominal with few treated units and grows conservative as the number of treated units increases, over-covering at $N_1 = 20$. Together these patterns validate Propositions 1--3 in finite samples within the regime where Assumptions W2--W3 are quantitatively reasonable.

```{r mc-bias-table, eval=mc_available}
bias_tbl <- mc_df %>%
  group_by(N1, gamma, hhi) %>%
  summarise(
    Bias_unwt = mean(est_unwt - true_att_equal, na.rm = TRUE),
    RMSE_unwt = sqrt(mean((est_unwt - true_att_equal)^2, na.rm = TRUE)),
    Bias_wt   = mean(est_wt - true_att_weighted, na.rm = TRUE),
    RMSE_wt   = sqrt(mean((est_wt - true_att_weighted)^2, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric) & !c(N1, gamma, hhi), ~ round(., 3)),
         hhi = round(hhi, 4))

kableExtra::kbl(
  bias_tbl, booktabs = TRUE,
  caption = "Monte Carlo bias and RMSE of the SDID estimators.", label = "mc-bias",
  col.names = c("$N_1$", "$\\gamma$", "HHI", "Bias", "RMSE", "Bias", "RMSE"),
  escape = FALSE,
  align = c("r", "r", "r", "r", "r", "r", "r")
) %>%
  kableExtra::add_header_above(c(" " = 3, "Equally weighted" = 2, "Population weighted" = 2)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position")) %>%
  kableExtra::footnote(
    general = "Bias and RMSE are computed against each estimator's own target ATT (the equally-weighted and population-weighted averages of the unit-level effects, respectively). 100 simulations per configuration. $N_1$ is the number of treated units, $\\\\gamma$ the size-effect correlation, and HHI the Herfindahl index of the treated-unit weights. Data-generating process described in Section 4.1.",
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```

```{r mc-coverage-table, eval=mc_available}
coverage <- function(est, se, true_val) {
  ci_lo <- est - 1.96 * se; ci_hi <- est + 1.96 * se
  mean(ci_lo <= true_val & true_val <= ci_hi, na.rm = TRUE)
}
cov_tbl <- mc_df %>%
  group_by(N1, gamma, hhi) %>%
  summarise(
    Boot_unwt = coverage(est_unwt, se_boot_unwt, true_att_equal),
    JK_unwt   = coverage(est_unwt, se_jk_unwt,   true_att_equal),
    Plac_unwt = coverage(est_unwt, se_plac_unwt, true_att_equal),
    Boot_wt   = coverage(est_wt,   se_boot_wt,   true_att_weighted),
    JK_wt     = coverage(est_wt,   se_jk_wt,     true_att_weighted),
    Plac_wt   = coverage(est_wt,   se_plac_wt,   true_att_weighted),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric) & !c(N1, gamma, hhi), ~ round(., 2)),
         hhi = round(hhi, 4))

kableExtra::kbl(
  cov_tbl, booktabs = TRUE,
  caption = "Monte Carlo 95\\% confidence-interval coverage.", label = "mc-coverage",
  col.names = c("$N_1$", "$\\gamma$", "HHI", "Boot", "JK", "Plac", "Boot", "JK", "Plac"),
  escape = FALSE, align = c("r", "r", "r", rep("r", 6))
) %>%
  kableExtra::add_header_above(c(" " = 3, "Equally weighted" = 3, "Population weighted" = 3)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kableExtra::footnote(
    general = "Nominal level 0.95. Bootstrap (Boot) and placebo (Plac) variance estimators use 50 replications; JK is the fixed-weight jackknife. 100 simulations per configuration. Coverage is evaluated against each estimator's own target ATT.",
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```

```{r mc-aca-summary, eval=mc_has_aca, include=FALSE}
## Scalars for the ACA-calibrated cell (the N1 = 1181 rows), referenced in the
## conditional paragraph below.
aca <- mc_df %>% filter(N1 > 100)
aca_bias_wt <- mean(aca$est_wt - aca$true_att_weighted, na.rm = TRUE)
aca_rmse_wt <- sqrt(mean((aca$est_wt - aca$true_att_weighted)^2, na.rm = TRUE))
aca_tau_wt  <- mean(aca$true_att_weighted, na.rm = TRUE)
aca_cov_boot <- coverage(aca$est_wt, aca$se_boot_wt, aca$true_att_weighted)
aca_cov_jk   <- coverage(aca$est_wt, aca$se_jk_wt,   aca$true_att_weighted)
aca_cov_plac <- coverage(aca$est_wt, aca$se_plac_wt, aca$true_att_weighted)
```

`r if (mc_has_aca) sprintf("The ACA-calibrated configuration (the $N_1 = 1{,}181$ rows of Tables \\ref{tab:mc-bias} and \\ref{tab:mc-coverage}) speaks directly to the application: at the application's own panel dimensions and weight concentration, the population-weighted estimator has bias %.3f against its target (RMSE %.3f, target ATT %.1f), and the bootstrap attains %.0f\\%% coverage, the jackknife %.0f\\%%, and the uniform-weight placebo %.0f\\%%. The bootstrap's performance in this cell is the simulation counterpart of the cluster-robust bootstrap we rely on for Table 2.", aca_bias_wt, aca_rmse_wt, aca_tau_wt, 100 * aca_cov_boot, 100 * aca_cov_jk, 100 * aca_cov_plac)`

`r if (mc_available && !mc_has_aca) "*The ACA-calibrated configuration described in Section 4.1 is not present in this build's* \\texttt{mc\\_results.csv}*; re-running* \\texttt{scripts/run\\_mc\\_simulations.R} *computes it (the completed grid configurations are skipped) and adds its row to the tables.*"`

```{r mc-trend-run, cache=FALSE}
## Trend cell (Section 4.1): a size-correlated trend added to L. Diagnosis only
## --- shows the population-weighted estimators inherit bias and the in-time
## placebo detects it. Generated by scripts/run_mc_trend.R (no remedy columns:
## the paper's use of this cell is diagnostic, not a remedy horse-race).
mc_trend_available <- file.exists(file.path("data", "mc_trend_results.csv"))
mc_trend_df <- if (mc_trend_available) {
  readr::read_csv(file.path("data", "mc_trend_results.csv"), show_col_types = FALSE)
} else NULL
```

`r if (mc_trend_available) "Table \\ref{tab:mc-trend} isolates the mechanism behind the application's placebo failure. A size-correlated trend leaves the equally-weighted estimator unbiased for its target but induces a growing bias in both the population-weighted SDID and the population-weighted DID estimators---the bias travels with the weights, not the estimator---while the in-time placebo, which is zero in expectation absent a trend, grows with $\\gamma_{\\text{trend}}$ and so detects it. This is the simulation counterpart of Section 3.6: a differential trend that loads on the largest units contaminates the population-weighted target under any level-matched design, and the estimand-specific placebo is what surfaces it."`

```{r mc-trend-table, eval=mc_trend_available}
trend_tbl <- mc_trend_df %>%
  group_by(gamma_trend) %>%
  summarise(
    Bias_eq      = mean(est_sdid_eq - true_att_equal,    na.rm = TRUE),
    Bias_sdid_wt = mean(est_sdid_wt - true_att_weighted, na.rm = TRUE),
    Bias_did_wt  = mean(est_did_wt  - true_att_weighted, na.rm = TRUE),
    Placebo_wt   = mean(est_plac_wt,                     na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(across(-gamma_trend, ~ round(., 3)))

kableExtra::kbl(
  trend_tbl, booktabs = TRUE,
  caption = "Monte Carlo: a size-correlated trend biases the population-weighted estimators and is detected by the in-time placebo.",
  label = "mc-trend",
  col.names = c("$\\gamma_{\\text{trend}}$", "Bias (eq.)",
                "Bias (pop., SDID)", "Bias (pop., DID)", "In-time placebo"),
  escape = FALSE, align = c("r", "r", "r", "r", "r")
) %>%
  kableExtra::kable_styling(latex_options = c("hold_position")) %>%
  kableExtra::footnote(
    general = "200 simulations per cell at $N_1 = 20$, $\\\\gamma = 0.5$, and moderate weight concentration (HHI $= 0.15$). Bias is computed against each estimator's own target ATT; the population-weighted target is held fixed across cells (the trend perturbs the factor structure $L$, not the treatment effect). The final column is the mean in-time placebo estimate under population weights, which is zero in expectation when $\\\\gamma_{\\\\text{trend}} = 0$. Detrended and stratified variants are omitted here and reported in the supplementary appendix.",
    general_title = "Notes:", threeparttable = TRUE, escape = FALSE)
```


# Discussion

We have shown that the SDID estimator of @arkhangelsky2021 admits a parsimonious extension to incorporate researcher-specified weights on treated units, with consistency and asymptotic normality results that inherit from the underlying factor-model framework subject to a no-dominant-weight regularity condition. We have also documented, through both formal characterization and direct application, that this regularity condition loses its practical bite in many policy settings, particularly those involving population weights on heterogeneously sized geographic units. In such settings, plug-in asymptotic standard errors are unreliable; cluster-robust bootstrap inference at a level of aggregation where the no-dominant-cluster condition is plausible is the appropriate finite-sample tool. Our application provides a concrete illustration: with $N_1^{\text{eff}} \approx `r round(n1_eff, 0)`$ and one county carrying $\approx `r round(top1_share * 100, 1)`\%$ of total weight, the asymptotic approximation cannot be trusted at face value, but state-level cluster bootstrapping across the `r n_states_treated` treated state clusters, complemented by randomization inference, provides a defensible alternative.

The ACA application illustrates more than the technical contribution. The equally-weighted and population-weighted estimands diverge by `r round(abs(as.numeric(tau_sdid_w) - as.numeric(tau_sdid)), 0)` deaths per 100,000---enough to swing the qualitative interpretation of the policy from "no detectable effect" to "substantial mortality reduction." The in-time placebo of Section 3.6 shows that the population-weighted *level* is not credibly identified in this county design: it loads on a size-correlated differential mortality trend that convex, level-matched estimators cannot remove, and the same falsification would challenge any level-matched population-weighted design on this panel, including the published benchmark. This is a property of the ACA county setting, not of the weighting; the recent triple-difference and administrative-data literature is built precisely to identify the population-scaled mortality effects that a single-difference county design cannot [@miller2021; @wyse2025]. What survives regardless is the paper's central point: the equally-weighted target is a clean, precisely-estimated null, so the *divergence* between estimands is real whatever the true population-weighted magnitude proves to be. The choice between these estimands is not a statistical matter but a substantive one. A researcher interested in the average effect across counties as administrative units might reasonably prefer equal weighting; a researcher interested in the per-capita effect on the population exposed to expanded Medicaid coverage, as is the natural reading for cost-benefit and welfare calculations, should prefer population weighting. Both estimands are well-defined and identifiable; the published mortality literature has implicitly favored the population-weighted target through cluster sampling, propensity-score weighting, or other means [@borgschulte2020; @miller2021]. Our contribution makes this choice transparent and provides a tool that respects it directly.

Three caveats merit discussion. First, our treated-unit weights are fixed by the researcher rather than estimated. We do not claim optimality in any decision-theoretic sense; the weights encode the researcher's definition of the estimand. Adapting the weights to optimally estimate a target subpopulation effect (in the spirit of @hirano2003 for propensity-score weighting) is a natural extension we leave for future work. Second, while we provide cluster-robust bootstrap and jackknife inference, our adaptation of the placebo variance estimator uses uniform weights for pseudo-treated units, which may understate variance when treatment effects are highly heterogeneous; the implications for finite-sample coverage are examined in the companion supplement. Third, our main development and application assume simultaneous adoption. Section 2.5 sketches the extension to staggered settings---cohort-wise weighted estimation against not-yet-treated controls, aggregated with researcher-chosen cohort weights, implemented in the software and formalized in the supplement---but we do not pursue a staggered application here. The interaction between within-cohort and across-cohort weighting connects naturally to the broader literature on aggregation in heterogeneous-effects panel designs [@callaway2021; @sun2021; @dechaisemartin2020], and a substantive staggered application is left to future work.


\section*{Declaration of generative AI and AI-assisted technologies in the manuscript preparation process}

During the preparation of this work the authors used Claude Code (Anthropic) for R package development, code development, data processing, replication of prior work, code review, and manuscript editing. The authors reviewed and verified all analyses, results, and text, and take full responsibility for the content of the published article.


\section*{Data availability}

Data used in this study are publicly available from CDC WONDER (mortality, Underlying Cause of Death file, all-cause mortality among adults aged 20--64, ICD-10 codes A00--Y89; county-year cells with 1--9 deaths are suppressed in the public file and are imputed by allocating exact state-year residuals, as described in Supplementary Appendix E.5) [@cdcwonder]; the SEER program (county-level population by age and sex) [@seer]; and the Bureau of Labor Statistics (county-level unemployment, Local Area Unemployment Statistics) [@bls_laus]. The BV reproduction in Supplementary Appendix E.5 additionally uses the ACS 2009--2013 5-year file and Census SAHIE 2013 (via the Census API), and county presidential returns from the MIT Election Data and Science Lab. Replication code, the weighted \texttt{synthdid} R package extension, and a companion supplementary appendix containing full proofs of Propositions 1--3, synthetic-control diagnostics, alternative weight specifications, leave-one-out sensitivity, and a complete description of the DLPS implementation are available at \url{https://github.com/johniselin-econ/synthdid_weights}.
