---
title: "Weighted Synthetic Difference-in-Differences"
author: "John Iselin\\thanks{The Budget Lab at Yale. Corresponding author: john.iselin@yale.edu} \\and Erica Ryan\\thanks{Amazon}"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: true
    keep_tex: true
    fig_caption: true
    citation_package: natbib
    includes:
      in_header: header.tex
bibliography: references.bib
biblio-style: elsarticle-harv
fontsize: 12pt
geometry: margin=1in
linestretch: 1.5
abstract: |
  The synthetic difference-in-differences estimator of @arkhangelsky2021 assigns equal weight to treated units, which may not match the estimand of interest when units differ in size. We propose a weighted SDID estimator incorporating user-specified weights into both synthetic control construction and effect estimation. Applying the method to ACA Medicaid expansion and county-level mortality, the equally-weighted estimate is near zero while the population-weighted estimate indicates a reduction of approximately 13.5 deaths per 100,000, demonstrating that the weighting choice can be consequential.

  \medskip\noindent
  \textbf{Keywords:} Synthetic difference-in-differences, causal inference, panel data, weighted estimation, Medicaid expansion

  \medskip\noindent
  \textbf{JEL Codes:} C21, C23, I13
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4,
  fig.align = "center",
  out.width = "90%",
  cache = TRUE
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(synthdid)

# Source weighted SDID functions
source("../R/utils.R")
source("../R/solver.R")
source("../R/synthdid.R")
source("../R/vcov.R")
source("../R/summary.R")
source("../R/print.R")

# Reproducibility
set.seed(20240101)
```

# Introduction

The synthetic difference-in-differences (SDID) estimator of @arkhangelsky2021 combines the strengths of the synthetic control method and difference-in-differences by using data-driven unit weights $\hat{\omega}$ and time weights $\hat{\lambda}$ while allowing level differences between groups. When multiple units are treated, the standard implementation averages them equally: each treated unit contributes $1/N_1$ to the estimand. This is an implicit design choice, not a requirement of the methodology.

In many empirical settings, equal weighting is inappropriate. When studying a state-level policy, a researcher interested in the per-capita treatment effect should weight by population. If treatment effects are heterogeneous and correlated with unit size, the equally-weighted and population-weighted estimands will diverge. We propose a weighted SDID estimator that replaces the uniform $1/N_1$ averaging with researcher-chosen weights $\tilde{\omega}_j$ in the collapsed form of the outcome matrix. The modification is minimal---one equation change---but it affects the estimation of both control unit weights $\hat{\omega}$ and time weights $\hat{\lambda}$, since the synthetic control is now constructed to match a differently-weighted treated average. We adapt all three variance estimators (bootstrap, jackknife, placebo) to handle weight renormalization during resampling. Software is available as an extension to the \texttt{synthdid} R package.\footnote{Available at \url{https://github.com/johniselin-econ/synthdid_weights}. See the Supplementary Appendix for inference details and extensions.}

We apply the method to the county-level mortality effects of the ACA's 2014 Medicaid expansion, following @borgschulte2020. Treated counties range from over 10 million to under 10,000 residents. The equally-weighted SDID estimate is near zero; the population-weighted estimate indicates a mortality reduction of approximately 13.5 deaths per 100,000---a sign change driven entirely by the weighting choice. This demonstrates that when treatment effects are heterogeneous across units of different sizes, the choice of treated-unit weights is economically consequential.


# The Weighted Estimator

## Setup and standard SDID

Consider a balanced panel with $N$ units over $T$ periods, where the first $N_0$ are untreated controls and the remaining $N_1 = N - N_0$ begin treatment at $T_0 + 1$. The SDID estimator of @arkhangelsky2021 operates on a *collapsed form* $\tilde{Y}$, an $(N_0+1) \times (T_0+1)$ matrix in which the last row averages treated units and the last column averages post-treatment periods:
\begin{equation} \label{eq:collapsed}
\tilde{Y}_{N_0+1, t} = \frac{1}{N_1} \sum_{j=1}^{N_1} Y_{N_0+j,t}, \quad
\tilde{Y}_{i, T_0+1} = \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it}.
\end{equation}
Control unit weights $\hat{\omega} \in \Delta^{N_0}$ and time weights $\hat{\lambda} \in \Delta^{T_0}$ are estimated from $\tilde{Y}$ via regularized Frank-Wolfe optimization on the simplex, and the treatment effect estimate is
\begin{equation} \label{eq:tau-sdid}
\hat{\tau}^{\text{sdid}} = \begin{pmatrix} -\hat{\omega} \\ \frac{1}{N_1}\mathbf{1}_{N_1} \end{pmatrix}^\top Y \begin{pmatrix} -\hat{\lambda} \\ \frac{1}{T_1}\mathbf{1}_{T_1} \end{pmatrix}.
\end{equation}
The key observation is that the target on the right-hand side of the weight estimation problem---the average treated outcome in \eqref{eq:collapsed}---uses equal weights $1/N_1$. This determines which synthetic control group is constructed.

## Weighted modification

Let $\tilde{\omega} = (\tilde{\omega}_1, \ldots, \tilde{\omega}_{N_1})$ be researcher-specified weights on treated units, normalized so $\sum_j \tilde{\omega}_j = 1$. The target estimand is $\tau^{(\tilde{\omega})} = \sum_j \tilde{\omega}_j \bar{\tau}_j$, where $\bar{\tau}_j = T_1^{-1}\sum_{t=T_0+1}^{T} \tau_{jt}$ is the average treatment effect for unit $j$ across post-treatment periods. Our modification replaces the uniform averages in \eqref{eq:collapsed} with weighted averages:
\begin{equation} \label{eq:collapsed-w}
\tilde{Y}^w_{N_0+1, t} = \sum_{j=1}^{N_1} \tilde{\omega}_j Y_{N_0+j,t}, \quad
\tilde{Y}^w_{i, T_0+1} = \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it}.
\end{equation}
Control weights $\hat{\omega}^w$ and time weights $\hat{\lambda}^w$ are estimated from $\tilde{Y}^w$ via the same algorithm---no modification to the solver is needed---and the treatment effect is
\begin{equation} \label{eq:tau-w}
\hat{\tau}^{w} = \begin{pmatrix} -\hat{\omega}^w \\ \tilde{\omega} \end{pmatrix}^\top Y \begin{pmatrix} -\hat{\lambda}^w \\ \frac{1}{T_1}\mathbf{1}_{T_1} \end{pmatrix}.
\end{equation}
Comparing \eqref{eq:tau-sdid} and \eqref{eq:tau-w}, the modification is concentrated in two places: the collapsed form that feeds the weight solver, and the treated-unit block of the final bilinear form. When $\tilde{\omega}_j = 1/N_1$ for all $j$, the weighted estimator reduces to the standard SDID (proof in Supplementary Appendix C). Weighted DiD ($\hat{\omega} = \mathbf{1}_{N_0}/N_0$, $\hat{\lambda} = \mathbf{1}_{T_0}/T_0$) and weighted SC ($\hat{\lambda} = \mathbf{0}$) emerge as special cases.

## Inference

We adapt all three variance estimators of @arkhangelsky2021 to handle weight renormalization during resampling. The key complication is the bootstrap: in a given draw $b$, treated unit $j$ appears $n_j^{(b)}$ times, and the bootstrap-adjusted weights are
\begin{equation} \label{eq:boot-w}
\tilde{\omega}_j^{(b)} = \frac{\tilde{\omega}_j \cdot n_j^{(b)}}{\sum_{k=1}^{N_1} \tilde{\omega}_k \cdot n_k^{(b)}}.
\end{equation}
The jackknife renormalizes analogously when leaving out a treated unit; the placebo method uses uniform weights for pseudo-treated control units, since the original $\tilde{\omega}$ do not map naturally to control units (see Supplementary Appendix B for details and alternative placebo weighting schemes).


# Application: ACA Medicaid Expansion

## Data

```{r load-data}
panel <- readr::read_csv(
  file.path("..", "data", "analysis_data.csv"),
  show_col_types = FALSE
) %>%
  select(
    unit = fips, time = year, y = crude_rate,
    treated_unit = expansion, pop = population,
    state_fips
  ) %>%
  mutate(post = as.integer(time >= 2014))

n_counties <- n_distinct(panel$unit)
n_treated <- n_distinct(panel$unit[panel$treated_unit == 1])
n_control <- n_distinct(panel$unit[panel$treated_unit == 0])
year_range <- range(panel$time)
n_years <- n_distinct(panel$time)
T0 <- sum(sort(unique(panel$time)) < 2014)
T1 <- n_years - T0

treated_pop <- panel %>%
  filter(treated_unit == 1, time == 2013) %>%
  summarise(
    total = sum(pop, na.rm = TRUE),
    median = median(pop, na.rm = TRUE),
    p10 = quantile(pop, 0.10, na.rm = TRUE),
    p90 = quantile(pop, 0.90, na.rm = TRUE)
  )
```

We use a balanced county-year panel covering `r year_range[1]`--`r year_range[2]` with `r format(n_counties, big.mark = ",")` counties (`r format(n_treated, big.mark = ",")` treated, `r format(n_control, big.mark = ",")` control), following @borgschulte2020. The outcome is the all-cause mortality rate per 100,000 among adults aged 20--64 (CDC WONDER). The treated-unit weights $\tilde{\omega}_j$ are each county's share of total treated-county population as of 2013. The population distribution is highly skewed: the 90th percentile county (`r format(round(treated_pop$p90), big.mark = ",")` residents) is an order of magnitude larger than the 10th percentile (`r format(round(treated_pop$p10), big.mark = ",")`).

## Results

```{r prepare-sdid-panel}
panel_sdid <- panel %>%
  arrange(unit, time) %>%
  select(unit, time, y, treated_unit, pop, state_fips) %>%
  mutate(.unit = as.factor(unit), .time = time,
         .W = as.integer(treated_unit == 1 & time >= 2014))

dfp <- panel_sdid %>% select(.unit, .time, y, .W) %>% as.data.frame()
setup <- synthdid::panel.matrices(dfp)

# Build population weights for treated counties, ordered to match setup$Y
pop_2013 <- panel %>% filter(treated_unit == 1, time == 2013) %>% select(unit, pop)
treated_names <- rownames(setup$Y)[(setup$N0 + 1):nrow(setup$Y)]
pop_ordered <- pop_2013 %>% mutate(unit_char = as.character(unit))
weight_order <- match(treated_names, pop_ordered$unit_char)
treated_weights <- pop_ordered$pop[weight_order]
treated_weights <- treated_weights / sum(treated_weights, na.rm = TRUE)

# Build state cluster vector matching setup$Y row order
unit_state <- panel_sdid %>%
  distinct(unit, state_fips) %>%
  mutate(unit_char = as.character(unit))
all_names <- rownames(setup$Y)
cluster_vec <- unit_state$state_fips[match(all_names, unit_state$unit_char)]
n_states <- n_distinct(cluster_vec)
```

```{r run-estimates, cache=TRUE}
tau_sdid   <- synthdid_estimate(setup$Y, setup$N0, setup$T0)
tau_did    <- did_estimate(setup$Y, setup$N0, setup$T0)

tau_sdid_w <- synthdid_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                          treated.weights = treated_weights,
                                          cluster = cluster_vec)
tau_did_w  <- did_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                     treated.weights = treated_weights,
                                     cluster = cluster_vec)
```

```{r compute-ses, cache=TRUE}
# State-clustered bootstrap SEs (treatment assigned at state level)
se_sdid   <- tryCatch(sqrt(vcov(tau_sdid,   method = "jackknife")), error = function(e) NA)
se_did    <- tryCatch(sqrt(vcov(tau_did,    method = "jackknife")), error = function(e) NA)
se_sdid_w <- tryCatch(suppressWarnings(sqrt(vcov(tau_sdid_w, method = "bootstrap", replications = 200))),
                       error = function(e) NA)
se_did_w  <- tryCatch(suppressWarnings(sqrt(vcov(tau_did_w,  method = "bootstrap", replications = 200))),
                       error = function(e) NA)
```

```{r results-table}
results <- tibble::tibble(
  Estimator = c("DiD", "SDID"),
  `Equally wtd.` = c(as.numeric(tau_did), as.numeric(tau_sdid)),
  `SE (eq.)` = c(se_did, se_sdid),
  `Pop.-weighted` = c(as.numeric(tau_did_w), as.numeric(tau_sdid_w)),
  `SE (wt.)` = c(se_did_w, se_sdid_w)
)

kableExtra::kbl(
  results,
  digits = 2,
  booktabs = TRUE,
  caption = "Effect of ACA Medicaid expansion on mortality (deaths per 100,000, ages 20--64). Equally-weighted SEs: county-level jackknife. Population-weighted SEs: state-clustered bootstrap (200 replications).",
  col.names = c("Estimator", "Estimate", "SE", "Estimate", "SE"),
  align = c("l", "r", "r", "r", "r")
) %>%
  kableExtra::add_header_above(c(" " = 1, "Equally weighted" = 2, "Population weighted" = 2)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position")) %>%
  kableExtra::footnote(general = sprintf("Sample: %s counties, %s--%s. Outcome: all-cause mortality per 100,000, ages 20--64 (CDC WONDER). %s states as clusters.",
    format(nrow(setup$Y), big.mark = ","), year_range[1], year_range[2], n_states),
    general_title = "", threeparttable = TRUE)
```

Table 1 reports the main results. The equally-weighted and population-weighted estimates diverge dramatically---indeed, they differ in sign. The equally-weighted (i.e., standard) SDID estimate is near zero, indicating no detectable mortality effect when all treated counties receive equal weight. The population-weighted SDID estimate is approximately $-13.5$ deaths per 100,000, indicating a substantial mortality reduction when the estimand weights counties by population. The DiD estimates show a similar pattern: a near-zero equally-weighted estimate and a large negative population-weighted estimate. This divergence arises because treatment effects are heterogeneous and correlated with county size: larger counties experience larger mortality reductions from Medicaid expansion, which may reflect greater baseline uninsurance rates and more scope for coverage gains in urban areas \citep{sommers2016, miller2021}.

```{r event-study-both, fig.cap="Event-study decomposition under equal and population weighting. Left panel: DiD. Right panel: SDID. Shaded bands show 95\\% bootstrap confidence intervals (200 replications). The dashed vertical line marks treatment onset (2014).", fig.height=4, fig.width=8, cache=TRUE}
es_sdid   <- synthdid_event_study(tau_sdid,   se.method = "bootstrap", replications = 200)
es_sdid_w <- synthdid_event_study(tau_sdid_w, se.method = "bootstrap", replications = 200)
es_did    <- synthdid_event_study(tau_did,    se.method = "bootstrap", replications = 200)
es_did_w  <- synthdid_event_study(tau_did_w,  se.method = "bootstrap", replications = 200)

es_df <- bind_rows(
  es_did    %>% mutate(Weighting = "Equally weighted", Estimator = "DiD"),
  es_did_w  %>% mutate(Weighting = "Population weighted", Estimator = "DiD"),
  es_sdid   %>% mutate(Weighting = "Equally weighted", Estimator = "SDID"),
  es_sdid_w %>% mutate(Weighting = "Population weighted", Estimator = "SDID")
) %>%
  mutate(
    year = as.numeric(time),
    Estimator = factor(Estimator, levels = c("DiD", "SDID"))
  )

ggplot(es_df, aes(x = year, y = estimate, color = Weighting, fill = Weighting)) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.15, color = NA) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 2013.5, linetype = "dashed", color = "gray30") +
  geom_point(aes(shape = Weighting), size = 1.3) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Estimator) +
  scale_color_manual(values = c("Equally weighted" = "steelblue", "Population weighted" = "firebrick")) +
  scale_fill_manual(values = c("Equally weighted" = "steelblue", "Population weighted" = "firebrick")) +
  scale_shape_manual(values = c("Equally weighted" = 16, "Population weighted" = 17)) +
  labs(x = NULL, y = "Effect on mortality\n(deaths per 100,000)", color = NULL, fill = NULL, shape = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom")
```

Figure 1 presents the event-study decomposition for both DiD (left) and SDID (right) under equal and population weighting. Pre-treatment coefficients cluster near zero under both methods and weighting schemes, supporting the parallel-trends assumption. Post-treatment, the two weighting schemes diverge sharply in both panels: the population-weighted coefficients are consistently negative ($-11$ to $-17$ deaths per 100,000), while the equally-weighted coefficients fluctuate near zero. The SDID pre-treatment fit is slightly tighter than DiD, reflecting the data-driven time weights, but the key finding is consistent across both estimators: the weighting choice---not the estimator choice---drives the main result.

```{r weight-stats}
hhi <- sum(treated_weights^2)
n1_eff <- 1 / hhi
top10_share <- sum(sort(treated_weights, decreasing = TRUE)[1:10])
```

The weight concentration statistics illuminate why the estimates diverge so dramatically. The population weights have a Herfindahl index of `r round(hhi, 4)`, implying an effective number of treated units $N_1^{\text{eff}} = `r round(n1_eff, 0)`$ compared to the raw $N_1 = `r nrow(setup$Y) - setup$N0`$. The ten largest counties account for `r round(top10_share * 100, 0)`\% of total treated-county weight. Standard errors are modestly larger under population weighting (`r round(se_sdid_w, 1)` vs. `r round(se_sdid, 1)` for SDID), reflecting greater heterogeneity in the reweighted sample. The consistency of results across DiD and SDID confirms that the finding is robust to the choice of estimator. As a further check, Supplementary Appendix F compares our population-weighted SDID to the double-lasso propensity score weighted DiD of @borgschulte2020, an entirely independent approach to constructing a population-relevant comparison group.

Because treatment is assigned at the state level while the unit of observation is the county, we report state-clustered bootstrap standard errors for the population-weighted estimates, resampling entire states with replacement following @clarke2023. The equally-weighted estimates use county-level jackknife SEs. Further discussion of clustered inference for SDID is in Supplementary Appendix G.


# Conclusion

We have proposed a simple but consequential extension to the SDID estimator: replacing uniform treated-unit averaging with researcher-chosen weights. The modification requires changing only the collapsed form of the outcome matrix; the optimization algorithm proceeds identically. The ACA application demonstrates that the weighting choice can determine not just the magnitude but the sign of the estimated treatment effect---an issue likely to arise in any setting where treatment effects are heterogeneous and correlated with unit size. Important extensions include staggered treatment adoption, where the weighted estimator can be applied within each adoption cohort, and cluster-robust inference for settings where treatment assignment occurs at a higher level than the unit of observation (Supplementary Appendix G). Monte Carlo simulations examining the finite-sample coverage properties of the adapted variance estimators are reported in Supplementary Appendix E.

\section*{Declaration of generative AI and AI-assisted technologies in the manuscript preparation process}

During the preparation of this work the authors used Claude Code (Anthropic) for R package development, code review, and manuscript formatting. The authors reviewed and edited all content and take full responsibility for the content of the published article.

\section*{Data availability}

Data used in this study are publicly available from CDC WONDER (mortality), SEER (population), and BLS (unemployment). Replication code and the weighted \texttt{synthdid} R package are available at \url{https://github.com/johniselin-econ/synthdid_weights}.
