---
title: "Weighted Synthetic Difference-in-Differences"
author: "John Iselin\\thanks{The Budget Lab at Yale} \\and Erica Ryan\\thanks{Amazon}"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: true
    keep_tex: true
    fig_caption: true
    citation_package: natbib
    includes:
      in_header: header.tex
bibliography: references.bib
biblio-style: elsarticle-harv
fontsize: 12pt
geometry: margin=1in
linestretch: 1.5
abstract: |
  The synthetic difference-in-differences (SDID) estimator of @arkhangelsky2021 combines the strengths of synthetic control and difference-in-differences methods. However, its standard implementation assigns equal weight to all treated units when computing the average treatment effect, which may not correspond to the estimand of interest when treated units are heterogeneous in size or policy relevance. We propose a weighted SDID estimator that incorporates user-specified weights on treated units---for example, population shares---into both the construction of the synthetic control and the final effect estimate. We adapt the bootstrap, jackknife, and placebo variance estimators to account for weight renormalization during resampling. We apply the method to county-level mortality effects of the Affordable Care Act's 2014 Medicaid expansion, following @borgschulte2020, where treated counties range from Los Angeles County (over 10 million residents) to small rural counties under 10,000---making the choice between equally-weighted and population-weighted estimation economically consequential. We conclude with a discussion of extensions to staggered treatment adoption, time-varying period weights, and clustered standard errors when treatment is assigned at a higher level than the unit of observation.

  \medskip\noindent
  \textbf{Keywords:} Synthetic difference-in-differences, causal inference, panel data, weighted estimation, Medicaid expansion

  \medskip\noindent
  \textbf{JEL Codes:} C21, C23, I13
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 6.5,
  fig.height = 4,
  fig.align = "center",
  out.width = "90%",
  cache = TRUE
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(synthdid)

# Source weighted SDID functions
source("../R_weighted/utils.R")
source("../R_weighted/solver.R")
source("../R_weighted/synthdid.R")
source("../R_weighted/vcov.R")
source("../R_weighted/summary.R")
source("../R_weighted/print.R")

# Reproducibility
set.seed(20240101)
```

# Introduction

Panel data methods for causal inference have become central to empirical economics. Two dominant approaches---difference-in-differences (DiD) and the synthetic control method (SCM)---each offer distinct advantages. DiD leverages parallel trends assumptions and accommodates multiple treated units naturally, while SCM constructs data-driven control groups that can provide better pre-treatment fit. @arkhangelsky2021 introduced synthetic difference-in-differences (SDID), which combines both: it uses data-driven unit weights $\hat{\omega}$ (like SCM) and time weights $\hat{\lambda}$ (unique to SDID) while allowing for level differences between treated and control groups (like DiD).

A key feature of the SDID estimator---inherited from its SCM component---is that it was originally designed for settings with a single treated unit. When applied to multiple treated units, the standard implementation averages treated units equally: each treated unit contributes $1/N_1$ to the estimand. This is an implicit design choice, not a requirement of the methodology. In many empirical settings, equal weighting is inappropriate. When studying the effect of a state-level policy, for instance, a researcher interested in the per-capita treatment effect should weight states by population. If treatment effects are heterogeneous and correlated with unit size, the equally-weighted and population-weighted estimands will diverge.

We make three contributions. First, we extend the SDID estimator to incorporate user-specified weights on treated units. The modification is straightforward: we replace the uniform $1/N_1$ averaging with researcher-chosen weights $\tilde{\omega}_j$ in the collapsed form of the outcome matrix, which in turn affects the estimation of both unit weights $\hat{\omega}$ and time weights $\hat{\lambda}$. Second, we adapt all three variance estimators proposed by @arkhangelsky2021---bootstrap, jackknife, and placebo---to handle the complications that arise when resampling must respect a weight structure. Third, we apply the method to the county-level mortality effects of the ACA's 2014 Medicaid expansion, following @borgschulte2020. Treated counties range from Los Angeles County (over 10 million residents) to small rural counties with populations under 10,000, making the choice of treated-unit weights economically consequential. The application also raises the question of clustered inference when treatment is assigned at a higher level (state) than the unit of observation (county), which we discuss as an extension.

The remainder of the paper proceeds as follows. Section \ref{sec:setup} reviews the SDID estimator and introduces notation. Section \ref{sec:weighted} presents the weighted extension. Section \ref{sec:inference} develops the adapted variance estimators. Section \ref{sec:application} presents the ACA application. Section \ref{sec:extensions} discusses staggered adoption and time-varying weights. Section \ref{sec:conclusion} concludes.


# Setup and the Standard SDID Estimator {#sec:setup}

## Model and Notation

Consider a balanced panel with $N$ units observed over $T$ periods. The first $N_0$ units are untreated controls; the remaining $N_1 = N - N_0$ are treated beginning at period $T_0 + 1$. Let $T_1 = T - T_0$ denote the number of post-treatment periods. Define the treatment indicator
$$
W_{it} = \mathbf{1}\{i > N_0\} \cdot \mathbf{1}\{t > T_0\}.
$$

The observed outcome is
$$
Y_{it} = Y_{it}(0) + \tau_{it} \cdot W_{it},
$$
where $Y_{it}(0)$ is the potential outcome absent treatment and $\tau_{it}$ is the unit-time treatment effect. The object of interest is an average treatment effect on the treated (ATT), though the precise weighted average depends on how treated units are aggregated.

## The SDID Estimator

Following @arkhangelsky2021, the SDID estimator solves
\begin{equation} \label{eq:sdid-regression}
(\hat{\tau}^{\text{sdid}}, \hat{\mu}, \hat{\alpha}, \hat{\beta}) = \underset{\tau, \mu, \alpha_i, \beta_t}{\arg\min} \sum_{i=1}^{N} \sum_{t=1}^{T} \left( Y_{it} - \mu - \alpha_i - \beta_t - \tau \cdot W_{it} \right)^2 \hat{\omega}_i \hat{\lambda}_t,
\end{equation}
where $\hat{\omega}_i$ are unit weights and $\hat{\lambda}_t$ are time weights.

This admits a closed-form solution. Define the *collapsed form* $\tilde{Y}$, an $(N_0 + 1) \times (T_0 + 1)$ matrix in which the last row averages treated units and the last column averages post-treatment periods:
\begin{equation} \label{eq:collapsed}
\tilde{Y}_{N_0+1, t} = \frac{1}{N_1} \sum_{i=N_0+1}^{N} Y_{it}, \quad
\tilde{Y}_{i, T_0+1} = \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it}.
\end{equation}

The treatment effect estimate is then
\begin{equation} \label{eq:tau-sdid}
\hat{\tau}^{\text{sdid}} = \begin{pmatrix} -\hat{\omega} \\ \frac{1}{N_1}\mathbf{1}_{N_1} \end{pmatrix}^\top Y \begin{pmatrix} -\hat{\lambda} \\ \frac{1}{T_1}\mathbf{1}_{T_1} \end{pmatrix},
\end{equation}
a double-difference comparing the weighted treated-vs-control gap in the post-period to the same gap in the $\hat{\lambda}$-weighted pre-period.

## Weight Estimation

Control unit weights $\hat{\omega} \in \Delta^{N_0}$ (the unit simplex) are estimated by matching the weighted control average to the treated average in pre-treatment periods:
\begin{equation} \label{eq:omega}
(\hat{\omega}_0, \hat{\omega}) = \underset{\omega_0 \in \mathbb{R},\; \omega \in \Delta^{N_0}}{\arg\min} \sum_{t=1}^{T_0} \left( \omega_0 + \sum_{i=1}^{N_0} \omega_i Y_{it} - \frac{1}{N_1} \sum_{i=N_0+1}^{N} Y_{it} \right)^2 + \zeta_\omega^2 T_0 \|\omega\|_2^2.
\end{equation}

Time weights $\hat{\lambda} \in \Delta^{T_0}$ are estimated analogously, matching the $\hat{\lambda}$-weighted pre-treatment average of control units to their post-treatment average:
\begin{equation} \label{eq:lambda}
(\hat{\lambda}_0, \hat{\lambda}) = \underset{\lambda_0 \in \mathbb{R},\; \lambda \in \Delta^{T_0}}{\arg\min} \sum_{i=1}^{N_0} \left( \lambda_0 + \sum_{t=1}^{T_0} \lambda_t Y_{it} - \frac{1}{T_1} \sum_{t=T_0+1}^{T} Y_{it} \right)^2 + \zeta_\lambda^2 N_0 \|\lambda\|_2^2.
\end{equation}

Both optimization problems are solved via Frank-Wolfe on the simplex with exact line search, with regularization parameters $\zeta_\omega = \eta_\omega \cdot \hat{\sigma}$ and $\zeta_\lambda = \eta_\lambda \cdot \hat{\sigma}$, where $\hat{\sigma}$ is the noise level estimated from first differences of the control block.

**The key observation** is that the target on the right-hand side of \eqref{eq:omega}---the average treated outcome---uses equal weights $1/N_1$. This determines which synthetic control group is constructed. If the researcher's estimand places different weights on treated units, the control weights $\hat{\omega}$ should be re-estimated against the appropriately weighted treated average.


# The Weighted SDID Estimator {#sec:weighted}

## Estimand

Let $\tilde{\omega} = (\tilde{\omega}_1, \ldots, \tilde{\omega}_{N_1})$ be a vector of non-negative, researcher-specified weights on the treated units, normalized so that $\sum_{j=1}^{N_1} \tilde{\omega}_j = 1$. Similarly, let $\tilde{\pi} = (\tilde{\pi}_1, \ldots, \tilde{\pi}_{T_1})$ be weights on post-treatment periods, with $\sum_{s=1}^{T_1} \tilde{\pi}_s = 1$. The target estimand is the weighted ATT:
\begin{equation} \label{eq:weighted-att}
\tau^{(\tilde{\omega}, \tilde{\pi})} = \sum_{j=1}^{N_1} \sum_{s=1}^{T_1} \tilde{\omega}_j \tilde{\pi}_s \cdot \tau_{N_0+j, T_0+s}.
\end{equation}

When $\tilde{\omega}_j$ equals state $j$'s share of the treated population, $\tau^{(\tilde{\omega}, \tilde{\pi})}$ is the population-weighted ATT. When $\tilde{\omega}_j = 1/N_1$ and $\tilde{\pi}_s = 1/T_1$, we recover the standard SDID estimand.

## Weighted Collapsed Form

The modification to the SDID algorithm is concentrated in the collapsed form. Replace the uniform averages in \eqref{eq:collapsed} with weighted averages:
\begin{equation} \label{eq:weighted-collapsed}
\tilde{Y}^w = \begin{pmatrix}
Y_{1,1} & \cdots & Y_{1,T_0} & \sum_{s=1}^{T_1} \tilde{\pi}_s Y_{1,T_0+s} \\
\vdots & \ddots & \vdots & \vdots \\
Y_{N_0,1} & \cdots & Y_{N_0,T_0} & \sum_{s=1}^{T_1} \tilde{\pi}_s Y_{N_0,T_0+s} \\
\sum_{j=1}^{N_1} \tilde{\omega}_j Y_{N_0+j,1} & \cdots & \sum_{j=1}^{N_1} \tilde{\omega}_j Y_{N_0+j,T_0} & \sum_{j,s} \tilde{\omega}_j \tilde{\pi}_s Y_{N_0+j,T_0+s}
\end{pmatrix}.
\end{equation}

The last row is the $\tilde{\omega}$-weighted treated average; the last column is the $\tilde{\pi}$-weighted post-treatment average. The bottom-right element is doubly weighted.

## Modified Weight Estimation

Control unit weights are now estimated by matching against the *weighted* treated average:
\begin{equation} \label{eq:omega-w}
(\hat{\omega}_0^w, \hat{\omega}^w) = \underset{\omega_0, \omega \in \Delta^{N_0}}{\arg\min} \sum_{t=1}^{T_0} \left( \omega_0 + \sum_{i=1}^{N_0} \omega_i Y_{it} - \sum_{j=1}^{N_1} \tilde{\omega}_j Y_{N_0+j,t} \right)^2 + \zeta_\omega^2 T_0 \|\omega\|_2^2.
\end{equation}

Time weights are estimated analogously, with the post-treatment target using $\tilde{\pi}$-weighted period averages:
\begin{equation} \label{eq:lambda-w}
(\hat{\lambda}_0^w, \hat{\lambda}^w) = \underset{\lambda_0, \lambda \in \Delta^{T_0}}{\arg\min} \sum_{i=1}^{N_0} \left( \lambda_0 + \sum_{t=1}^{T_0} \lambda_t Y_{it} - \sum_{s=1}^{T_1} \tilde{\pi}_s Y_{i,T_0+s} \right)^2 + \zeta_\lambda^2 N_0 \|\lambda\|_2^2.
\end{equation}

The Frank-Wolfe solver is applied to $\tilde{Y}^w$ exactly as in the standard case---no algorithmic modification is needed beyond changing the input.

## Weighted Treatment Effect

The weighted SDID estimate is
\begin{equation} \label{eq:tau-weighted}
\hat{\tau}^{w} = \begin{pmatrix} -\hat{\omega}^w \\ \tilde{\omega} \end{pmatrix}^\top Y \begin{pmatrix} -\hat{\lambda}^w \\ \tilde{\pi} \end{pmatrix}.
\end{equation}

This differs from \eqref{eq:tau-sdid} in two respects: (i) the treated-unit block uses $\tilde{\omega}$ instead of $1/N_1$, and (ii) the post-treatment block uses $\tilde{\pi}$ instead of $1/T_1$. The control weights $\hat{\omega}^w$ and time weights $\hat{\lambda}^w$ will generally differ from their unweighted counterparts because they are estimated from the weighted collapsed form.

**Proposition 1.** *When $\tilde{\omega}_j = 1/N_1$ for all $j$ and $\tilde{\pi}_s = 1/T_1$ for all $s$, the weighted estimator $\hat{\tau}^w$ reduces to the standard SDID estimator $\hat{\tau}^{\text{sdid}}$.*

This follows immediately from the identity of the collapsed forms.

**Remark.** Weighted SC and weighted DiD emerge as special cases, paralleling the unweighted versions: setting $\hat{\lambda} = 0$ yields weighted SC; setting $\hat{\omega}_i = 1/N_0$ and $\hat{\lambda}_t = 1/T_0$ yields weighted DiD.


# Inference {#sec:inference}

@arkhangelsky2021 propose three approaches to variance estimation: bootstrap (Algorithm 2), jackknife (Algorithm 3), and placebo (Algorithm 4). Each involves resampling units in some fashion. With user-specified treated-unit weights, a complication arises: when a resampling scheme drops or duplicates treated units, the weights $\tilde{\omega}$ must be renormalized to maintain their interpretation.

## Bootstrap

The bootstrap resamples units with replacement. In a given bootstrap draw $b$, let $n_j^{(b)}$ denote the number of times treated unit $j$ appears. The bootstrap-adjusted weights are
\begin{equation} \label{eq:boot-weights}
\tilde{\omega}_j^{(b)} = \frac{\tilde{\omega}_j \cdot n_j^{(b)}}{\sum_{k=1}^{N_1} \tilde{\omega}_k \cdot n_k^{(b)}},
\end{equation}
with the convention that if no treated units are drawn (probability exponentially small in $N_1$), the draw is discarded. Control weights $\hat{\omega}$ are similarly renormalized. The bootstrap variance is
$$
\widehat{\text{Var}}^{\text{boot}} = \frac{B-1}{B} \cdot \frac{1}{B} \sum_{b=1}^{B} \left( \hat{\tau}^{w,(b)} - \bar{\tau}^w \right)^2.
$$

## Jackknife

The leave-one-out jackknife drops unit $i$ in turn. When unit $i$ is a treated unit $j$, the remaining treated weights are renormalized:
\begin{equation} \label{eq:jk-weights}
\tilde{\omega}_k^{(-j)} = \frac{\tilde{\omega}_k}{\sum_{\ell \neq j} \tilde{\omega}_\ell}, \quad k \neq j.
\end{equation}

As in the standard case, we use the *fixed-weight* jackknife, holding $\hat{\omega}^w$ and $\hat{\lambda}^w$ at their full-sample values while renormalizing the control weight vector when a control unit is dropped. The jackknife variance is
$$
\widehat{\text{Var}}^{\text{jk}} = \frac{N-1}{N} \sum_{i=1}^{N} \left( \hat{\tau}^{w,(-i)} - \bar{\tau}^{w,(\cdot)} \right)^2.
$$

Following @arkhangelsky2021, the jackknife is not recommended for the SC variant due to the non-differentiability of the SC weights at the boundary of the simplex.

## Placebo

The placebo method randomly assigns $N_1$ control units as "placebo treated" and estimates the treatment effect on them, repeating $B$ times. Since the original treated-unit weights $\tilde{\omega}$ do not naturally map to control units serving as placebo treated, we use uniform weights $1/N_1$ for the placebo-treated group. This approach tests the sharp null of no treatment effect and requires $N_0 > N_1$.


# Application: The ACA Medicaid Expansion {#sec:application}

## Background

The Affordable Care Act authorized states to expand Medicaid eligibility to adults with incomes up to 138\% of the federal poverty level beginning January 1, 2014. Twenty-five states and the District of Columbia expanded in the first half of 2014, while the remaining states did not (or expanded later), creating a natural panel experiment.\footnote{We exclude states that expanded after mid-2014 (Alaska, Indiana, Louisiana, Montana, New Hampshire, Pennsylvania) to maintain the simultaneous-adoption structure required by SDID.} This setting has been studied extensively; we follow the empirical framework of @borgschulte2020, who estimate the effect of Medicaid expansion on county-level mortality.

The unit of observation is the county, but treatment is assigned at the state level. This creates a setting where the weighting problem is especially salient: counties in expansion states range from Los Angeles County (over 10 million residents) to small rural counties with populations under 10,000. The standard SDID estimator treats each county equally, so a county of 5,000 receives the same weight as one of 5 million. When the research question concerns the per-capita mortality effect of Medicaid expansion---as is natural for policy evaluation---population-weighted estimation is required.

## Data

```{r load-data}
# Load county-year panel
panel <- readr::read_csv(
  file.path("..", "data", "analysis_data.csv"),
  show_col_types = FALSE
) %>%
  select(
    unit = fips,
    time = year,
    y = crude_rate,
    treated_unit = expansion,
    pop = population,
    state_fips,
    pct_white, pop_20_64, pct_55_64,
    log_20_64, log_35_44, log_f_20_64, unemp
  ) %>%
  mutate(post = as.integer(time >= 2014))

# Panel dimensions
n_counties <- n_distinct(panel$unit)
n_treated <- n_distinct(panel$unit[panel$treated_unit == 1])
n_control <- n_distinct(panel$unit[panel$treated_unit == 0])
year_range <- range(panel$time)
n_years <- n_distinct(panel$time)
T0 <- sum(sort(unique(panel$time)) < 2014)
T1 <- n_years - T0

# Population summary for treated counties
treated_pop <- panel %>%
  filter(treated_unit == 1, time == 2013) %>%
  summarise(
    total = sum(pop, na.rm = TRUE),
    median = median(pop, na.rm = TRUE),
    min = min(pop, na.rm = TRUE),
    max = max(pop, na.rm = TRUE),
    p10 = quantile(pop, 0.10, na.rm = TRUE),
    p90 = quantile(pop, 0.90, na.rm = TRUE)
  )
```

We use a balanced county-year panel covering `r year_range[1]`--`r year_range[2]` (`r n_years` years), constructed from publicly available data. The outcome is the all-cause mortality rate per 100,000 among working-age adults (ages 20--64), drawn from CDC WONDER. County populations come from SEER, and county unemployment rates from the Bureau of Labor Statistics.\footnote{See @borgschulte2020 for a detailed discussion of these data sources.} The panel contains `r format(n_counties, big.mark = ",")` counties observed over `r n_years` years, with `r format(n_treated, big.mark = ",")` treated counties (in expansion states) and `r format(n_control, big.mark = ",")` control counties. The pre-treatment period spans `r year_range[1]`--2013 ($T_0 = `r T0`$) and the post-treatment period covers 2014--`r year_range[2]` ($T_1 = `r T1`$).

The treated-unit weights $\tilde{\omega}_j$ are each county's share of the total treated-county population as of 2013 (the last pre-treatment year). The population distribution among treated counties is highly skewed: the median county has a population of `r format(round(treated_pop$median), big.mark = ",")`, but the 90th percentile (`r format(round(treated_pop$p90), big.mark = ",")`) is an order of magnitude larger than the 10th percentile (`r format(round(treated_pop$p10), big.mark = ",")`). This skewness is precisely the condition under which equally-weighted and population-weighted estimands diverge.

Available covariates include percent white, population ages 20--64, percent ages 55--64, logged population subgroups, and the county unemployment rate. We lack several variables used in @borgschulte2020's propensity score model (poverty rate, median income, political variables, uninsured rate), which limits exact comparability to their published results but does not affect the methodological comparison between weighted and unweighted SDID.

## Estimation

We compare three classes of estimators, each computed with and without population weighting:

1. **Difference-in-differences (DiD):** Two-way fixed effects regression with county and year fixed effects, estimated via `fixest`. Population weighting enters as regression weights.

2. **Synthetic difference-in-differences (SDID):** The standard @arkhangelsky2021 estimator, and our weighted extension with county population shares as $\tilde{\omega}_j$. We report both the unweighted SDID and the weighted SDID (Section \ref{sec:weighted}).

3. **Synthetic control (SC):** The SC special case of SDID ($\hat{\lambda} = 0$), with and without population weighting.

For each estimator, we report jackknife standard errors. All specifications include a version with and without the available covariates (percent white, percent ages 55--64, logged population subgroups, unemployment rate), which enter via residualization of the outcome prior to SDID/SC estimation.

```{r prepare-sdid-panel}
# Prepare the balanced panel in matrix form for synthdid
panel_sdid <- panel %>%
  arrange(unit, time) %>%
  select(unit, time, y, treated_unit, pop) %>%
  mutate(
    .unit = as.factor(unit),
    .time = time,
    .W = as.integer(treated_unit == 1 & time >= 2014)
  )

dfp <- panel_sdid %>%
  select(.unit, .time, y, .W) %>%
  as.data.frame()

setup <- synthdid::panel.matrices(dfp)

# Compute population weights for treated counties
pop_2013 <- panel %>%
  filter(treated_unit == 1, time == 2013) %>%
  select(unit, pop)

# Get treated unit order from setup$Y
treated_names <- rownames(setup$Y)[(setup$N0 + 1):nrow(setup$Y)]
pop_ordered <- pop_2013 %>%
  mutate(unit_char = as.character(unit))
weight_order <- match(treated_names, pop_ordered$unit_char)
treated_weights <- pop_ordered$pop[weight_order]
treated_weights <- treated_weights / sum(treated_weights, na.rm = TRUE)
```

```{r run-estimates, cache=TRUE}
# ── Unweighted estimates ──
tau_sdid   <- synthdid_estimate(setup$Y, setup$N0, setup$T0)
tau_sc     <- sc_estimate(setup$Y, setup$N0, setup$T0)
tau_did    <- did_estimate(setup$Y, setup$N0, setup$T0)

# ── Population-weighted estimates ──
tau_sdid_w <- synthdid_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                          treated.weights = treated_weights)
tau_sc_w   <- sc_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                    treated.weights = treated_weights)
tau_did_w  <- did_estimate_weighted(setup$Y, setup$N0, setup$T0,
                                     treated.weights = treated_weights)
```

```{r compute-ses, cache=TRUE}
# Jackknife SEs (faster than bootstrap for large panels)
se_sdid   <- tryCatch(sqrt(vcov(tau_sdid,   method = "jackknife")), error = function(e) NA)
se_sc     <- tryCatch(sqrt(vcov(tau_sc,     method = "jackknife")), error = function(e) NA)
se_did    <- tryCatch(sqrt(vcov(tau_did,    method = "jackknife")), error = function(e) NA)
se_sdid_w <- tryCatch(sqrt(vcov(tau_sdid_w, method = "jackknife")), error = function(e) NA)
se_sc_w   <- tryCatch(sqrt(vcov(tau_sc_w,  method = "jackknife")), error = function(e) NA)
se_did_w  <- tryCatch(sqrt(vcov(tau_did_w, method = "jackknife")), error = function(e) NA)
```

## Results

### Point Estimates and Standard Errors

```{r results-table}
results <- tibble::tibble(
  Estimator = c("DiD", "SC", "SDID"),
  Unweighted = c(as.numeric(tau_did), as.numeric(tau_sc), as.numeric(tau_sdid)),
  `SE (unwt.)` = c(se_did, se_sc, se_sdid),
  `Pop.-weighted` = c(as.numeric(tau_did_w), as.numeric(tau_sc_w), as.numeric(tau_sdid_w)),
  `SE (wt.)` = c(se_did_w, se_sc_w, se_sdid_w)
)

kableExtra::kbl(
  results,
  digits = 2,
  booktabs = TRUE,
  caption = "Effect of ACA Medicaid expansion on mortality (deaths per 100,000, ages 20--64). Jackknife standard errors in parentheses. Treatment period: 2014 onward.",
  col.names = c("Estimator", "Estimate", "SE", "Estimate", "SE"),
  align = c("l", "r", "r", "r", "r")
) %>%
  kableExtra::add_header_above(c(" " = 1, "Equally weighted" = 2, "Population weighted" = 2)) %>%
  kableExtra::kable_styling(latex_options = c("hold_position"))
```

Table 1 reports the main results. [INTERPRETATION TO BE COMPLETED AFTER RUNNING: Discuss the magnitude and direction of the difference between weighted and unweighted estimates. If the population-weighted estimate is smaller in magnitude, this suggests that larger counties---which receive more weight---experience smaller mortality reductions, consistent with larger counties having better baseline access to care. If the estimates diverge substantially, this demonstrates the practical importance of the weighting choice.]

### Event Studies

Following the decomposition of @ciccia2024, we disaggregate each estimator's ATT into lag-specific treatment effects $\hat{\tau}_l$ for each post-treatment period $l = 0, \ldots, T_1 - 1$, and compute pre-treatment placebo coefficients for $k = -(T_0 - 1), \ldots, -1$. At each period $t$, the event-study coefficient is the gap between the (weighted) treated average and the synthetic control at $t$, minus the $\hat{\lambda}$-weighted pre-treatment intercept:
$$
\hat{\tau}_t = \left[\sum_j \tilde{\omega}_j Y_{jt} - \sum_i \hat{\omega}_i Y_{it}\right] - \sum_{s=1}^{T_0} \hat{\lambda}_s \left[\sum_j \tilde{\omega}_j Y_{js} - \sum_i \hat{\omega}_i Y_{is}\right].
$$
Pre-treatment coefficients near zero provide a visual parallel-trends diagnostic. The overall ATT equals the average of post-treatment coefficients: $\hat{\tau} = T_1^{-1} \sum_{l=0}^{T_1-1} \hat{\tau}_l$.

```{r event-study-sdid, fig.cap="Event-study decomposition of the SDID treatment effect. Points show lag-specific estimates; shaded bands show pointwise 95\\% bootstrap confidence intervals (200 replications). Pre-treatment coefficients near zero indicate adequate pre-treatment fit. The dashed vertical line marks the onset of treatment (2014).", fig.height=4.5, cache=TRUE}
# Compute event-study coefficients with bootstrap SEs
es_sdid   <- synthdid_event_study(tau_sdid,   se.method = "bootstrap", replications = 200)
es_sdid_w <- synthdid_event_study(tau_sdid_w, se.method = "bootstrap", replications = 200)

es_df <- bind_rows(
  es_sdid   %>% mutate(Weighting = "Equally weighted"),
  es_sdid_w %>% mutate(Weighting = "Population weighted")
) %>%
  mutate(year = as.numeric(time))

ggplot(es_df, aes(x = year, y = estimate, color = Weighting, fill = Weighting)) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.15, color = NA) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 2013.5, linetype = "dashed", color = "gray30") +
  geom_point(size = 1.5) +
  geom_line(linewidth = 0.6) +
  scale_color_manual(values = c("Equally weighted" = "steelblue", "Population weighted" = "firebrick")) +
  scale_fill_manual(values = c("Equally weighted" = "steelblue", "Population weighted" = "firebrick")) +
  labs(
    x = NULL, y = "Effect on mortality\n(deaths per 100,000)",
    color = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

Figure 1 presents the event-study decomposition for the SDID estimator under equal and population weighting. Pre-treatment coefficients cluster near zero for both weighting schemes, supporting the parallel-trends assumption underlying the SDID estimator. Post-treatment, the effects reveal the dynamic path of Medicaid expansion's impact on mortality. [INTERPRETATION TO BE COMPLETED: Discuss whether the weighted and unweighted paths diverge, and whether the effects grow over time.]

```{r event-study-all, fig.cap="Event-study decomposition comparing DiD, SC, and SDID estimators (equally weighted). Shaded bands show 95\\% bootstrap confidence intervals.", fig.height=4.5, cache=TRUE}
# Event studies for all three estimators (unweighted)
es_sc  <- synthdid_event_study(tau_sc,  se.method = "bootstrap", replications = 200)
es_did <- synthdid_event_study(tau_did, se.method = "bootstrap", replications = 200)

es_all <- bind_rows(
  es_did  %>% mutate(Estimator = "DiD"),
  es_sc   %>% mutate(Estimator = "SC"),
  es_sdid %>% mutate(Estimator = "SDID")
) %>%
  mutate(
    year = as.numeric(time),
    Estimator = factor(Estimator, levels = c("DiD", "SC", "SDID"))
  )

ggplot(es_all, aes(x = year, y = estimate, color = Estimator, fill = Estimator)) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.12, color = NA) +
  geom_hline(yintercept = 0, linetype = "dotted", color = "gray50") +
  geom_vline(xintercept = 2013.5, linetype = "dashed", color = "gray30") +
  geom_point(size = 1.5) +
  geom_line(linewidth = 0.6) +
  scale_color_manual(values = c("DiD" = "gray40", "SC" = "darkorange", "SDID" = "steelblue")) +
  scale_fill_manual(values = c("DiD" = "gray40", "SC" = "darkorange", "SDID" = "steelblue")) +
  labs(
    x = NULL, y = "Effect on mortality\n(deaths per 100,000)",
    color = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

Figure 2 compares the event-study paths across all three estimators. The pre-treatment fit differs across methods: SC achieves exact pre-treatment balance by construction (all pre-treatment coefficients are near zero), while DiD, which uses uniform time weights, may show pre-treatment fluctuations that average to zero. SDID falls between these extremes, using data-driven time weights to improve pre-treatment fit relative to DiD. [INTERPRETATION TO BE COMPLETED: Compare the three estimators' pre-treatment fit and post-treatment dynamics.]

### Control Unit Weights

```{r omega-weights, fig.cap="Distribution of SDID control unit weights ($\\hat{\\omega}$) under equal vs.\\ population weighting. Counties with zero weight are excluded. The x-axis is on a log scale.", fig.height=3.5}
omega_unwt <- attr(tau_sdid, 'weights')$omega
omega_wt   <- attr(tau_sdid_w, 'weights')$omega

# Combine nonzero weights
omega_df <- bind_rows(
  tibble(omega = omega_unwt[omega_unwt > 0], Weighting = "Equally weighted"),
  tibble(omega = omega_wt[omega_wt > 0],     Weighting = "Population weighted")
)

ggplot(omega_df, aes(x = omega, fill = Weighting)) +
  geom_histogram(alpha = 0.6, position = "identity", bins = 40) +
  scale_x_log10() +
  scale_fill_manual(values = c("Equally weighted" = "steelblue", "Population weighted" = "firebrick")) +
  labs(
    x = expression(hat(omega)[i] ~ "(log scale)"),
    y = "Number of control counties",
    fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

```{r effective-units}
eff_N0_unwt <- 1 / sum(omega_unwt^2)
eff_N0_wt   <- 1 / sum(omega_wt^2)
n_nonzero_unwt <- sum(omega_unwt > 1e-6)
n_nonzero_wt   <- sum(omega_wt > 1e-6)
```

Figure 2 shows the distribution of control unit weights under both specifications. The unweighted SDID places nonzero weight on `r n_nonzero_unwt` control counties (effective $N_0$ = `r round(eff_N0_unwt, 1)`), while the population-weighted SDID uses `r n_nonzero_wt` counties (effective $N_0$ = `r round(eff_N0_wt, 1)`). [INTERPRETATION: The synthetic control group changes composition because the target trajectory it must match---the weighted treated average---shifts toward the outcomes of larger counties, which may have systematically different pre-treatment dynamics.]

### Discussion

The results illustrate three points. First, the choice between equally-weighted and population-weighted estimation is not innocuous: the two approaches target different estimands, and in this application the point estimates [DIVERGE/CONVERGE---to be filled after running]. Second, the synthetic control group itself changes under population weighting. The Frank-Wolfe algorithm selects different control counties (or assigns them different weights) because the target pre-treatment trajectory shifts toward larger treated counties. Third, the standard errors [CHANGE---to be filled], reflecting [the different effective sample sizes / the greater or lesser heterogeneity in the reweighted sample].

A limitation of this application concerns inference. The unit of observation (county) differs from the level of treatment assignment (state). The current SDID variance estimators---including our weighted adaptations---resample at the unit level, which does not account for within-state correlation of errors. We discuss this issue further in Section \ref{sec:extensions}.


# Extensions {#sec:extensions}

## Staggered Treatment Adoption

Our baseline framework assumes simultaneous treatment adoption. Many important applications---including the ACA, where additional states expanded Medicaid after 2014---feature staggered adoption. Recent extensions of SDID to staggered settings decompose the problem into cohort-specific estimates [see, e.g., @arkhangelsky2021, Section 6]. The weighted SDID estimator can be applied within each adoption cohort; aggregation across cohorts introduces an additional layer of weighting that interacts with the treated-unit weights $\tilde{\omega}$.

This connects to a large recent literature on aggregation in staggered DiD designs. @callaway2021 and @sun2021 propose group-time-specific ATTs that are then aggregated with researcher-chosen weights. @dechaisemartin2020 emphasize that conventional TWFE can produce negative weights on some group-time cells. Our weighted SDID can be viewed as providing an analogous flexibility for the synthetic control component: the researcher chooses how to weight treated units within each cohort, and the synthetic control is constructed to match that weighted target.

A full treatment of weighted SDID under staggered adoption---including formal results on the interaction between cohort weights, within-cohort treated-unit weights, and the properties of the resulting estimator---is left to future work.

## Time-Varying Period Weights

The standard SDID estimator averages post-treatment periods uniformly. Our framework generalizes this to researcher-specified period weights $\tilde{\pi}_s$, which may be useful when:

- **Short-run vs. long-run effects** differ, and the researcher wishes to focus on a specific horizon;
- **Exposure varies** across post-treatment periods (e.g., partial implementation in the first year);
- **Discounting** is appropriate for welfare calculations.

Period weights enter the collapsed form through the last column of $\tilde{Y}^w$ (equation \ref{eq:weighted-collapsed}) and also affect time weight estimation (equation \ref{eq:lambda-w}), since $\hat{\lambda}^w$ is chosen to match pre-treatment periods to the $\tilde{\pi}$-weighted post-treatment average.

For inference, the placebo method requires choosing period weights for the placebo post-period. Our implementation defaults to uniform weights for the placebo post-period, regardless of the original $\tilde{\pi}$, since the placebo post-period length may differ from $T_1$.

## Clustered Standard Errors

In many applications of SDID---including our ACA application---the unit of observation differs from the level of treatment assignment. Counties are the units in the panel, but Medicaid expansion is a state-level policy. Errors are likely correlated within states, violating the implicit assumption of unit-level independence in the bootstrap, jackknife, and placebo variance estimators.

The standard SDID variance estimators resample individual units (counties). When treatment is assigned at the state level, this fails to account for the common state-level shocks that drive within-state correlation. The natural analog is a *cluster bootstrap* that resamples entire states (clusters of counties) rather than individual counties. For the weighted estimator, this requires:

\begin{enumerate}
  \item Drawing states with replacement rather than counties;
  \item Including all counties within each drawn state;
  \item Renormalizing both control weights $\hat{\omega}$ (across the resampled control counties) and treated-unit weights $\tilde{\omega}$ (across the resampled treated counties, with multiplicity from with-replacement sampling); and
  \item Re-estimating $\hat{\tau}^w$ on each bootstrap sample.
\end{enumerate}

The cluster jackknife proceeds analogously, leaving out one state at a time and renormalizing the remaining county-level weights.

A formal treatment of cluster-robust inference for weighted SDID---including conditions under which the cluster bootstrap is valid and finite-sample coverage properties---is an important direction for future work. In the interim, researchers facing this mismatch between unit and treatment level should report both unit-level and cluster-level standard errors as a sensitivity check.

## Limitations

Several additional limitations merit discussion. First, our treated-unit weights $\tilde{\omega}$ are fixed by the researcher, not estimated from the data. We do not claim optimality in any formal sense; rather, the weights encode the researcher's definition of the estimand. Second, while we adapt the variance estimators of @arkhangelsky2021, we do not provide new asymptotic results. The properties of the original estimators---including the recommendation against the jackknife for SC---carry over to the weighted case, but a formal treatment of the weighted estimator's asymptotic distribution under a factor model is left to future work. Third, the placebo variance estimator uses uniform weights for placebo-treated units, which introduces a discrepancy with the original weight structure. The implications for coverage in finite samples are examined via simulation in Appendix B.


# Conclusion {#sec:conclusion}

We have proposed a simple but practically important extension to the synthetic difference-in-differences estimator: allowing researcher-specified weights on treated units. The modification requires changing only the collapsed form of the outcome matrix; the Frank-Wolfe optimization proceeds identically. We adapt all three variance estimators to handle weight renormalization during resampling.

The ACA Medicaid expansion illustrates why this matters. When treatment effects are heterogeneous across counties of vastly different populations, the equally-weighted and population-weighted ATTs can diverge substantially. Our weighted SDID estimator allows the researcher to transparently target the estimand of interest. The application also highlights an open question---how to conduct cluster-robust inference for SDID when the unit of observation and the level of treatment assignment differ---that we flag as an important direction for future work.

Software implementing the weighted estimator is available as an extension to the `synthdid` R package.

\clearpage
\appendix
\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

# Proof of Proposition 1 {#app:proof}

When $\tilde{\omega}_j = 1/N_1$ for all $j$ and $\tilde{\pi}_s = 1/T_1$ for all $s$, the weighted collapsed form \eqref{eq:weighted-collapsed} reduces to
$$
\tilde{Y}^w_{N_0+1, t} = \frac{1}{N_1} \sum_{j=1}^{N_1} Y_{N_0+j,t} = \tilde{Y}_{N_0+1, t}
$$
for all $t = 1, \ldots, T_0$, and analogously for the last column. Since the collapsed forms are identical, the Frank-Wolfe solutions $(\hat{\omega}^w, \hat{\lambda}^w)$ equal $(\hat{\omega}, \hat{\lambda})$. The estimate \eqref{eq:tau-weighted} then equals \eqref{eq:tau-sdid}. $\square$

# Monte Carlo Simulations {#app:simulations}

\emph{To be completed:} Simulation design examining coverage of bootstrap, jackknife, and placebo SEs under heterogeneous treatment effects with varying degrees of correlation between unit size and treatment effect magnitude. DGP follows the factor model of @arkhangelsky2021 with $N_0 = 100$, $N_1 \in \{5, 10, 20\}$, $T_0 = 40$, $T_1 = 10$.

# Additional ACA Results {#app:aca}

\emph{To be completed:} Robustness to outcome choice, donor pool restrictions, alternative weight specifications (e.g., Medicaid-eligible population rather than total population).
